### 2010-10-29 Paper_1.tex
Initial commit
	new file:   .gitignore
	new file:   Paper_1.tex

\documentclass[manuscript]{aastex} % one-column, double-spaced document
%\documentclass[preprint2]{aastex}  %% double-column, single-spaced document
% \documentclass[12pt,preprint]{aastex}
% \documentclass[aaspp4]{aastex}
% \documentclass[aas2pp4]{aastex}
% \documentclass[aj_pt4]{aastex}
% \documentclass[apjpt4]{aastex}
% \documentstyle[12pt,aasms4]{article}
% \documentclass[12pt]{article}
%\documentstyle[12pt]{article}
% \tighten
%
%
\newcommand{\ie}{{\it i.e.}}
\newcommand{\cf}{{\it cf.}}
\newcommand{\eg}{{\it e.g.}}
\newcommand{\etal}{{\it et al.}}
\newcommand{\rd}{{\rm d}}
\newcommand{\ag}{\mbox{ \raisebox{-.8ex}{$\stackrel{\textstyle >}{\sim}$} }}
\newcommand{\al}{\mbox{ \raisebox{-.8ex}{$\stackrel{\textstyle <}{\sim}$} }}
\newcommand{\vdag}{(v)^\dagger}
\newcommand{\myemail}{tomsc@astro.indiana.edu}
\newcommand{\mso}{$M_{\odot}$ }
\newcommand{\rso}{$R_{\odot}$ }
\newcommand{\msoyr}{$M_{\odot}/{\rm year}$}

\newcounter{subequation}

\usepackage{color}
\usepackage{graphicx}
\usepackage{amsmath, amsthm, amssymb}

\citestyle{aa}

% \slugcomment{Not to appear in Nonlearned J., 45.}

\shorttitle{Effective Disk Alphas}
\shortauthors{Scott et al.}

%
\begin{document}

\title{Convergence Studies and Effective Alphas for Disks with Gravitational Instabilities.
 \\
I. The Constant Cooling Case}
\author{Scott Michael}
\affil{Astronomy Department, Indiana University, Bloomington, IN 47405}
\email{scamicha@astro.indiana.edu}

\author{Thomas Y. Steiman-Cameron}
\affil{Astronomy Department, Indiana University, Bloomington, IN 47405}
\email{tomsc@astro.indiana.edu}

\author{Richard H. Durisen }
\affil{Astronomy Department, Indiana University, Bloomington, IN 47405}
\email{durisen@astro.indiana.edu}

\and

\author{Aaron Boley}
\affil{Department of Astronomy, University of Florida, Gainesville, FL 32611}
\email{aaron.boley@gmail.com}

\begin{abstract}

Using 3D hydrodynamic simulations of a cooling protostellar disk undergoing gravitational instabilities (GIs), we compute the effective Shakura and Sunyaev (1973) $\alpha$ due to gravitational stresses and compare them to predictions from an analytic local theory for thin disks by Gammie (2001).  Our goal is to determine how accurately a locally defined alpha can characterize mass and angular momentum transport by GIs in disks.  We examine disks with an imposed constant global cooling time (Mej\'{i}a et al. 2005).  Models were computed with grids of different spatial resolution to investigate how the computed alpha is affected by numerical resolution.

\end{abstract}

%\keywords{ }

%%%%%% INTRODUCTION %%%%%%%%%%%%%%%
\section{Introduction}

While detailed three-dimensional hydrodynamic modeling of protostellar disks can
provide considerable insight into the physical and thermal states of these disks,
computing resources limit such studies to snapshots
spanning timescales much shorter than a disk's evolutionary lifetime.  In contrast, 
approaches that replace detailed physics with statistical parameters that
encapsulate this physics allow for calculations covering
much longer timeframes.  Both approaches can play important roles in broadening our 
understanding of these complex nonlinear systems, provided a connection between the detailed physics 
and statistical parameters can be demonstrated and understood.   In what follows, we
examine the applicability of an $\alpha$-disk formulation (Shakura \& Shaviv 1973) to  the
evolution of protostellar disks subject to gravitational instabilities.

Gravitational instabilities (GIs) play an important, and at times dominant,
role in driving the structural and thermal
evolution of  protostellar disks \citep[for a review, see][]{Dur10}.  
%In particular, torques arising from GIs provide a very effective means of redistributing matter and 
%angular momentum within these disks 
%\citep{Lod04, Lod05, Vor05, Vor06, Bol09}.  
Thermal processes play the primary role in regulating the
amplitude and outcome of these instabilities  (Pickett
et al. 1998, 2000a,b, 2003, see also Nelson et al. 1998, 2000). 
A link between a disk's thermodynamic state and its susceptibility to GIs
is provided by the Toomre $Q$-parameter (Toomre 1981);  $Q = c_s \kappa / \pi G \Sigma$,
where $c_s$ is the sound speed, $\kappa$ is the epicyclic frequency ($\sim$ the rotation frequency 
$\Omega$ in a nearly Keplerian disk), and $\Sigma$ is the disk surface mass density. 
For $Q \al$  1.5 to 1.7,  small density perturbations in a disk
grow exponentially on a time scale comparable to the rotation period.
These perturbations manifest themselves as multi-arm spirals with a predominantly trailing pattern that
transports angular momentum outward by
gravitational torques (e.g.,Larson 1984, Boss 1984, Durisen et al. 1986, Papaloizou \& Savonije
1991, Laughlin \& Bodenheimer 1994, Laughlin et al., 1998; Nelson
et al., 1998; Pickett et al., 1998, add refs?).
The long range nature of these torques cause nonlinear
spiral structures to develop over a broad range of radii (Laughlin et al. 1998, Nelson et al. 1998,
2000; Pickett et al. 1998, 2000, 2003).

Numerous studies
have shown that thermal physics can control GI amplitudes
by a {\it balance} of heating and cooling (e.g., Tomley et al. 1991, 1994;
Pickett et al., 1998, 2000,
2003; Gammie, 2001; Boss, 2002; Rice et al., 2003; Mej\'{i}a et al., 2005; Boley
et al., 2006, 2007a; Stamatellos and Whitworth, 2008; Cossins et al., 2009).
Heating by GIs tends to increase $c_s$, thus increasing $Q$. If disk heating persists,
the instability will shut off once $Q$ becomes sufficiently large.  However, if the
disk cools radiatively, then heating and cooling can reach a balance at nonlinear
wave amplitude. GI-activity can sustain this balance  at a relatively constant,
but unstable, value of $Q$ (e.g., Paczynski, 1978; Lin and Pringle; Pringle, 1981;
see also Goldreich and  Lynden-Bell 1965).  
For sufficiently low cooling rates, the disk can exist in a state of quasi-equilibrium, 
where cooling is 
balanced by heating arising from the inward flow of matter and shocks induced by the GIs
\citep{Gam01, Lod04, Ric05, Cos09, Vor10}.

Fully nonlinear hydrodynamic simulations of low-$Q$ disks show that multiple
modes can become unstable in the linear regime (e.g., Nelson et al.,
1998; Pickett et al., 1998; additional refs?).  Even if the initial growth is dominated by
one mode, numerous modes usually appear in the nonlinear regime. As a consequence, 
if a disk achieves a quasi-steady balance between heating and cooling, 
then there tends to be power at
all $m$-values resolved by a given numerical method (e.g., Mej\'{i}õa et al., 2005;
Boley et al., 2006).  
In this  {\it asymptotic state}, GI-active disks develop overlapping
density structures of different strengths, geometries and coherence,
producing
strongly nonlinear turbulence that pervades the entire disk
(e.g., Pickett et al. 2003, Mej\'{i}a et al. 2005, Boley et al. 2006).
Gammie (2001) refers to this state as {\it gravitoturbulence}.  
Gravitoturbulence provides a possible bridge between
the detailed physics of GIs and the viscous transport parameter $\alpha$.

Shakura \& Syunyaev (1973) first parameterized turbulent transport in a steady-state
disk with the  parameter $\alpha$, where the source of the
turbulence is unspecified.
Although their original work was directed towards X-ray bright
accretion disks around black holes, this heuristic
approach has proved fruitful in many 
astrophysical situations by allowing for
analytic solutions and relatively easy numerical modeling
over time-spans inaccessible to detailed modeling.
In the $\alpha$ disk formalism, when the accretion rate
$\dot{M}$ is constant in both time and radius, $\dot{M} = 
3\pi\alpha c_s^2 \Sigma / \Omega$ where $c_s$ is the sound speed, $\Sigma$ the 
surface density, and $\Omega$ the rotational frequency. 
The $\alpha$ 
formalism makes no assumption
about the nature of the angular momentum transport, only that {\it heating and cooling
are dominated by local processes}; the validity
of an $\alpha$-disk picture requires self-consistency of local energy dissipation.

Dissipative locality in GI-active disks has not been rigorously tested and 
remains an open question. 
Several studies support the idea that GI
transport in real disks is, in many important respects, an intrinsically global phenomenon and
cannot be properly treated by a local $\alpha$-like prescription (Laughlin \& Rozyczka 1996, Balbus
\& Papaloizou 1999, Lodato \& Rice 2005, Mej\'{i}a et al. 2005, Boley et al. 2006, Cai et al.
2007).  This almost certainly applies to the embedded phase of both high and low-mass
stars when infall from the collapsing protostellar cloud feeds mass into the disk at a rapid
rate (e.g., Yorke et al. 1993, Laughlin \& Bodenhiemer 1994, Yorke \& Bodenheimer 1999,
Vorobyov \& Basu 2005, 2006, Kratter \& Matzner 2007), and disks observed in the embedded
phase can be quite massive relative to their central stars (e.g., Osorio et al. 2003). 
However, despite the long-distance nature of gravitational torques, theoretical
and numerical results suggest that in protostellar disks where heating and cooling are in
balance it may be valid to represent GI-induced transport by an {\it effective viscosity}, a local process
(e.g., Gammie 2001; Rice et al. 2005; Lodato \& Rice 2004).
Indeed, Gammie (2001) derived an effective alpha for the case of a thin (effectively 2D)
gravitationally unstable disk where the heating due to local turbulent dissipation caused by
GIs was balanced by the local cooling rate.  Modifications to this derivation have since been
proposed by {\bf (Tom: include other derivations and refs)}.   Recent simulations (e.g., Cossins
et al. 2009; Vorobyov 2010) suggest that, at least in the case of imposed cooling,
$\alpha_{eff}$ roughly converges to the value predicted by Gammie.

{\bf (Tom: rework and finish this paragraph, or toss it)}
Three possibilities exist: 1) GI-transport is predominately a long-range effect and thus not 
amenable to an $\alpha$-disk representation, 2) GI-transport can be well represented as
a short-range effect and thus is directly amenable to an $\alpha$-disk representation,
If long-range energy transport and/or angular momentum transport are negligible,  or if
the long-range transport could be shown to represent a known fraction of the 
***. 
Identifying the principal angular momentum transport mechanism
and how it behaves is critical to understanding whether disks can be described by an alpha disk model.

{\bf (Tom: add text concerning need for convergence approach)}
Here we report the results of a convergence study designed to determine
whether GI-active disks, chaotic by nature, achieve statistical equilibrium; if so, to what
level do these disks approximate the gravitoturbulence of Gammie (2001) and with what distribution
of $\alpha_{eff}$; and to characterize how the GI cascades from low-order to high-order modes. 
To this end, we follow the evolution of a GI-active disk, subject to constant cooling, in its asymptotic state,
where cooling and heating are essentially balanced.  Simulations are
performed using four different resolutions in the azimuthal direction.

{\bf (Tom: add paragraph providing outline of paper)}

%Nonlinear evolutions of disks undergoing GIs, even local thin-disk treatments, 
%\citep[e.g.,][]{Joh03} 
%are sufficiently complex that they must be done numerically,  
%typically using a finite element approach, where
%a set of coupled equations is solved simultaneously on a computational grid.  
%Computational constraints prohibit a full examination of GI unstable disks over 
%the life of the disk.
%Indeed, existing computational resources limit 
%detailed three-dimensional 
%simulations of disk evolution to tens of
%outer disk rotations.  
%We cannot yet follow the entire dynamical evolution of a 
%protoplanetary disk nor
%model all the radiative and particle processes of interest at once in full 3D. However, 
%by following GI simulations to their asymptotic states, when these exist, we
%can extrapolate the consequences over much longer times (e.g., Boley et al 2006).

%For 
%grids of a fixed {\it physical} size,
%grids with more elements provide finer meshes, which typically produce a more accurate solution. 
%However, this increased accuracy if offset by temporal sacrifices.
%The number of equations to be solved at each timestep 
%increases as the grid size is increased.  At the same time, the Courant condition 
%forces these timesteps to be shorter in duration.  A mesh convergence study 
%provides a common approach to balancing accuracy and computing resources.
%In the following sections we follow the evolution of a disk subject to GUs
%using four different
%resolutions in the azimuthal direction.  The number of azimuthal grid points used in simulations
%may be particularly important, because higher resolution 
%in azimuth allows GI power to spread to higher-order modes 
%that behave more locally. We test this by examining the power in various 
%modes to see how they are affected by the choice of grid.

\section{Hydrodynamical Simulations}

Simulations were conducted with the three-dimensional hydrodynamics 
code used by the
Indiana University Hydrodynamics Group in several previous studies
\citep{Pic98, Pic00, Mej04, Mej05, Bol06, Cai08}.  This code
uses a second-order, explicit, Eulerian scheme to solve Poisson's equation, 
an equation of state,
and the equations of hydrodynamics in conservative
form on a uniform cylindrical grid.  The code assumes mirror symmetry about the equatorial plane.
It includes both
self-gravity and artificial viscosity; the latter serves to 
mediate shocks.
Source and flux terms (Norman \&Winkler 1986) are computed separately in an
explicit, second-order time integration (van Albada et al. 1982; Christodoulou 1991; Yang 1992).

We assume an equation of state $P = (\gamma -1)\epsilon$, 
where $P$ is the pressure, $\epsilon$ is the internal energy density, and $\gamma$ 
is the ratio of specific heats, given here by $\gamma = 5/3$.  
The model disks are cooled by decreasing their internal energy according to 
the prescription $d\epsilon/dt = \epsilon / t_{cool}$, where $t_{cool}$ is the global cooling time in Outer Rotation Periods (orp).  For the grid resolutions used here, one orp is defined as the rotational period at radial zone 200.

%\subsection{Models}

Because this is a polytropic disk with an idealized cooling time, only the ratio of disk to stellar mass, given here by 0.153, is specified.  In what follows, quantities specified with real physical units are derived assuming a 1 $M_\odot$ central star and a 0.153  $M_\odot$ disk.   In this case, the orp is defined at $r = 33$ AU and
one orp $\approx$ 303 years.  Models have $t_{cool} = 2$ orp.   {\bf (Scott: Please check these numbers)}
The disk has
an initial surface density $\Sigma \sim r^{ -1}$, and initial inner and outer radii of 2.3 and 40 AU, respectively.

Initial conditions for the disk's structure and thermodynamic state
were set using an equilibrium star plus disk model generated using a modified Haschisu (1986)
self-consistent field (SCF) relaxation method (Pickett et al. 1996, 2003;
Mej\'{i}a (2004);  Mej\'{i}a et al. (2005); Cai 2006).  
Here  the density and angular
momentum distributions are iteratively
solved, using the specified equation of state,  until convergence is achieved. 
This  equilibrium disk was then seeded with small random perturbations to
allow GIs to grow.  The resultant marginally unstable disk served as the $t = 0$ load for the hydro code.

The initially unstable disk model passed through several phases of evolution
(see also Pickett et al., 2003; Mej\'{i}a et al., 2005).  The initial load is unstable to
the growth of axisymmetric structures. During this axisymmetric phase, 
which lasts several orps,
the disk shrinks slightly, in an axisymmetric manner, until it becomes unstable
to nonaxisymmetric perturbations.    When this condition is met, the disk undergoes a strong burst 
of GIs that predominantly manifest themselves in
one or two discrete global spiral modes.  The disk expands violently during the burst phase,
producing a significant rearrangement of the disk's mass distribution on a timescale of one or
two orps. 
This is then followed by a period of several orps where heating temporarily washes
out some of the nonaxisymmetry in the disk.  Finally the disk settles into
a quasi-steady, long-lived asymptotic state of sustained GI-activity
over a large part of the disk, with an overall balance of heating and cooling.

Simulations were run with four different angular resolutions, with grid sizes in $(r,\phi,z)$
of  $(j_{max}, k_{max}, l_{max}) =$
 (512,64,64), (512,128,64), (512,256,64), and (512,512,64).
Because of the large computational overhead associated with starting each simulation from
$t = 0$,
the $l_{max} =$ 128 simulation was run through the axisymmetric, burst and transition
phases. Simulations using $l_{max} =$ 64, 256, and 512 
all begin by interpolating the $l_{max} =$ 128 simulation
to higher or lower azimuthal resolution at 9.6 ORPs, near the end of the transition
phase. Although this strategy saves a large amount of computational time, it limits
the scope of the analysis to the asymptotic phase and
it introduces a perturbation to the disk which takes several
orps to subside. By 12 orps, the perturbation has dissipated and all four disks
have transitioned to the asymptotic state.  All simulations were 
run through $\sim 18$ orps.

Figure 1 shows $Q$, as a function of radius, at $t \sim$ 18 orp, while
Table 1 lists $Q_{avg}$, the value of $Q$ of for the asymptotic disk, time-averaged of
12 -- 19 orp and spatially averaged over 10 -- 40 AU, as a function of radius
for each resolution, 
{\bf (Scott: Figure 1 needs to be redone - I have extracted the 
plot from your chapter 3 and it only shows the 128 model.  We should show all four resolutions
or ditch this figure)}.

%%%%%%%%%%%% RESULTS & DISCUSSION %%%%%%%%%%%%%%%%

\section{Results and Discussion}

In the asymptotic state, GI-active disks develop complex
density structures arising from the superposition of
all modes present in the disk.  Since the number of
modes accessible
to the disk is a function of the azimuthal resolution,
we expect that
higher resolution simulations will manifest more power
in higher-order modes 
that behave more locally. We test this by examining the power in various 
modes to see how they are affected by the choice of grid.

\subsection{Modes}

Figure 2
%\ref{fig:DensityPlots}, 
shows midplane and radial densities  at $t \sim 18$ orps  
for the $l_{max} =$ 64, 128, 256, and 512 simulations. 
The lower resolution (smaller $l_{max}$) simulations
display longer wavelength spiral structures, with larger amplitudes and more coherence than those
seen in higher resolution 
simulations.  
Higher angular resolution
allows GI power to spread to higher-order azimuthal modes, as demonstrated 
by a Fourier decomposition of the density.
The strengths of these modes are given by their global Fourier amplitudes,
\begin{equation}
A_m = \frac{(a_m^2 + b_m^2)^{1/2}}{\pi\int\rho_o \varpi {\rm d}\varpi {\rm d}z},
\end{equation}
where
\renewcommand{\theequation}{\arabic{equation}\alph{subequation}}
\setcounter{subequation}{2}
\begin{equation}
a_m = \int \rho \cos(m\phi)\varpi {\rm d}\varpi {\rm d}z {\rm d}\phi,
\end{equation}
\addtocounter{equation}{-1}
\addtocounter{subequation}{1}
\begin{equation}
b_m = \int \rho \sin(m\phi)\varpi {\rm d}\varpi {\rm d}z {\rm d}\phi.
\end{equation}
\renewcommand{\theequation}{\arabic{equation}}
Here  $\rho_o$ is the axisymmetric (zero-order)  component of the density.
In what follows, $A_m$ is averaged over time to 
suppress fluctuations on the dynamic time scale due
to the chaotic nature of GIs in the asymptotic  phase.
In particular, $\langle A_{tot} \rangle$ will represent 
the power over the radial range 10 -- 40 AU,
integrated from $ m = 2$ to $l_{max}/2$ and
averaged over the time range 12 -- 19 orps.  
The density distribution cannot be resolved in azimuth
on angular scales less than $2\pi/l_{max}$, therefore $A_m$ is
limited to modes $m \le l_{max}/2$.  
Because the central star is artificially fixed to the grid center, the $m = 1$ mode 
cannot be be realistically treated; it is excluded from
 $\langle A_{tot} \rangle$.

{\bf (Note to all: Throughout the paper I use integrals in places where we 
formally have used summations.  I note that this is a common practice, though
not strictly correct.  I find the use of integrals more elegant, but am willing to
revert to summations if people push for it.)}

Previous studies have shown that disks in the asymptotic state have power at
all resolvable $m$-values  (e.g., Mej\'{i}õa et al., 2005;
Boley et al., 2006).  Table 1 lists $\langle A_{tot} \rangle$ along with 
$\langle A_{2-7} \rangle  /  \langle A_{tot} \rangle$,  the fraction of the total
power falling into low-order ($m = 2$ -- 7) modes, for each resolution. 
While the total power is approximately independent of
resolution, $\langle A_{tot} \rangle = 1.90 \pm 0.07$, 
the modal distribution of this power shows marked differences.
In particular, the fractional power in lower order modes decreases monotonically
with increasing resolution.  Over three quarters of $\langle A_{tot} \rangle$ resides
in modes $m =$ 2 -- 7 for the $l_{max} = 64$ simulation, while approximately
half resides in these modes when $l_{max} =$ 512.

The shift of power from lower to higher order modes as the resolution of
the computational grid is increased arises from the fact that, while
$\langle A_{tot} \rangle$ is approximately independent of $l_{max}$, 
the number of degrees of freedom available for this power, $l_{max}/2$, 
increases linearly with the resolution.  
Thus power is naturally spread from lower order to higher order modes
as the available $m$-values increases.  
Indeed, even if the disk is dominated by one or two low-order modes at the
time of outburst, this power cascades to higher-order modes (e.g., 
Laughlin et al. 1997, 1998; see also Laughlin and Korchagin 1996).

Given the resolution dependence of how power is
distributed,  we must examine 
what resolution is sufficient to properly capture the evolution of these disks.
Table 1 shows that while the fractional power in lower order modes
decreases with each doubling of $l_{max}$, the amount of change in going from $l_{max} = 256$ to 512 
is significantly less than the change going from $l_{max} = 64$ to 128,
with a strong suggestion
that this diminution will not proceed much further with higher azimuthal resolution.  
Indeed, a  Richardson extrapolation {\bf (add reference)} 
of the tabulated values of  $\langle A_{2-7} \rangle / \langle A_{tot} \rangle$
suggests a limiting value of  $\langle A_{2-7} \rangle / \langle A_{tot} \rangle \sim$ 0.43  
for increasingly higher resolutions.
This is displayed graphically in Figure 4, where the
fractional power in lower order modes is plotted against the  normalized grid 
spacing for each value of $l_{max}$.
The grid spacing  is normalized such that spacing of the $l_{max}$ = 512 grid
equals unity.  Convergence occurs as the normalized spacing goes to zero.
The Richardson Extrapolation value  of $\langle A_{2-7}\rangle  / \langle A_{tot}\rangle  = 0.43$ is
plotted at a normalized grid spacing  equal to zero.   More than half of the modal power in
a converged disk resides in higher-order modes, modes 
that help define a state of gravitoturbulence.
The fact that
the $l_{max} =$ 512 simulation has
$\langle A_{2-7} \rangle / \langle A_{tot} \rangle$ near the convergence value
lends some confidence in the use of the $l_{max} =$ 512 model in
examining the convergence of $\alpha$.

\subsection{Gravitational Torques}

Coherent density structures in a disk produce gravitational torques.  
Low-order modes in $\Delta\rho / \rho$  have longer wavelengths and, if coherent in
radius, have longer lever arms to produce torques. On the other hand, high-order
modes have relatively short wavelengths which, particularly if they lack coherence,
produce more localized effects. The high-order
modes tend to produce pockets of local GI activity, which can cancel each other
over the whole disk, leading to the expectation that  gravitational torques in the disk 
will decrease as $l_{max}$ is increased.    This can be tested by examining the
torque contributions from  various Fourier modes.

The torque {\bf C} on a cylindrical surface of the disk at radius $\varpi$
can be obtained by integrating the stress tensor $T$ over the
surface of the cylinder (e.g., Lynden-Bell \& Kalnajs 1972), i.e.,
\begin{equation}
{\rm {\bf C}} = \int {\bf r} \times T \cdot {\rm d}S.
\end{equation}
If the stress tensor includes only gravitational stresses,
the surface integral can be replaced with the volume integral
\begin{equation}
{\rm \bf C} = \int \rho {\bf r} \times \nabla \Phi {\rm d}V,
\end{equation}
where $\Phi$ is the gravitational potential.  Here we are interested
only in the $z$-component of torque,  
\begin{equation}
\label{torque1}
{\rm \bf C}_Z = \int \rho \frac{\partial\Phi} {\partial \phi} {\rm d}V,
\end{equation}
as only this component  
drives mass and angular momentum transfer.  
The torque contribution from each mode $m$ can be calculated 
by replacing $\rho$ in Equation (5) with the corresponding modal density component
\begin{equation}
\rho_m = \slantfrac{1}{2} \rho_o + a_m \cos(m\phi) + b_m \sin(m\phi),
\end{equation}
where $a_m$ and $b_m$ are provided by Equation 2,
and only the gravitational potential produced by the mass distribution
given by $\rho_m$ is included in $\Phi$.
{\bf (Scott: Equation (6) is the form you use in your chapter 3, but it
does not appear to be correct; it does not yield a correct value for $m=0$)}

Figure 5 displays $\langle {\rm \bf C}_{Ztot} \rangle$, the time-averaged gravitational 
torque summed over all disk modes
(excluding $m=1$),
along with the torques summed over low-order modes,
$\langle {\rm \bf C}_{Z(2-n)} \rangle$, where $n = $ 2, 3, 4, and 5, as a function of radius
and azimuthal resolution. 
{\bf (Note to all: the version of Figure 5 included with this draft  shows the torques
time-averaged over for four different time-frames.  I did this because I wanted to see if the disk was 
"settling down" with time.  I think the figure suggests  that the 12-14 orp disk is not
yet in an asymptotic state - or further from it than at later times - I'm not sure this is
a significant difference.  Looking forward to Figure 7, it can be argued that the
disk is reasonably settled, in a statistical sense, shortly after 12 orp.
Dick has told me he only believes the 12-18 orp result - Dick, I hope 
I'm not misquoting you).  
} 
The scale of the $y$-axis in each subplot 
is the same, with the exception of the $l_{max} = 64$ simulation. 
In this case the scale is doubled.  Two features are immediately apparent:
1) the total gravitational torque is smaller for higher resolution simulations,
and 2) the total torque is dominated by low-order modes.
Both reflect the fact that a while higher fraction of the total power in
larger $l_{max}$ simulations is in higher order modes, these modes make only
minimal contributions to the gravitational torque.
As resolution increases,
non-axisymmetric power is shifted from low-order global modes, 
which produce relatively large gravitational torques,
to high-order local modes, which tend to cancel each other out when integrated over the
whole disk.

\subsection{Effective Alphas}

An $\alpha$-framework treats disks as essentially two-dimensional objects
subject to local energy dissipation.  Therefore, in order
to compare our results with predicted 
and published results, we must extract two-dimensional information from our 
three-dimensional disks.  Following Lodato \& Rice (2004), 
\begin{equation}
\alpha_{eff}(\varpi) = \left| \frac{{\rm d} \ln \Omega}{{\rm d} \ln \varpi} \right|^{-1} 
\frac {\langle T_{\varpi\phi}^{grav}\rangle + \langle T_{\varpi\phi}^{Reyn}\rangle}
{\Sigma c_s^2},
\end{equation}
where $\Omega$ is the azimuthally averaged rotation speed, 
$T_{\varpi\phi}^{grav}$ and $T_{\varpi\phi}^{Reyn}$ represent the 
azimuthally averaged
$\varpi,\phi$-component of the specific (per unit mass)
gravitational 
(or ``Newton'') and hydrodynamic (or ``Reynolds'') stress tensors, respectively,
$\Sigma$ the surface density,
and $c_s$ the azimuthally averaged midplane sound speed.  Brackets
surrounding the stress terms indicate that these are integrated over
the $z$-direction.
The denominator $\Sigma c_c^2$ was evaluated as $\int_z \rho c_c^2 {\rm d}z$.
{\bf (Scott: please provide some text explaining why this choice was made for
${\Sigma c_s^2},$)}

Numerical uncertainties render the Reynolds stress difficult to determine. 
Fortunately, this difficulty is mitigated by the fact
that previous studies have shown the
magnitude of the Reynolds stress to be small relative to the gravitational stress
(references?).
Thus it can be safely neglected in Equation (7).

The gravitational torque from equation \ref{torque1} can be used to determine the
gravitational stress by
\begin{equation}
\begin{split}
 T_{\varpi\phi}^{grav}  {\varpi}& = \frac{1}{2\pi\varpi^2}C_Z
\\
                                & = - \frac{1}{2\pi\varpi^2} \int \rho\frac{\partial\Phi}{\partial\phi} {\rm d}V.
\end{split}
\end{equation}
{\bf (There is a problem with this equation)}
Figure 6 shows the effective $\alpha$s for the four different grid resolutions, time
averaged from 12 to 19 orps.  
For comparison, the predicted values by Gammie (2001)
for the strongly self-gravitating limit (lower curve) and the non-self-gravitating limit
(upper curve) are shown. Note that predictions from are based upon a thin
disk approximation, where the local balance of heating and cooling dominates the
energetics.

,

In an $\alpha$ disk formalism the mechanism giving rise to the transport of angular
momentum also acts to dissipate energy. Lodato \& Rice (2005) postulated that a
disk in a state of gravitoturbulence would have locally balanced heating and cooling
rates. If the heating were due to local dissipation, then $\alpha$ and $t_{cool}$ are related by
\begin{equation}
\alpha = \left| \frac{{\rm d}\ln \Omega}{{\rm d} \ln R} \right|^{-2}
\frac{1}{\gamma(\gamma - 1) t_{cool}\Omega}.
\end{equation}
Rice et al. (2005) concluded that for a given $t_{cool}$  there exists some maximum stress
quantified by $\alpha_{max}$, which if exceeded would cause a disk to fragment. The horizontal
dot dashed line in figure 6 indicates the $\alpha_{max} \approx 0.06$ found by Rice et al. (2005).
Clarke et al. (2007) have revised this limit to $\alpha_{max} = 0.12$ for cases where $t_{cool}$  varies
with time.

Figure 7 shows the time-variation of $\alpha$, normalized to its value at $t = 18$ orp), as a function of 
radius and  time.  This plot suggests that the disk has not yet achieved statistical equilibrium until 
$\sim $14 orp {\bf (Everyone:  What do you think?  We are using averages over12-18 orp,
are we comfortable with those limits?)}

\begin{center}
\line(1,0){250}
\end{center}

{\bf From this point on the draft is unfinished; the text that follows is extracted from 
Scott's Chapter 3, with little to no modification, so for now consider it as filler.   
I am still working on the rest of this section, which I expect to be the concluding section
of the paper (excluding a summary).  Rather than further delay getting
the draft to you, I am sending it as is.   I will be adding text and incorporating results from a 
few recent papers and will send the rest soon.}

\begin{center}
\line(1,0){250}
\end{center}

As the azimuthal resolution increases, the effective $\alpha$ in the GI active region
decreases. This is due to the decrease in the total gravitational torque, as illustrated
in Figure 5. This decrease in gravitational torque is due, in turn, to the shift of
non-axisymmetric power from low-order global modes, which produce large torques,
to high-order local modes, which can cancel each other out when integrated over the
whole disk. In fact, the decrease in the ratio of low-order power to the total power
$\langle A_{2-7}\rangle / \langle A_\Sigma \rangle$ 
closely tracks the decrease in the time averaged $\alpha$ averaged from 10 to
40 AU, which can be seen in Table 1 as the $\alpha_{avg}$ value. Although the 
average $\alpha$ values
have not converged to a single value for my highest two resolutions, the percentage
change continues to decrease. In addition, Figure 6 shows that the $l_{max} =$ 256
and 512 simulations have qualitatively similar time averaged profiles over a large
range of disk radii. These profiles also fall within the predicted Gammie curves, with
tendency toward the lower (strongly self-gravitating) curve. Indeed the $\alpha_{avg}$
value
of 0.020 for the $l_{max} = $ 512 simulation is squarely between the $\alpha_{avg}$ values for the
Gammie curves 0.014 (strongly self-gravitating) and 0.027 (non-self-gravitating).

Although the effective $\alpha$ curves do correspond to the Gammie prediction, the
results from these simulations differ from Gammie's in one significant way. The local
shearing box simulations performed by Gammie (2001) show that the total 
effective $\alpha$ is composed of a gravitational stress component and a Reynolds stress component
that are nearly equal (see Gammie (2001) Figure 3). However, all of the analyses
presented here only consider the gravitational stress. When I attempt to measure
the Reynolds stress, I find it is quite small compared to the gravitational stress. As
noted previously, the Reynolds stress is very hard to determine in global simulations.
However, the prediction made by Gammie relies only on the idea that heating from
GI activity balances cooling. So the exact mix of stresses contributing to the total
amount of heating could be different depending on the type of simulation, the physics
considered in the simulation, and the simulation parameters.

\section{Summary}

When considering the effect of azimuthal resolution on the asymptotic behavior
of a GI unstable disk I found that without sufficient azimuthal resolution of at least
$l_{max} =$ 512 gravitational torques and effective gravitational $\alpha$
will be underestimated
for disk with a constant global cooling rate. I also found that, when properly resolved,
more non-axisymmetric power propagates to higher order modes, and the amplitude
of low-order modes decreases. This results in a decrease of the gravitational torque
and effective gravitational $\alpha$. For $l_{max} =$ 512 the effective gravitational 
$\alpha$ measured
is in very good agreement with the prediction of Gammie (2001). Gammie's 
prediction follows from an argument of balance of GI heating and cooling, resulting in
``gravitoturbulence''. However, his detailed simulations found that the torques and
heating in the disk were due in equal parts to gravitational stresses and Reynolds
stresses, this is in contrast to my measurements which only consider gravitational
stresses.

\acknowledgements

\begin{thebibliography}{}

\bibitem
[Armitage et al. (2001)]{Arm01}
Armitage, P.J., Livio, M., \& Pringle, J.E. 2001, \mnras, 347, 959

\bibitem[Balbus \& Papaloizou (1999)]{Bal99}
Balbus, S. A. \& Papaloizou, J. C. B. 1999, \apj, 521, 650

\bibitem
% PAPER 3
[Boley et al.(2006)]{Bol06}
Boley, A.C., Mej\'{i}a, A.C., Durisen, R.H., Cai, K., Pickett, M.K., 
\& DAlessio, P. 2006, \apj, 651, 517 (Paper III)

\bibitem
[Boley \& Durisen ((2006)]{BoDu06}
Boley, A.C., \& Durisen, R.H. 2006, \apj, 641, 534

\bibitem
[Boley \& Durisen ((2008)]{Bol08}
Boley, A.C., \& Durisen, R.H. 2008, \apj, 685, 1193

\bibitem
[Boley (2009)]{Bol09}
Boley, A.C. 2009, \apjl, 695, L53

\bibitem
[Boley et al. (2010)]{Bol10}
Boley, A.C., Hayfield, T.,  Mayer, L., \& Durisen, R.H. 2010, Icarus, 207, 509

\bibitem
[Boss (1998)]{Bos98}
Boss, A.P. 1998, \apj, 503, 923

%\bibitem
%[Boss (2000)]{Bos00}
%Boss, A.P. 2000a, \apjl, 536, L101
%Boss, A.P. 2000b, \apj, 545, L61

\bibitem
[Boss (2001)]{Bos01}
Boss, A.P. 2001, \apj, 563, 367

%\bibitem
%[Boss (2002)]{Bos02}
%Boss, A.P. 2002a, \apjl, 567, L149
%Boss, A.P. 2002b, \apj, 576, 462

%\bibitem
%[Boss (2003)]{Bos03}

\bibitem
[Boss (2005)]{Bos05}
Boss, A.P. 2005, \apj, 629, 535

\bibitem
[Boss (2007)]{Bos07}
Boss, A.P. 2007, \apjl, 661, L73

\bibitem
% PAPER 4
[Cai et al. (2008)]{Cai08}
Cai, K., Durisen, R. H., Boley, A. C., Pickett, M. K., \& Mej\'{i}a, A. C. 2008, \apj, 673, 1138 (Paper IV)

\bibitem
[Clarke et al. (2007)]{Cla07}
Clarke, C. J., Harper-Clark, E., \& Lodato, G. 2007, \mnras, 381, 1543

\bibitem
[Clarke (2009)]{Cla09}
Clarke, C.J.  2009, \mnras, 396,1066

\bibitem
[Cossins et al. (2009)]{Cos09}
Cossins, P., Lodato, G. \& Clarke, C.  2009, \mnras, 393, 1157

\bibitem
[D'Angelo et al. (2010)]{DAn10}
D'Angelo, G., Durisen, R.H., \& Lissauer, J.J. 2010, in Exoplanets, ed. S. Seager (U. Arizona Press),
in press.

\bibitem
[Durisen (2010)]{Dur10}
Durisen, R.H. 2010, in Physical Processes in Circumstellar Disks Around Young Stars, ed. P. Garcia
(Chicago University Press), in press

\bibitem
[Gammie (1996)]{Gam96}
Gammie, C.F. 1996, \apj, 457, 355

\bibitem
[Gammie (2001)]{Gam01}
Gammie, C. F. 2001, \apj, 553, 174

\bibitem
[Imamura et al. (2000)]{Ima00}
Imamura, J. N., Durisen, R. H., \& Pickett, B. K. 2000, \apj, 528, 946

\bibitem
[Johnson \& Gammie (2003)]{Joh03}
Johnson, B.M., \& Gammie, C.F. 2003, \apj, 597, 131

\bibitem
[Lodato \& Rice (2004)]{Lod04}
Lodato, G., \& Rice, W. K. M. 2004, \mnras, 351,630

\bibitem
[Lodato \& Rice (2005)]{Lod05}
Lodato, G., \& Rice, W.K.M. 2005, \mnras, 358, 1489

\bibitem
[Mayer et al. (2002)]{May02}
Mayer, L., Quinn, T., Wadsley, J., \& Stadel, J. 2002, Science, 298, 1756

\bibitem
[Mayer et al. (2004)]{May04}
Mayer, L., Quinn, T., Wadsley, J., \& Stadel, J. 2004, \apj, 609, 1045

\bibitem
[Mayer et al. (2007)]{May07}
Mayer, L., Lufkin, G., Quinn, T., \& Wadsley, J., 2007, \apjl, 661, L77

\bibitem
[Mej\'{i}a (2004)]{Mej04}
Mej\'{i}a, A. C. 2004, Ph.D. Thesis, Indiana Univ.

\bibitem
% PAPER 2
[Mej\'{i}a et al. (2005)]{Mej05}
Mej\'{i}a, A. C., Durisen, R. H., Pickett, \& M. K., Cai, K. 2005, \apj, 619, 1098 (Paper II)

\bibitem
[Papaloizou \& Lin (1995)]{Pap95}
Papaloizou, J.C.B. \& Lin, D.N.C. 1995, Ann. Rev. \aap, 33, 505

\bibitem
[Pickett (1995)]{Pic95}
Pickett, B. K. 1995, Ph.D. thesis, Indiana Univ.

\bibitem
[Pickett et al. (1998)]{Pic98}
Pickett, B. K., Cassen, P. M., Durisen, R. H., \& Link, R. 1998, ApJ, 504, 468

\bibitem
[Pickett et al. (2000)]{Pic00}
\rule[0.1cm]{1.6cm}{0.01cm}. 2000, ApJ, 529, 1034

\bibitem
%PAPER 1
[Pickett et al. (2003)]{Pic03}
Pickett, B. K., Mej\'{i}a, A. C., Durisen, R. H., Cassen, P. M., Berry, D. K., \&
Link, R. P. 2003, ApJ, 590, 1060 (Paper I)

\bibitem
[Rafikov (2009)]{Raf09}
Rafikov, R.R. 2009, \apj, 704, 281

\bibitem
[Rice et al. (2003)]{Ric03}
Rice, W.K.M., Armitage, P.J., Bate, M.R., \& Bonnell, I.A. 2003, \mnras, 339, 1025

\bibitem
[Rice et al. (2005)]{Ric05}
Rice, W. K. M., Lodato, G., \& Armitage, P. J. 2005, \mnras, 364, L56

\bibitem
[Shakura & Sunyaev (1973)]{Sha73}
Shakura, N. I., \& Sunyaev, R. A. 1973, \aap, 24, 337

\bibitem
[Stamatellos \& Whitworth (2008)]{Sta08}
Stamatellos, D., \& Whitworth, A.P. 2008, \aap, 480, 879

\bibitem
[Vorobyov \& Basu (2005)]{Vor05}
Vorobyov, E.I., \& Basu, S. 2005, \apjl, 633, L137

\bibitem
[Vorobyov \& Basu (2006)]{Vor06}
Vorobyov, E.I., \& Basu, S. 2005, \apj, 650, 956

\bibitem
[Vorobyov (2010)]{Vor10}
Vorobyov, E. I. 2010, New Astron., 15, 24

\bibitem
[Zhu et al. (2009)]{Zhu09}
Zhu, Z., Hartmann, L., Gammie, C.J., \& McKinney, J.C. 2009, \apj, 701, 620

%\hline[3em]  2000, ApJ, 529, 1034

\end{thebibliography}

\newpage

% \begin{deluxetable}{cccc} 
% \tablecolumns{4} 
% \tablewidth{0pc} 
% \tablecaption{Effective Alphas} 
% \tablehead{ 
% \colhead{}    &  \multicolumn{3}{c}{Non-shell Stars} &   \colhead{}   & 
%\multicolumn{3}{c}{Shell Stars} \\ 
%\cline{2-4} \cline{6-8} \\ 
%\cline{2-8} \\ 
% \colhead{Azimuthal Zones} & \colhead{$\alpha$}   & \colhead{$A_{tot}$}    
% & \colhead{$A_{2-5} / A_{tot}$} \\ 
%\colhead{} & \colhead{} &\colhead{} &  \colhead{angle} & \colhead{$10^{-28}$ ergs /}  \\
%\colhead{} & \colhead{} &\colhead{} &  \colhead{angle} 
%&  \multicolumn{2}{c}{$(10^{-28}$ ergs / cm$^3$ s sr)}  \\
% } 
% \startdata
% 64	 &    0.0587	 &        2.02 	 & 0.752	      \\
% 128	   &    0.0402	  &       1.99  	 & 0.608      \\
% 256	  &     0.0237	 &        1.87	 & 0.535     \\
% 512	  &     0.0241	 &        2.28	 & 0.467     \\
% \enddata 
% \tablecomments{
% Table 1. The $\alpha$ of Figure 1, averaged over the radial region from 10 -- 40 AU, tabulated for different 
% azimuthal resolutions, along with amplitudes of Fourier components for non-axisymmetric density structures. 
% The radially averaged $\alpha$ predicted by Gammie (2001) are tablulated for comparison.
% }
% \end{deluxetable}

\begin{deluxetable}{cccccc} 
\tablecolumns{5} 
\tablewidth{0pc} 
\tablecaption{$Q$, Fourier Amplitudes, and $\alpha$} 
\tablehead{ 
   \colhead{$l_{max}$} 
& \colhead{$Q_{avg}$}   
& \colhead{$\langle A_{tot} \rangle$}    
& \colhead{$\langle A_{2-7} \rangle / \langle A_{tot} \rangle$} 
& \colhead{$\alpha_{avg}$} \\ 
} 
\startdata
64   & 1.26 & 1.96 &  0.76 & 0.069 \\
128 & 1.24 & 1.91 &  0.62 & 0.044 \\
256 & 1.39 & 1.83 &  0.54 & 0.026 \\
512 & 1.48 & 1.95 &  0.49 & 0.020 \\
\enddata 
\tablecomments{
All quantities are time-averaged over 12 -- 19 orp.
$Q_{avg}$ and $\alpha_{avg}$ are also spatially
averaged over 10 - 40 AU.
}
\end{deluxetable}
\newpage

\begin{figure}\label{Final_Q}
\plotone{Final_Q_Values.pdf}
\caption{Final Q values - {\bf \it NEEDS REPLOTTING}
}
\end{figure}

\begin{figure}\label{fig:DensityPlots}
\plotone{Figure_1.pdf}
\caption
{
Midplane disk densities at $t \sim 18$ orpS for simulations with 64, 128, 256 and 512 azimuthal grid elements.
The simulations differ only in the number of  azimuthal zones used in the calculations.  Densities out of the plane, along an azimuthal cut through the disk, are displayed below the face-on views.  The color scale is  logarithmic, and axis units are AU. 
}
\end{figure}
\newpage

\begin{figure}\label{fig:Am_vs_log_m}
\plotone{Figure_3.pdf}
\caption
{
The logarithm of the amplitudes for the global Fourier components of the non-axisymmetric density structure plotted for different $m$-values. Notice that the components with the highest global amplitudes are those with low $m$-values.  
}
\end{figure}

\begin{figure}
\plotone{Power_Extrapolation.pdf}
\caption
{
Convergence of fractional power.  The $x$-axis represents
the azimuthal ``spacing'' between grid points,  normalized such that spacing of the $l_{max}$ = 512 grid
equals unity.
The $y$-axis displays the fraction of the total power in the low-order ($m = $ 2--7) Fourier modes. 
Also plotted is the Richardson Extrapolation for the asymptotic value  
$\langle A_{2-7} \rangle / \langle A_{tot} \rangle = 0.43$ as the grid spacing goes to zero.  
}
\end{figure}

\begin{figure}
\plotone{Torque_Array.pdf}
\caption
{
Torque Array}
\end{figure}

%\begin{figure}
%\plotone{Mode_CascadeRatios.pdf}
%\caption
%{
%Mode Cascade Ratios .  Note that by the end of the time frame shown, there is rough parity between 
%torques arising from tower order ($m = 2$ -- 5) mades and torques from higher order modes.}
%\end{figure}

\begin{figure}
\label{fig:alpha_v_radius}
\plotone{Figure_2.pdf}
\caption
{Effective $\alpha$'s for the $t_{cool} =$ 2 orp, $\gamma = 5/3$, $\Sigma \sim r^{ -1}$ disk, averaged between 12 and 18 orps, for four different different azimuthal grid resolutions, given by the number of azimuthal grid-points Lmax. The parallel dashed lines show the predictions of Gammie (2001) for $\alpha$ (upper: strongly self-gravitating limit, lower: non-self-gravitating limit), while the dashed line at $\log(\alpha) = -1.22$ represents the maximum stress a disk can maintain without fragmenting (Rice et al. 2005) . 
}
\end{figure}
\newpage

\begin{figure}
\plotone{Alpha_variations_stack512.pdf}
\caption
{Alpha variations.}
\end{figure}

\end{document}

### 2010-10-29 Paper_1.tex
Made changes to eqns 5 & 6
	modified:   Paper_1.tex
	new file:   figures/.gitignore

@@
 \usepackage{graphicx}
+\usepackage{epstopdf}
 \usepackage{amsmath, amsthm, amssymb}
@@
 \begin{equation}
-b_m = \int \rho \sin(m\phi)\varpi {\rm d}\varpi {\rm d}z {\rm d}\phi.
+b_m =\int \rho \sin(m\phi)\varpi {\rm d}\varpi {\rm d}z {\rm d}\phi.
 \end{equation}
@@
 The torque contribution from each mode $m$ can be calculated 
-by replacing $\rho$ in Equation (5) with the corresponding modal density component
+by replacing $\rho$ in Equation (5) with the density distribution reconstructed from a single Fourier component
 \begin{equation}
-\rho_m = \slantfrac{1}{2} \rho_o + a_m \cos(m\phi) + b_m \sin(m\phi),
+\rho_m = a_{\phi m} \cos(m\phi) + b_{\phi m} \sin(m\phi),
 \end{equation}
-where $a_m$ and $b_m$ are provided by Equation 2,
+where $a_{\phi m} =\slantfrac{1}{\pi} \int \rho \cos(m\phi){\rm d}\phi$ and $b_{\phi m} =\slantfrac{1}{\pi} \int \rho \sin(m\phi){\rm d}\phi$,
 and only the gravitational potential produced by the mass distribution
@@
 \begin{split}
- T_{\varpi\phi}^{grav}  {\varpi}& = \frac{1}{2\pi\varpi^2}C_Z
-\\
+ T_{\varpi\phi}^{grav}\left( \varpi \right) & = \frac{1}{2\pi\varpi^2}C_Z\\
                                 & = - \frac{1}{2\pi\varpi^2} \int \rho\frac{\partial\Phi}{\partial\phi} {\rm d}V.

### 2010-11-01 Paper_1.tex
Fixed references, converted to BibTex
	modified:   Paper_1.tex
	new file:   general.bib

@@
 \shorttitle{Effective Disk Alphas}
-\shortauthors{Scott et al.}
+\shortauthors{Michael et al.}
 
@@
 
-Using 3D hydrodynamic simulations of a cooling protostellar disk undergoing gravitational instabilities (GIs), we compute the effective Shakura and Sunyaev (1973) $\alpha$ due to gravitational stresses and compare them to predictions from an analytic local theory for thin disks by Gammie (2001).
+Using 3D hydrodynamic simulations of a cooling protostellar disk undergoing gravitational instabilities (GIs), we compute the effective \citet{shakura1973} $\alpha$ due to gravitational stresses and compare them to predictions from an analytic local theory for thin disks by \citet{gammie2001}.
 Our goal is to determine how accurately a locally defined alpha can characterize mass and angular momentum transport by GIs in disks.
-We examine disks with an imposed constant global cooling time (Mej\'{i}a et al.
-2005).
+We examine disks with an imposed constant global cooling time \citep{mejia2005}.
 Models were computed with grids of different spatial resolution to investigate how the computed alpha is affected by numerical resolution.
@@
 
-While detailed three-dimensional hydrodynamic modeling of protostellar disks can
-provide considerable insight into the physical and thermal states of these disks,
-computing resources limit such studies to snapshots
-spanning timescales much shorter than a disk's evolutionary lifetime.  In contrast, 
-approaches that replace detailed physics with statistical parameters that
-encapsulate this physics allow for calculations covering
-much longer timeframes.  Both approaches can play important roles in broadening our 
-understanding of these complex nonlinear systems, provided a connection between the detailed physics 
-and statistical parameters can be demonstrated and understood.   In what follows, we
-examine the applicability of an $\alpha$-disk formulation (Shakura \& Shaviv 1973) to  the
+While detailed three-dimensional hydrodynamic modeling of protostellar disks can provide considerable insight into the physical and thermal states of these disks, computing resources limit such studies to snapshots spanning timescales much shorter than a disk's evolutionary lifetime.
+In contrast, approaches that replace detailed physics with statistical parameters that encapsulate this physics allow for calculations covering much longer timeframes.
+Both approaches can play important roles in broadening our understanding of these complex nonlinear systems, provided a connection between the detailed physics and statistical parameters can be demonstrated and understood.
+In what follows, we examine the applicability of an $\alpha$-disk formulation \citep{shakura1973} to  the
 evolution of protostellar disks subject to gravitational instabilities.
 
-Gravitational instabilities (GIs) play an important, and at times dominant,
-role in driving the structural and thermal
-evolution of  protostellar disks \citep[for a review, see][]{Dur10}.  
+Gravitational instabilities (GIs) play an important, and at times dominant, role in driving the structural and thermal
+evolution of  protostellar disks \citep[for a review, see][]{durisen2005}.
 %In particular, torques arising from GIs provide a very effective means of redistributing matter and 
@@
 %\citep{Lod04, Lod05, Vor05, Vor06, Bol09}.  
-Thermal processes play the primary role in regulating the
-amplitude and outcome of these instabilities  (Pickett
-et al. 1998, 2000a,b, 2003, see also Nelson et al. 1998, 2000). 
-A link between a disk's thermodynamic state and its susceptibility to GIs
-is provided by the Toomre $Q$-parameter (Toomre 1981);  $Q = c_s \kappa / \pi G \Sigma$,
-where $c_s$ is the sound speed, $\kappa$ is the epicyclic frequency ($\sim$ the rotation frequency 
+Thermal processes play the primary role in regulating the amplitude and outcome of these instabilities \citep{pickett1998,pickett2000,mejia2005,nelson1998,nelson2000a}.
+A link between a disk's thermodynamic state and its susceptibility to GIs is provided by the Toomre $Q$-parameter \citep{toomre1981};  $Q = c_s \kappa / \pi G \Sigma$, where $c_s$ is the sound speed, $\kappa$ is the epicyclic frequency ($\sim$ the rotation frequency
 $\Omega$ in a nearly Keplerian disk), and $\Sigma$ is the disk surface mass density.
-For $Q \al$  1.5 to 1.7,  small density perturbations in a disk
-grow exponentially on a time scale comparable to the rotation period.
-These perturbations manifest themselves as multi-arm spirals with a predominantly trailing pattern that
-transports angular momentum outward by
-gravitational torques (e.g.,Larson 1984, Boss 1984, Durisen et al. 1986, Papaloizou \& Savonije
-1991, Laughlin \& Bodenheimer 1994, Laughlin et al., 1998; Nelson
-et al., 1998; Pickett et al., 1998, add refs?).
+For $Q \al$  1.5 to 1.7,  small density perturbations in a disk grow exponentially on a time scale comparable to the rotation period.
+These perturbations manifest themselves as multi-arm spirals with a predominantly trailing pattern that transports angular momentum outward by gravitational torques \citep[e.g.]{larson1984,boss1984,durisen1986,papaloizou1991, laughlin1994,nelson1998,pickett1998}.
 The long range nature of these torques cause nonlinear
-spiral structures to develop over a broad range of radii (Laughlin et al. 1998, Nelson et al. 1998,
-2000; Pickett et al. 1998, 2000, 2003).
+spiral structures to develop over a broad range of radii \citep{laughlin1998,nelson1998,nelson2000a,pickett1998, pickett2000,pickett2003}.
 
-Numerous studies
-have shown that thermal physics can control GI amplitudes
-by a {\it balance} of heating and cooling (e.g., Tomley et al. 1991, 1994;
-Pickett et al., 1998, 2000,
-2003; Gammie, 2001; Boss, 2002; Rice et al., 2003; Mej\'{i}a et al., 2005; Boley
-et al., 2006, 2007a; Stamatellos and Whitworth, 2008; Cossins et al., 2009).
-Heating by GIs tends to increase $c_s$, thus increasing $Q$. If disk heating persists,
-the instability will shut off once $Q$ becomes sufficiently large.  However, if the
-disk cools radiatively, then heating and cooling can reach a balance at nonlinear
-wave amplitude. GI-activity can sustain this balance  at a relatively constant,
-but unstable, value of $Q$ (e.g., Paczynski, 1978; Lin and Pringle; Pringle, 1981;
-see also Goldreich and  Lynden-Bell 1965).  
-For sufficiently low cooling rates, the disk can exist in a state of quasi-equilibrium, 
-where cooling is 
+Numerous studies have shown that thermal physics can control GI amplitudes by a {\it balance} of heating and cooling \citep[for example]{tomley1991,tomley1994,pickett1998,pickett2000, pickett2003,gammie2001,boss2002,rice2003b,mejia2005,boley2006,boley2007, stamatellos2008,cossins2009}.
+Heating by GIs tends to increase $c_s$, thus increasing $Q$.
+If disk heating persists, the instability will shut off once $Q$ becomes sufficiently large.
+However, if the disk cools radiatively, then heating and cooling can reach a balance at nonlinear wave amplitude.
+GI-activity can sustain this balance  at a relatively constant,
+but unstable, value of $Q$ \citep{paczynski1978, lin1981,goldreich1965}.  
+For sufficiently low cooling rates, the disk can exist in a state of quasi-equilibrium, where cooling is 
 balanced by heating arising from the inward flow of matter and shocks induced by the GIs
-\citep{Gam01, Lod04, Ric05, Cos09, Vor10}.
+\citep{gammie2001, lodato2004, rice2005, cossins2009, vorobyov2010}.
 
 Fully nonlinear hydrodynamic simulations of low-$Q$ disks show that multiple
-modes can become unstable in the linear regime (e.g., Nelson et al.,
-1998; Pickett et al., 1998; additional refs?).  Even if the initial growth is dominated by
-one mode, numerous modes usually appear in the nonlinear regime. As a consequence, 
-if a disk achieves a quasi-steady balance between heating and cooling, 
-then there tends to be power at
-all $m$-values resolved by a given numerical method (e.g., Mej\'{i}õa et al., 2005;
-Boley et al., 2006).  
-In this  {\it asymptotic state}, GI-active disks develop overlapping
-density structures of different strengths, geometries and coherence,
-producing
-strongly nonlinear turbulence that pervades the entire disk
-(e.g., Pickett et al. 2003, Mej\'{i}a et al. 2005, Boley et al. 2006).
-Gammie (2001) refers to this state as {\it gravitoturbulence}.  
-Gravitoturbulence provides a possible bridge between
-the detailed physics of GIs and the viscous transport parameter $\alpha$.
+modes can become unstable in the linear regime \citep[e.g.]{nelson1998, pickett1998} {\bf additional refs?}.
+Even if the initial growth is dominated by one mode, numerous modes usually appear in the nonlinear regime.
+As a consequence, if a disk achieves a quasi-steady balance between heating and cooling, then there tends to be power at all $m$-values resolved by a given numerical method \citep{mejia2005,boley2006}.
+In this  {\it asymptotic state}, GI-active disks develop overlapping density structures of different strengths, geometries and coherence,producing strongly nonlinear turbulence that pervades the entire disk \citep[e.g.]{pickett2003, mejia2005,boley2006}.
+\citet{gammie2001} refers to this state as {\it gravitoturbulence}. Gravitoturbulence provides a possible bridge between the detailed physics of GIs and the viscous transport parameter $\alpha$.
 
-Shakura \& Syunyaev (1973) first parameterized turbulent transport in a steady-state
-disk with the  parameter $\alpha$, where the source of the
-turbulence is unspecified.
+\citet{shakura1973} first parameterized turbulent transport in a steady-state disk with the  parameter $\alpha$, where the source of the turbulence is unspecified.
 Although their original work was directed towards X-ray bright
-accretion disks around black holes, this heuristic
-approach has proved fruitful in many 
-astrophysical situations by allowing for
-analytic solutions and relatively easy numerical modeling
-over time-spans inaccessible to detailed modeling.
-In the $\alpha$ disk formalism, when the accretion rate
-$\dot{M}$ is constant in both time and radius, $\dot{M} = 
-3\pi\alpha c_s^2 \Sigma / \Omega$ where $c_s$ is the sound speed, $\Sigma$ the 
-surface density, and $\Omega$ the rotational frequency. 
-The $\alpha$ 
-formalism makes no assumption
-about the nature of the angular momentum transport, only that {\it heating and cooling
-are dominated by local processes}; the validity
-of an $\alpha$-disk picture requires self-consistency of local energy dissipation.
+accretion disks around black holes, this heuristic approach has proved fruitful in many astrophysical situations by allowing for analytic solutions and relatively easy numerical modeling over time-spans inaccessible to detailed modeling.
+In the $\alpha$ disk formalism, when the accretion rate $\dot{M}$ is constant in both time and radius, $\dot{M} = 3\pi\alpha c_s^2 \Sigma / \Omega$ where $c_s$ is the sound speed, $\Sigma$ the surface density, and $\Omega$ the rotational frequency.
+The $\alpha$ formalism makes no assumption about the nature of the angular momentum transport, only that {\it heating and cooling are dominated by local processes}; the validity of an $\alpha$-disk picture requires self-consistency of local energy dissipation.
 
-Dissipative locality in GI-active disks has not been rigorously tested and 
-remains an open question. 
-Several studies support the idea that GI
-transport in real disks is, in many important respects, an intrinsically global phenomenon and
-cannot be properly treated by a local $\alpha$-like prescription (Laughlin \& Rozyczka 1996, Balbus
-\& Papaloizou 1999, Lodato \& Rice 2005, Mej\'{i}a et al. 2005, Boley et al. 2006, Cai et al.
-2007).  This almost certainly applies to the embedded phase of both high and low-mass
-stars when infall from the collapsing protostellar cloud feeds mass into the disk at a rapid
-rate (e.g., Yorke et al. 1993, Laughlin \& Bodenhiemer 1994, Yorke \& Bodenheimer 1999,
-Vorobyov \& Basu 2005, 2006, Kratter \& Matzner 2007), and disks observed in the embedded
-phase can be quite massive relative to their central stars (e.g., Osorio et al. 2003). 
-However, despite the long-distance nature of gravitational torques, theoretical
-and numerical results suggest that in protostellar disks where heating and cooling are in
-balance it may be valid to represent GI-induced transport by an {\it effective viscosity}, a local process
-(e.g., Gammie 2001; Rice et al. 2005; Lodato \& Rice 2004).
-Indeed, Gammie (2001) derived an effective alpha for the case of a thin (effectively 2D)
-gravitationally unstable disk where the heating due to local turbulent dissipation caused by
-GIs was balanced by the local cooling rate.  Modifications to this derivation have since been
-proposed by {\bf (Tom: include other derivations and refs)}.   Recent simulations (e.g., Cossins
-et al. 2009; Vorobyov 2010) suggest that, at least in the case of imposed cooling,
-$\alpha_{eff}$ roughly converges to the value predicted by Gammie.
+Dissipative locality in GI-active disks has not been rigorously tested and remains an open question. 
+Several studies support the idea that GI transport in real disks is, in many important respects, an intrinsically global phenomenon and cannot be properly treated by a local $\alpha$-like prescription \citep{laughlin1996, balbus1999, lodato2005, mejia2005, boley2006, cai2008}.
+This almost certainly applies to the embedded phase of both high and low-mass stars when infall from the collapsing protostellar cloud feeds mass into the disk at a rapid
+rate \citep[e.g.]{yorke1993, laughlin1994, yorke1999,vorobyov2005,vorobyov2006}, and disks observed in the embedded phase can be quite massive relative to their central stars \citep[e.g.]{osorio2003}.
+However, despite the long-distance nature of gravitational torques, theoretical and numerical results suggest that in protostellar disks where heating and cooling are in balance it may be valid to represent GI-induced transport by an {\it effective viscosity}, a local process \citep[e.g.]{gammie2001, rice2005, lodato2004}.
+Indeed, \citet{gammie2001} derived an effective alpha for the case of a thin 2D gravitationally unstable disk where the heating due to local turbulent dissipation caused by GIs was balanced by the local cooling rate.
+Modifications to this derivation have since been proposed by {\bf (Tom: include other derivations and refs)}.
+Recent simulations \citep[e.g.]{cossins2009, vorobyov2010} suggest that, at least in the case of imposed cooling, $\alpha_{eff}$ roughly converges to the value predicted by \citeauthor{gammie2001}.
 
@@
 {\bf (Tom: add text concerning need for convergence approach)}
-Here we report the results of a convergence study designed to determine
-whether GI-active disks, chaotic by nature, achieve statistical equilibrium; if so, to what
-level do these disks approximate the gravitoturbulence of Gammie (2001) and with what distribution
-of $\alpha_{eff}$; and to characterize how the GI cascades from low-order to high-order modes. 
-To this end, we follow the evolution of a GI-active disk, subject to constant cooling, in its asymptotic state,
-where cooling and heating are essentially balanced.  Simulations are
-performed using four different resolutions in the azimuthal direction.
+Here we report the results of a convergence study designed to determine whether GI-active disks, chaotic by nature, achieve statistical equilibrium; if so, to what level do these disks approximate the gravitoturbulence of \citet{gammie2001} and with what distribution of $\alpha_{eff}$; and to characterize how the GI cascades from low-order to high-order modes.
+To this end, we follow the evolution of a GI-active disk, subject to constant cooling, in its asymptotic state, where cooling and heating are essentially balanced.
+Simulations are performed using four different resolutions in the azimuthal direction.
 
@@
 
-Simulations were conducted with the three-dimensional hydrodynamics 
-code used by the
-Indiana University Hydrodynamics Group in several previous studies
-\citep{Pic98, Pic00, Mej04, Mej05, Bol06, Cai08}.  This code
-uses a second-order, explicit, Eulerian scheme to solve Poisson's equation, 
-an equation of state,
-and the equations of hydrodynamics in conservative
-form on a uniform cylindrical grid.  The code assumes mirror symmetry about the equatorial plane.
-It includes both
-self-gravity and artificial viscosity; the latter serves to 
-mediate shocks.
-Source and flux terms (Norman \&Winkler 1986) are computed separately in an
-explicit, second-order time integration (van Albada et al. 1982; Christodoulou 1991; Yang 1992).
+Simulations were conducted with the three-dimensional hydrodynamics code used by the Indiana University Hydrodynamics Group in several previous studies \citep{pickett1998, pickett2000, mejiaphd2004, mejia2005, boley2006, cai2008}.
+This code uses a second-order, explicit, Eulerian scheme to solve Poisson's equation, an equation of state, and the equations of hydrodynamics in conservative form on a uniform cylindrical grid.
+The code assumes mirror symmetry about the equatorial plane.
+It includes both self-gravity and artificial viscosity; the latter serves to
+mediate shocks. Source and flux terms \citep{norman1986} are computed separately in an explicit, second-order time integration \citep{albada1982,christodoulou1991,yangphd1992}.
 
-We assume an equation of state $P = (\gamma -1)\epsilon$, 
-where $P$ is the pressure, $\epsilon$ is the internal energy density, and $\gamma$ 
-is the ratio of specific heats, given here by $\gamma = 5/3$.  
-The model disks are cooled by decreasing their internal energy according to 
-the prescription $d\epsilon/dt = \epsilon / t_{cool}$, where $t_{cool}$ is the global cooling time in Outer Rotation Periods (orp).
-For the grid resolutions used here, one orp is defined as the rotational period at radial zone 200.
+We assume an equation of state $P = (\gamma -1)\epsilon$, where $P$ is the pressure, $\epsilon$ is the internal energy density, and $\gamma$ is the ratio of specific heats, given here by $\gamma = 5/3$.
+The model disks are cooled by decreasing their internal energy according to the prescription $d\epsilon/dt = \epsilon / t_{cool}$, where $t_{cool}$ is the global cooling time in Outer Rotation Periods (ORP).
+For the grid resolutions used here, one ORP is defined as the rotational period at radial zone 200.
 
@@
 In what follows, quantities specified with real physical units are derived assuming a 1 $M_\odot$ central star and a 0.153  $M_\odot$ disk.
-In this case, the orp is defined at $r = 33$ AU and
-one orp $\approx$ 303 years.  Models have $t_{cool} = 2$ orp.   {\bf (Scott: Please check these numbers)}
-The disk has
-an initial surface density $\Sigma \sim r^{ -1}$, and initial inner and outer radii of 2.3 and 40 AU, respectively.
+In this case, the ORP is defined at $r = 33$ AU and one ORP $\approx$ 303 years.
+Models have $t_{cool} = 2$ ORP.
+{\bf (Scott: Please check these numbers)}
+The disk has an initial surface density $\Sigma \sim r^{ -1}$, and initial inner and outer radii of 2.3 and 40 AU, respectively.
 
-Initial conditions for the disk's structure and thermodynamic state
-were set using an equilibrium star plus disk model generated using a modified Haschisu (1986)
-self-consistent field (SCF) relaxation method (Pickett et al. 1996, 2003;
-Mej\'{i}a (2004);  Mej\'{i}a et al. (2005); Cai 2006).  
-Here  the density and angular
-momentum distributions are iteratively
-solved, using the specified equation of state,  until convergence is achieved. 
-This  equilibrium disk was then seeded with small random perturbations to
-allow GIs to grow.  The resultant marginally unstable disk served as the $t = 0$ load for the hydro code.
+Initial conditions for the disk's structure and thermodynamic state were set using an equilibrium star plus disk model generated using a modified \citet{hachisu1986} self-consistent field (SCF) relaxation method \citep{pickett1996,pickett2003,mejiaphd2004,mejia2005,cai2006}.
+Here  the density and angular momentum distributions are iteratively solved, using the specified equation of state,  until convergence is achieved.
+This  equilibrium disk was then seeded with small random perturbations to allow GIs to grow.
+The resultant marginally unstable disk served as the $t = 0$ load for the hydro code.
 
-The initially unstable disk model passed through several phases of evolution
-(see also Pickett et al., 2003; Mej\'{i}a et al., 2005).  The initial load is unstable to
-the growth of axisymmetric structures. During this axisymmetric phase, 
-which lasts several orps,
-the disk shrinks slightly, in an axisymmetric manner, until it becomes unstable
-to nonaxisymmetric perturbations.    When this condition is met, the disk undergoes a strong burst 
-of GIs that predominantly manifest themselves in
-one or two discrete global spiral modes.  The disk expands violently during the burst phase,
-producing a significant rearrangement of the disk's mass distribution on a timescale of one or
-two orps. 
-This is then followed by a period of several orps where heating temporarily washes
-out some of the nonaxisymmetry in the disk.  Finally the disk settles into
-a quasi-steady, long-lived asymptotic state of sustained GI-activity
-over a large part of the disk, with an overall balance of heating and cooling.
+The initially unstable disk model passed through several phases of evolution \citep[see also]{pickett2003, mejia2005}.
+The initial load is unstable to the growth of axisymmetric structures.
+During this axisymmetric phase,
+which lasts several ORPs, the disk shrinks slightly, in an axisymmetric manner, until it becomes unstable to nonaxisymmetric perturbations.
+When this condition is met, the disk undergoes a strong burst of GIs that predominantly manifest themselves in one or two discrete global spiral modes.
+The disk expands violently during the burst phase, producing a significant rearrangement of the disk's mass distribution on a timescale of one or
+two ORPs.
+This is then followed by a period of several ORPs where heating temporarily washes out some of the nonaxisymmetry in the disk.
+Finally the disk settles into a quasi-steady, long-lived asymptotic state of sustained GI-activity over a large part of the disk, with an overall balance of heating and cooling.
 
-Simulations were run with four different angular resolutions, with grid sizes in $(r,\phi,z)$
-of  $(j_{max}, k_{max}, l_{max}) =$
- (512,64,64), (512,128,64), (512,256,64), and (512,512,64).
-Because of the large computational overhead associated with starting each simulation from
-$t = 0$,
-the $l_{max} =$ 128 simulation was run through the axisymmetric, burst and transition
-phases. Simulations using $l_{max} =$ 64, 256, and 512 
-all begin by interpolating the $l_{max} =$ 128 simulation
-to higher or lower azimuthal resolution at 9.6 ORPs, near the end of the transition
-phase. Although this strategy saves a large amount of computational time, it limits
-the scope of the analysis to the asymptotic phase and
-it introduces a perturbation to the disk which takes several
-orps to subside. By 12 orps, the perturbation has dissipated and all four disks
-have transitioned to the asymptotic state.  All simulations were 
-run through $\sim 18$ orps.
+Simulations were run with four different angular resolutions, with grid sizes in $(r,\phi,z)$ of  $(j_{max}, k_{max}, l_{max}) =$ (512,64,64), (512,128,64), (512,256,64), and (512,512,64).
+Because of the large computational overhead associated with starting each simulation from$t = 0$, the $l_{max} =$ 128 simulation was run through the axisymmetric, burst and transition phases.
+Simulations using $l_{max} =$ 64, 256, and 512 all begin by interpolating the $l_{max} =$ 128 simulation to higher or lower azimuthal resolution at 9.6 ORPs, near the end of the transition
+phase.
+Although this strategy saves a large amount of computational time, it limits the scope of the analysis to the asymptotic phase and it introduces a perturbation to the disk which takes several ORPs to subside.
+By 12 ORPs, the perturbation has dissipated and all four disks have transitioned to the asymptotic state.
+All simulations were
+run through $\sim 18$ ORPs.
 
-Figure 1 shows $Q$, as a function of radius, at $t \sim$ 18 orp, while
-Table 1 lists $Q_{avg}$, the value of $Q$ of for the asymptotic disk, time-averaged of
-12 -- 19 orp and spatially averaged over 10 -- 40 AU, as a function of radius
-for each resolution, 
+Figure \ref{Final_Q} shows $Q$, as a function of radius, at $t \sim$ 18 ORP, while table \ref{tbl:ams} lists $Q_{avg}$, the value of $Q$ of for the asymptotic disk, time-averaged of 12 -- 19 ORP and spatially averaged over 10 -- 40 AU, as a function of radius for each resolution,
 {\bf (Scott: Figure 1 needs to be redone - I have extracted the 
@@
 
-In the asymptotic state, GI-active disks develop complex
-density structures arising from the superposition of
-all modes present in the disk.  Since the number of
-modes accessible
-to the disk is a function of the azimuthal resolution,
-we expect that
-higher resolution simulations will manifest more power
-in higher-order modes 
-that behave more locally. We test this by examining the power in various 
-modes to see how they are affected by the choice of grid.
+In the asymptotic state, GI-active disks develop complex density structures arising from the superposition of
+all modes present in the disk.
+Since the number of modes accessible to the disk is a function of the azimuthal resolution, we expect that higher resolution simulations will manifest more power in higher-order modes
+that behave more locally. We test this by examining the power in various modes to see how they are affected by the choice of grid.
 
@@
 
-Figure 2
-%\ref{fig:DensityPlots}, 
-shows midplane and radial densities  at $t \sim 18$ orps  
-for the $l_{max} =$ 64, 128, 256, and 512 simulations. 
-The lower resolution (smaller $l_{max}$) simulations
-display longer wavelength spiral structures, with larger amplitudes and more coherence than those
-seen in higher resolution 
-simulations.  
-Higher angular resolution
-allows GI power to spread to higher-order azimuthal modes, as demonstrated 
-by a Fourier decomposition of the density.
+Figure \ref{fig:DensityPlots}, shows midplane and radial densities  at $t \sim 18$ ORPs for the $l_{max} =$ 64, 128, 256, and 512 simulations.
+The lower resolution (smaller $l_{max}$) simulations display longer wavelength spiral structures, with larger amplitudes and more coherence than those seen in higher resolution simulations.
+Higher angular resolution allows GI power to spread to higher-order azimuthal modes, as demonstrated by a Fourier decomposition of the density.
 The strengths of these modes are given by their global Fourier amplitudes,
+
 \begin{equation}
@@
 \renewcommand{\theequation}{\arabic{equation}}
+
 Here  $\rho_o$ is the axisymmetric (zero-order)  component of the density.
-In what follows, $A_m$ is averaged over time to 
-suppress fluctuations on the dynamic time scale due
-to the chaotic nature of GIs in the asymptotic  phase.
-In particular, $\langle A_{tot} \rangle$ will represent 
-the power over the radial range 10 -- 40 AU,
-integrated from $ m = 2$ to $l_{max}/2$ and
-averaged over the time range 12 -- 19 orps.  
-The density distribution cannot be resolved in azimuth
-on angular scales less than $2\pi/l_{max}$, therefore $A_m$ is
-limited to modes $m \le l_{max}/2$.  
-Because the central star is artificially fixed to the grid center, the $m = 1$ mode 
-cannot be be realistically treated; it is excluded from
- $\langle A_{tot} \rangle$.
+In what follows, $A_m$ is averaged over time to suppress fluctuations on the dynamic time scale due to the chaotic nature of GIs in the asymptotic  phase.
+In particular, $\langle A_{tot} \rangle$ will represent the power over the radial range 10 -- 40 AU, integrated from $ m = 2$ to $l_{max}/2$ and averaged over the time range 12 -- 19 ORPs.
+The density distribution cannot be resolved in azimuth on angular scales less than $2\pi/l_{max}$, therefore $A_m$ is limited to modes $m \le l_{max}/2$.
+Because the central star is artificially fixed to the grid center, the $m = 1$ mode cannot be be realistically treated; it is excluded from $\langle A_{tot} \rangle$.
 
@@
 
-Previous studies have shown that disks in the asymptotic state have power at
-all resolvable $m$-values  (e.g., Mej\'{i}õa et al., 2005;
-Boley et al., 2006).  Table 1 lists $\langle A_{tot} \rangle$ along with 
-$\langle A_{2-7} \rangle  /  \langle A_{tot} \rangle$,  the fraction of the total
-power falling into low-order ($m = 2$ -- 7) modes, for each resolution. 
-While the total power is approximately independent of
-resolution, $\langle A_{tot} \rangle = 1.90 \pm 0.07$, 
+Previous studies have shown that disks in the asymptotic state have power at all resolvable $m$-values  \citep[e.g.]{mejia2005,boley2006}.
+Table \ref{tbl:ams} lists $\langle A_{tot} \rangle$ along with $\langle A_{2-7} \rangle  /  \langle A_{tot} \rangle$,  the fraction of the total power falling into low-order ($m = 2$ -- 7) modes, for each resolution.
+While the total power is approximately independent of resolution, $\langle A_{tot} \rangle = 1.90 \pm 0.07$,
 the modal distribution of this power shows marked differences.
-In particular, the fractional power in lower order modes decreases monotonically
-with increasing resolution.  Over three quarters of $\langle A_{tot} \rangle$ resides
-in modes $m =$ 2 -- 7 for the $l_{max} = 64$ simulation, while approximately
-half resides in these modes when $l_{max} =$ 512.
+In particular, the fractional power in lower order modes decreases monotonically with increasing resolution.
+Over three quarters of $\langle A_{tot} \rangle$ resides
+in modes $m =$ 2 -- 7 for the $l_{max} = 64$ simulation, while approximately half resides in these modes when $l_{max} =$ 512.
 
-The shift of power from lower to higher order modes as the resolution of
-the computational grid is increased arises from the fact that, while
-$\langle A_{tot} \rangle$ is approximately independent of $l_{max}$, 
-the number of degrees of freedom available for this power, $l_{max}/2$, 
-increases linearly with the resolution.  
-Thus power is naturally spread from lower order to higher order modes
-as the available $m$-values increases.  
-Indeed, even if the disk is dominated by one or two low-order modes at the
-time of outburst, this power cascades to higher-order modes (e.g., 
-Laughlin et al. 1997, 1998; see also Laughlin and Korchagin 1996).
+The shift of power from lower to higher order modes as the resolution of the computational grid is increased arises from the fact that, while $\langle A_{tot} \rangle$ is approximately independent of $l_{max}$, the number of degrees of freedom available for this power, $l_{max}/2$, increases linearly with the resolution.
+Thus power is naturally spread from lower order to higher order modes as the available $m$-values increases.
+Indeed, even if the disk is dominated by one or two low-order modes at the time of outburst, this power cascades to higher-order modes \citep[for example]{laughlin1997, laughlin1998,laughlin1996}.
 
-Given the resolution dependence of how power is
-distributed,  we must examine 
-what resolution is sufficient to properly capture the evolution of these disks.
-Table 1 shows that while the fractional power in lower order modes
-decreases with each doubling of $l_{max}$, the amount of change in going from $l_{max} = 256$ to 512 
-is significantly less than the change going from $l_{max} = 64$ to 128,
-with a strong suggestion
-that this diminution will not proceed much further with higher azimuthal resolution.  
-Indeed, a  Richardson extrapolation {\bf (add reference)} 
-of the tabulated values of  $\langle A_{2-7} \rangle / \langle A_{tot} \rangle$
-suggests a limiting value of  $\langle A_{2-7} \rangle / \langle A_{tot} \rangle \sim$ 0.43  
-for increasingly higher resolutions.
-This is displayed graphically in Figure 4, where the
-fractional power in lower order modes is plotted against the  normalized grid 
-spacing for each value of $l_{max}$.
-The grid spacing  is normalized such that spacing of the $l_{max}$ = 512 grid
-equals unity.  Convergence occurs as the normalized spacing goes to zero.
-The Richardson Extrapolation value  of $\langle A_{2-7}\rangle  / \langle A_{tot}\rangle  = 0.43$ is
-plotted at a normalized grid spacing  equal to zero.   More than half of the modal power in
-a converged disk resides in higher-order modes, modes 
-that help define a state of gravitoturbulence.
-The fact that
-the $l_{max} =$ 512 simulation has
-$\langle A_{2-7} \rangle / \langle A_{tot} \rangle$ near the convergence value
-lends some confidence in the use of the $l_{max} =$ 512 model in
-examining the convergence of $\alpha$.
+Given the resolution dependence of how power is distributed,  we must examine what resolution is sufficient to properly capture the evolution of these disks.
+Table \ref{tbl:ams} shows that while the fractional power in lower order modes decreases with each doubling of $l_{max}$, the amount of change in going from $l_{max} = 256$ to 512
+is significantly less than the change going from $l_{max} = 64$ to 128, with a strong suggestion that this diminution will not proceed much further with higher azimuthal resolution.
+Indeed, a  Richardson extrapolation {\bf (add reference)} of the tabulated values of  $\langle A_{2-7} \rangle / \langle A_{tot} \rangle$ suggests a limiting value of  $\langle A_{2-7} \rangle / \langle A_{tot} \rangle \sim$ 0.43 for increasingly higher resolutions.
+This is displayed graphically in Figure 4, where the fractional power in lower order modes is plotted against the  normalized grid
+spacing for each value of $l_{max}$. The grid spacing  is normalized such that spacing of the $l_{max}$ = 512 grid
+equals unity.
+Convergence occurs as the normalized spacing goes to zero.
+The Richardson extrapolation value  of $\langle A_{2-7}\rangle  / \langle A_{tot}\rangle  = 0.43$ is plotted at a normalized grid spacing  equal to zero.
+More than half of the modal power in a converged disk resides in higher-order modes, modes that help define a state of gravitoturbulence.
+The fact that the $l_{max} =$ 512 simulation has $\langle A_{2-7} \rangle / \langle A_{tot} \rangle$ near the convergence value lends some confidence in the use of the $l_{max} =$ 512 model in examining the convergence of $\alpha$.
 
@@
 Coherent density structures in a disk produce gravitational torques.
-Low-order modes in $\Delta\rho / \rho$  have longer wavelengths and, if coherent in
-radius, have longer lever arms to produce torques. On the other hand, high-order
-modes have relatively short wavelengths which, particularly if they lack coherence,
-produce more localized effects. The high-order
-modes tend to produce pockets of local GI activity, which can cancel each other
-over the whole disk, leading to the expectation that  gravitational torques in the disk 
-will decrease as $l_{max}$ is increased.    This can be tested by examining the
-torque contributions from  various Fourier modes.
+Low-order modes in $\Delta\rho / \rho$  have longer wavelengths and, if coherent in radius, have longer lever arms to produce torques.
+On the other hand, high-order modes have relatively short wavelengths which, particularly if they lack coherence, produce more localized effects.
+The high-order modes tend to produce pockets of local GI activity, which can cancel each other
+over the whole disk, leading to the expectation that  gravitational torques in the disk will decrease as $l_{max}$ is increased.
+This can be tested by examining the torque contributions from  various Fourier modes.
 
-The torque {\bf C} on a cylindrical surface of the disk at radius $\varpi$
-can be obtained by integrating the stress tensor $T$ over the
-surface of the cylinder (e.g., Lynden-Bell \& Kalnajs 1972), i.e.,
+The torque {\bf C} on a cylindrical surface of the disk at radius $\varpi$ can be obtained by integrating the stress tensor $T$ over the surface of the cylinder \citep{lyndenbell1972}, i.e.,
 \begin{equation}
@@
 \end{equation}
-If the stress tensor includes only gravitational stresses,
-the surface integral can be replaced with the volume integral
+If the stress tensor includes only gravitational stresses, the surface integral can be replaced with the volume integral
 \begin{equation}
@@
 \end{equation}
-where $\Phi$ is the gravitational potential.  Here we are interested
-only in the $z$-component of torque,  
+where $\Phi$ is the gravitational potential.  Here we are interested only in the $z$-component of torque,  
 \begin{equation}
@@
 \end{equation}
-as only this component  
-drives mass and angular momentum transfer.  
-The torque contribution from each mode $m$ can be calculated 
-by replacing $\rho$ in Equation (5) with the density distribution reconstructed from a single Fourier component
+as only this component drives mass and angular momentum transfer.
+The torque contribution from each mode $m$ can be calculated by replacing $\rho$ in Equation (5) with the density distribution reconstructed from a single Fourier component
 \begin{equation}
@@
 \end{equation}
-where $a_{\phi m} =\slantfrac{1}{\pi} \int \rho \cos(m\phi){\rm d}\phi$ and $b_{\phi m} =\slantfrac{1}{\pi} \int \rho \sin(m\phi){\rm d}\phi$,
-and only the gravitational potential produced by the mass distribution
-given by $\rho_m$ is included in $\Phi$.
+where $a_{\phi m} =\slantfrac{1}{\pi} \int \rho \cos(m\phi){\rm d}\phi$ and $b_{\phi m} =\slantfrac{1}{\pi} \int \rho \sin(m\phi){\rm d}\phi$, and only the gravitational potential produced by the mass distribution given by $\rho_m$ is included in $\Phi$.
 {\bf (Scott: Equation (6) is the form you use in your chapter 3, but it
@@
 
-Figure 5 displays $\langle {\rm \bf C}_{Ztot} \rangle$, the time-averaged gravitational 
-torque summed over all disk modes
-(excluding $m=1$),
-along with the torques summed over low-order modes,
-$\langle {\rm \bf C}_{Z(2-n)} \rangle$, where $n = $ 2, 3, 4, and 5, as a function of radius
-and azimuthal resolution. 
+Figure 5 displays $\langle {\rm \bf C}_{Ztot} \rangle$, the time-averaged gravitational torque summed over all disk modes (excluding $m=1$), along with the torques summed over low-order modes, $\langle {\rm \bf C}_{Z(2-n)} \rangle$, where $n = $ 2, 3, 4, and 5, as a function of radius and azimuthal resolution.
 {\bf (Note to all: the version of Figure 5 included with this draft  shows the torques
 time-averaged over for four different time-frames.  I did this because I wanted to see if the disk was 
-"settling down" with time.  I think the figure suggests  that the 12-14 orp disk is not
+"settling down" with time.  I think the figure suggests  that the 12-14 ORP disk is not
 yet in an asymptotic state - or further from it than at later times - I'm not sure this is
 a significant difference.  Looking forward to Figure 7, it can be argued that the
-disk is reasonably settled, in a statistical sense, shortly after 12 orp.
-Dick has told me he only believes the 12-18 orp result - Dick, I hope 
+disk is reasonably settled, in a statistical sense, shortly after 12 ORP.
+Dick has told me he only believes the 12-18 ORP result - Dick, I hope 
 I'm not misquoting you).  
 } 
-The scale of the $y$-axis in each subplot 
-is the same, with the exception of the $l_{max} = 64$ simulation. 
-In this case the scale is doubled.  Two features are immediately apparent:
-1) the total gravitational torque is smaller for higher resolution simulations,
-and 2) the total torque is dominated by low-order modes.
-Both reflect the fact that a while higher fraction of the total power in
-larger $l_{max}$ simulations is in higher order modes, these modes make only
-minimal contributions to the gravitational torque.
-As resolution increases,
-non-axisymmetric power is shifted from low-order global modes, 
-which produce relatively large gravitational torques,
-to high-order local modes, which tend to cancel each other out when integrated over the
-whole disk.
+The scale of the $y$-axis in each subplot is the same, with the exception of the $l_{max} = 64$ simulation. 
+In this case the scale is doubled.
+Two features are immediately apparent: 1) the total gravitational torque is smaller for higher resolution simulations, and 2) the total torque is dominated by low-order modes.
+Both reflect the fact that a while higher fraction of the total power in larger $l_{max}$ simulations is in higher order modes, these modes make only minimal contributions to the gravitational torque.
+As resolution increases, non-axisymmetric power is shifted from low-order global modes, which produce relatively large gravitational torques, to high-order local modes, which tend to cancel each other out when integrated over the whole disk.
 
@@
 
-An $\alpha$-framework treats disks as essentially two-dimensional objects
-subject to local energy dissipation.  Therefore, in order
-to compare our results with predicted 
-and published results, we must extract two-dimensional information from our 
-three-dimensional disks.  Following Lodato \& Rice (2004), 
+An $\alpha$-framework treats disks as essentially two-dimensional objects subject to local energy dissipation.
+Therefore, in order to compare our results with predicted and published results, we must extract two-dimensional information from our three-dimensional disks.
+Following \citet{lodato2004},
 \begin{equation}
@@
 \end{equation}
-where $\Omega$ is the azimuthally averaged rotation speed, 
-$T_{\varpi\phi}^{grav}$ and $T_{\varpi\phi}^{Reyn}$ represent the 
-azimuthally averaged
-$\varpi,\phi$-component of the specific (per unit mass)
-gravitational 
-(or ``Newton'') and hydrodynamic (or ``Reynolds'') stress tensors, respectively,
-$\Sigma$ the surface density,
-and $c_s$ the azimuthally averaged midplane sound speed.  Brackets
-surrounding the stress terms indicate that these are integrated over
-the $z$-direction.
+where $\Omega$ is the azimuthally averaged rotation speed, $T_{\varpi\phi}^{grav}$ and $T_{\varpi\phi}^{Reyn}$ represent the azimuthally averaged $\varpi,\phi$-component of the specific (per unit mass) gravitational
+(or ``Newton'') and hydrodynamic (or ``Reynolds'') stress tensors, respectively, $\Sigma$ the surface density,
+and $c_s$ the azimuthally averaged midplane sound speed.
+Brackets surrounding the stress terms indicate that these are integrated over the $z$-direction.
 The denominator $\Sigma c_c^2$ was evaluated as $\int_z \rho c_c^2 {\rm d}z$.
@@
 Numerical uncertainties render the Reynolds stress difficult to determine.
-Fortunately, this difficulty is mitigated by the fact
-that previous studies have shown the
-magnitude of the Reynolds stress to be small relative to the gravitational stress
-(references?).
+Fortunately, this difficulty is mitigated by the fact that previous studies have shown the magnitude of the Reynolds stress to be small relative to the gravitational stress (references?).
 Thus it can be safely neglected in Equation (7).
@@
 {\bf (There is a problem with this equation)}
-Figure 6 shows the effective $\alpha$s for the four different grid resolutions, time
-averaged from 12 to 19 orps.  
-For comparison, the predicted values by Gammie (2001)
-for the strongly self-gravitating limit (lower curve) and the non-self-gravitating limit
-(upper curve) are shown. Note that predictions from are based upon a thin
-disk approximation, where the local balance of heating and cooling dominates the
-energetics.
+Figure 6 shows the effective $\alpha$s for the four different grid resolutions, time averaged from 12 to 19 ORPs.
+For comparison, the predicted values by \citet{gammie2001} for the strongly self-gravitating limit (lower curve) and the non-self-gravitating limit (upper curve) are shown.
+Note that predictions from are based upon a thin
+disk approximation, where the local balance of heating and cooling dominates the energetics.
 
-,
-
-In an $\alpha$ disk formalism the mechanism giving rise to the transport of angular
-momentum also acts to dissipate energy. Lodato \& Rice (2005) postulated that a
-disk in a state of gravitoturbulence would have locally balanced heating and cooling
-rates. If the heating were due to local dissipation, then $\alpha$ and $t_{cool}$ are related by
+In an $\alpha$ disk formalism the mechanism giving rise to the transport of angular momentum also acts to dissipate energy.
+Lodato \& Rice (2005) postulated that a disk in a state of gravitoturbulence would have locally balanced heating and cooling rates.
+If the heating were due to local dissipation, then $\alpha$ and $t_{cool}$ are related by
 \begin{equation}
@@
 \end{equation}
-Rice et al. (2005) concluded that for a given $t_{cool}$  there exists some maximum stress
-quantified by $\alpha_{max}$, which if exceeded would cause a disk to fragment. The horizontal
-dot dashed line in figure 6 indicates the $\alpha_{max} \approx 0.06$ found by Rice et al. (2005).
-Clarke et al. (2007) have revised this limit to $\alpha_{max} = 0.12$ for cases where $t_{cool}$  varies
-with time.
+\citet{rice2005} concluded that for a given $t_{cool}$  there exists some maximum stress quantified by $\alpha_{max}$, which if exceeded would cause a disk to fragment.
+The horizontal dot dashed line in figure 6 indicates the $\alpha_{max} \approx 0.06$ found by \citeauthor{rice2005}.
+\citet{clarke2007} have revised this limit to $\alpha_{max} = 0.12$ for cases where $t_{cool}$  varies with time.
 
-Figure 7 shows the time-variation of $\alpha$, normalized to its value at $t = 18$ orp), as a function of 
-radius and  time.  This plot suggests that the disk has not yet achieved statistical equilibrium until 
-$\sim $14 orp {\bf (Everyone:  What do you think?  We are using averages over12-18 orp,
+Figure 7 shows the time-variation of $\alpha$, normalized to its value at $t = 18$ ORP), as a function of radius and  time.
+This plot suggests that the disk has not yet achieved statistical equilibrium until $\sim $14 ORP {\bf (Everyone:  What do you think?
+We are using averages over12-18 ORP,
 are we comfortable with those limits?)}
@@
 of the paper (excluding a summary).  Rather than further delay getting
-the draft to you, I am sending it as is.   I will be adding text and incorporating results from a 
+the draft to you, I am sending it as is.   I will be adding text and incORPorating results from a 
 few recent papers and will send the rest soon.}
@@
 
-As the azimuthal resolution increases, the effective $\alpha$ in the GI active region
-decreases. This is due to the decrease in the total gravitational torque, as illustrated
-in Figure 5. This decrease in gravitational torque is due, in turn, to the shift of
-non-axisymmetric power from low-order global modes, which produce large torques,
-to high-order local modes, which can cancel each other out when integrated over the
-whole disk. In fact, the decrease in the ratio of low-order power to the total power
-$\langle A_{2-7}\rangle / \langle A_\Sigma \rangle$ 
-closely tracks the decrease in the time averaged $\alpha$ averaged from 10 to
-40 AU, which can be seen in Table 1 as the $\alpha_{avg}$ value. Although the 
-average $\alpha$ values
-have not converged to a single value for my highest two resolutions, the percentage
-change continues to decrease. In addition, Figure 6 shows that the $l_{max} =$ 256
-and 512 simulations have qualitatively similar time averaged profiles over a large
-range of disk radii. These profiles also fall within the predicted Gammie curves, with
-tendency toward the lower (strongly self-gravitating) curve. Indeed the $\alpha_{avg}$
-value
-of 0.020 for the $l_{max} = $ 512 simulation is squarely between the $\alpha_{avg}$ values for the
-Gammie curves 0.014 (strongly self-gravitating) and 0.027 (non-self-gravitating).
+As the azimuthal resolution increases, the effective $\alpha$ in the GI active region decreases.
+This is due to the decrease in the total gravitational torque, as illustrated in Figure 5.
+This decrease in gravitational torque is due, in turn, to the shift of non-axisymmetric power from low-order global modes, which produce large torques, to high-order local modes, which can cancel each other out when integrated over the whole disk.
+In fact, the decrease in the ratio of low-order power to the total power $\langle A_{2-7}\rangle / \langle A_\Sigma \rangle$ closely tracks the decrease in the time averaged $\alpha$ averaged from 10 to 40 AU, which can be seen in table \ref{tbl:ams} as the $\alpha_{avg}$ value.
+Although the average $\alpha$ values have not converged to a single value for my highest two resolutions, the percentage change continues to decrease.
+In addition, Figure 6 shows that the $l_{max} =$ 256
+and 512 simulations have qualitatively similar time averaged profiles over a large range of disk radii.
+These profiles also fall within the predicted Gammie curves, with tendency toward the lower (strongly self-gravitating) curve.
+Indeed the $\alpha_{avg}$ value of 0.020 for the $l_{max} = $ 512 simulation is squarely between the $\alpha_{avg}$ values for the Gammie curves 0.014 (strongly self-gravitating) and 0.027 (non-self-gravitating).
 
-Although the effective $\alpha$ curves do correspond to the Gammie prediction, the
-results from these simulations differ from Gammie's in one significant way. The local
-shearing box simulations performed by Gammie (2001) show that the total 
-effective $\alpha$ is composed of a gravitational stress component and a Reynolds stress component
-that are nearly equal (see Gammie (2001) Figure 3). However, all of the analyses
+Although the effective $\alpha$ curves do correspond to the Gammie prediction, the results from these simulations differ from Gammie's in one significant way.
+The local shearing box simulations performed by \citet{gammie2001} show that the total effective $\alpha$ is composed of a gravitational stress component and a Reynolds stress component that are nearly equal (see Gammie (2001) Figure 3).
+However, all of the analyses
 presented here only consider the gravitational stress. When I attempt to measure
@@
 
-\begin{thebibliography}{}
-
-\bibitem
-[Armitage et al. (2001)]{Arm01}
-Armitage, P.J., Livio, M., \& Pringle, J.E. 2001, \mnras, 347, 959
-
-\bibitem[Balbus \& Papaloizou (1999)]{Bal99}
-Balbus, S. A. \& Papaloizou, J. C. B. 1999, \apj, 521, 650
-
-\bibitem
-% PAPER 3
-[Boley et al.(2006)]{Bol06}
-Boley, A.C., Mej\'{i}a, A.C., Durisen, R.H., Cai, K., Pickett, M.K., 
-\& DAlessio, P. 2006, \apj, 651, 517 (Paper III)
-
-\bibitem
-[Boley \& Durisen ((2006)]{BoDu06}
-Boley, A.C., \& Durisen, R.H. 2006, \apj, 641, 534
-
-\bibitem
-[Boley \& Durisen ((2008)]{Bol08}
-Boley, A.C., \& Durisen, R.H. 2008, \apj, 685, 1193
-
-\bibitem
-[Boley (2009)]{Bol09}
-Boley, A.C. 2009, \apjl, 695, L53
-
-\bibitem
-[Boley et al. (2010)]{Bol10}
-Boley, A.C., Hayfield, T.,  Mayer, L., \& Durisen, R.H. 2010, Icarus, 207, 509
-
-\bibitem
-[Boss (1998)]{Bos98}
-Boss, A.P. 1998, \apj, 503, 923
-
-%\bibitem
-%[Boss (2000)]{Bos00}
-%Boss, A.P. 2000a, \apjl, 536, L101
-%Boss, A.P. 2000b, \apj, 545, L61
-
-\bibitem
-[Boss (2001)]{Bos01}
-Boss, A.P. 2001, \apj, 563, 367
-
-%\bibitem
-%[Boss (2002)]{Bos02}
-%Boss, A.P. 2002a, \apjl, 567, L149
-%Boss, A.P. 2002b, \apj, 576, 462
-
-%\bibitem
-%[Boss (2003)]{Bos03}
-
-\bibitem
-[Boss (2005)]{Bos05}
-Boss, A.P. 2005, \apj, 629, 535
-
-\bibitem
-[Boss (2007)]{Bos07}
-Boss, A.P. 2007, \apjl, 661, L73
-
-\bibitem
-% PAPER 4
-[Cai et al. (2008)]{Cai08}
-Cai, K., Durisen, R. H., Boley, A. C., Pickett, M. K., \& Mej\'{i}a, A. C. 2008, \apj, 673, 1138 (Paper IV)
-
-\bibitem
-[Clarke et al. (2007)]{Cla07}
-Clarke, C. J., Harper-Clark, E., \& Lodato, G. 2007, \mnras, 381, 1543
-
-\bibitem
-[Clarke (2009)]{Cla09}
-Clarke, C.J.  2009, \mnras, 396,1066
-
-\bibitem
-[Cossins et al. (2009)]{Cos09}
-Cossins, P., Lodato, G. \& Clarke, C.  2009, \mnras, 393, 1157
-
-\bibitem
-[D'Angelo et al. (2010)]{DAn10}
-D'Angelo, G., Durisen, R.H., \& Lissauer, J.J. 2010, in Exoplanets, ed. S. Seager (U. Arizona Press),
-in press.
-
-\bibitem
-[Durisen (2010)]{Dur10}
-Durisen, R.H. 2010, in Physical Processes in Circumstellar Disks Around Young Stars, ed. P. Garcia
-(Chicago University Press), in press
-
-\bibitem
-[Gammie (1996)]{Gam96}
-Gammie, C.F. 1996, \apj, 457, 355
-
-\bibitem
-[Gammie (2001)]{Gam01}
-Gammie, C. F. 2001, \apj, 553, 174
-
-\bibitem
-[Imamura et al. (2000)]{Ima00}
-Imamura, J. N., Durisen, R. H., \& Pickett, B. K. 2000, \apj, 528, 946
-
-\bibitem
-[Johnson \& Gammie (2003)]{Joh03}
-Johnson, B.M., \& Gammie, C.F. 2003, \apj, 597, 131
-
-\bibitem
-[Lodato \& Rice (2004)]{Lod04}
-Lodato, G., \& Rice, W. K. M. 2004, \mnras, 351,630
-
-\bibitem
-[Lodato \& Rice (2005)]{Lod05}
-Lodato, G., \& Rice, W.K.M. 2005, \mnras, 358, 1489
-
-\bibitem
-[Mayer et al. (2002)]{May02}
-Mayer, L., Quinn, T., Wadsley, J., \& Stadel, J. 2002, Science, 298, 1756
-
-\bibitem
-[Mayer et al. (2004)]{May04}
-Mayer, L., Quinn, T., Wadsley, J., \& Stadel, J. 2004, \apj, 609, 1045
-
-\bibitem
-[Mayer et al. (2007)]{May07}
-Mayer, L., Lufkin, G., Quinn, T., \& Wadsley, J., 2007, \apjl, 661, L77
-
-\bibitem
-[Mej\'{i}a (2004)]{Mej04}
-Mej\'{i}a, A. C. 2004, Ph.D. Thesis, Indiana Univ.
-
-\bibitem
-% PAPER 2
-[Mej\'{i}a et al. (2005)]{Mej05}
-Mej\'{i}a, A. C., Durisen, R. H., Pickett, \& M. K., Cai, K. 2005, \apj, 619, 1098 (Paper II)
-
-\bibitem
-[Papaloizou \& Lin (1995)]{Pap95}
-Papaloizou, J.C.B. \& Lin, D.N.C. 1995, Ann. Rev. \aap, 33, 505
-
-\bibitem
-[Pickett (1995)]{Pic95}
-Pickett, B. K. 1995, Ph.D. thesis, Indiana Univ.
-
-\bibitem
-[Pickett et al. (1998)]{Pic98}
-Pickett, B. K., Cassen, P. M., Durisen, R. H., \& Link, R. 1998, ApJ, 504, 468
-
-\bibitem
-[Pickett et al. (2000)]{Pic00}
-\rule[0.1cm]{1.6cm}{0.01cm}. 2000, ApJ, 529, 1034
-
-\bibitem
-%PAPER 1
-[Pickett et al. (2003)]{Pic03}
-Pickett, B. K., Mej\'{i}a, A. C., Durisen, R. H., Cassen, P. M., Berry, D. K., \&
-Link, R. P. 2003, ApJ, 590, 1060 (Paper I)
-
-\bibitem
-[Rafikov (2009)]{Raf09}
-Rafikov, R.R. 2009, \apj, 704, 281
-
-\bibitem
-[Rice et al. (2003)]{Ric03}
-Rice, W.K.M., Armitage, P.J., Bate, M.R., \& Bonnell, I.A. 2003, \mnras, 339, 1025
-
-\bibitem
-[Rice et al. (2005)]{Ric05}
-Rice, W. K. M., Lodato, G., \& Armitage, P. J. 2005, \mnras, 364, L56
-
-\bibitem
-[Shakura & Sunyaev (1973)]{Sha73}
-Shakura, N. I., \& Sunyaev, R. A. 1973, \aap, 24, 337
-
-\bibitem
-[Stamatellos \& Whitworth (2008)]{Sta08}
-Stamatellos, D., \& Whitworth, A.P. 2008, \aap, 480, 879
-
-\bibitem
-[Vorobyov \& Basu (2005)]{Vor05}
-Vorobyov, E.I., \& Basu, S. 2005, \apjl, 633, L137
-
-\bibitem
-[Vorobyov \& Basu (2006)]{Vor06}
-Vorobyov, E.I., \& Basu, S. 2005, \apj, 650, 956
-
-\bibitem
-[Vorobyov (2010)]{Vor10}
-Vorobyov, E. I. 2010, New Astron., 15, 24
-
-\bibitem
-[Zhu et al. (2009)]{Zhu09}
-Zhu, Z., Hartmann, L., Gammie, C.J., \& McKinney, J.C. 2009, \apj, 701, 620
-
-%\hline[3em]  2000, ApJ, 529, 1034
-
-\end{thebibliography}
+\bibliographystyle{apj}
+\bibliography{general}
 
@@
 
-\begin{deluxetable}{cccccc} 
+\begin{deluxetable}{cccccc}\label{tbl:ams}
 \tablecolumns{5} 
@@
 \tablecomments{
-All quantities are time-averaged over 12 -- 19 orp.
+All quantities are time-averaged over 12 -- 19 ORP.
 $Q_{avg}$ and $\alpha_{avg}$ are also spatially
@@
 \begin{figure}\label{Final_Q}
-\plotone{Final_Q_Values.pdf}
+%\plotone{Final_Q_Values.eps}
 \caption{Final Q values - {\bf \it NEEDS REPLOTTING}
@@
 \begin{figure}\label{fig:DensityPlots}
-\plotone{Figure_1.pdf}
+%\plotone{Figure_1.eps}
 \caption
 {
-Midplane disk densities at $t \sim 18$ orpS for simulations with 64, 128, 256 and 512 azimuthal grid elements.
+Midplane disk densities at $t \sim 18$ ORPs for simulations with 64, 128, 256 and 512 azimuthal grid elements.
 The simulations differ only in the number of  azimuthal zones used in the calculations.
@@
 \begin{figure}\label{fig:Am_vs_log_m}
-\plotone{Figure_3.pdf}
+%\plotone{Figure_3.eps}
 \caption
@@
 \begin{figure}
-\plotone{Power_Extrapolation.pdf}
+%\plotone{Power_Extrapolation.eps}
 \caption
@@
 \begin{figure}
-\plotone{Torque_Array.pdf}
+%\plotone{Torque_Array.eps}
 \caption
@@
 \label{fig:alpha_v_radius}
-\plotone{Figure_2.pdf}
+%\plotone{Figure_2.eps}
 \caption
-{Effective $\alpha$'s for the $t_{cool} =$ 2 orp, $\gamma = 5/3$, $\Sigma \sim r^{ -1}$ disk, averaged between 12 and 18 orps, for four different different azimuthal grid resolutions, given by the number of azimuthal grid-points Lmax.
+{Effective $\alpha$'s for the $t_{cool} =$ 2 ORP, $\gamma = 5/3$, $\Sigma \sim r^{ -1}$ disk, averaged between 12 and 18 ORPs, for four different different azimuthal grid resolutions, given by the number of azimuthal grid-points Lmax.
 The parallel dashed lines show the predictions of Gammie (2001) for $\alpha$ (upper: strongly self-gravitating limit, lower: non-self-gravitating limit), while the dashed line at $\log(\alpha) = -1.22$ represents the maximum stress a disk can maintain without fragmenting (Rice et al.
@@
 \begin{figure}
-\plotone{Alpha_variations_stack512.pdf}
+%\plotone{Alpha_variations_stack512.eps}
 \caption

### 2010-11-04 Paper_1.tex
test commit
Edits to Paper. 1st pass.
	modified:   Paper_1.tex
	new file:   figures/qcompareres.eps
	modified:   general.bib

@@
 
-\newcounter{subequation}
-
+\usepackage{natbib}
 \usepackage{color}
@@
 \usepackage{epstopdf}
+\usepackage{grffile}
 \usepackage{amsmath, amsthm, amssymb}
+\usepackage{amssymb}
 
@@
 Using 3D hydrodynamic simulations of a cooling protostellar disk undergoing gravitational instabilities (GIs), we compute the effective \citet{shakura1973} $\alpha$ due to gravitational stresses and compare them to predictions from an analytic local theory for thin disks by \citet{gammie2001}.
-Our goal is to determine how accurately a locally defined alpha can characterize mass and angular momentum transport by GIs in disks.
+Our goal is to determine how accurately a locally defined $\alpha$ can characterize mass and angular momentum transport by GIs in disks.
 We examine disks with an imposed constant global cooling time \citep{mejia2005}.
-Models were computed with grids of different spatial resolution to investigate how the computed alpha is affected by numerical resolution.
+Models were computed with grids of different spatial resolution to investigate how the computed $\alpha$ is affected by numerical resolution.
 
@@
 
-Gravitational instabilities (GIs) play an important, and at times dominant, role in driving the structural and thermal
+Gravitational instabilities (GIs) play an important, {\bf SM: I'm not sure this is true in general.
+Perhaps reword to ``GIs can play an important role...and may dominate the structure for certain phases of evolution''}and at times dominant, role in driving the structural and thermal
 evolution of  protostellar disks \citep[for a review, see][]{durisen2005}.
@@
 Thermal processes play the primary role in regulating the amplitude and outcome of these instabilities \citep{pickett1998,pickett2000,mejia2005,nelson1998,nelson2000a}.
-A link between a disk's thermodynamic state and its susceptibility to GIs is provided by the Toomre $Q$-parameter \citep{toomre1981};  $Q = c_s \kappa / \pi G \Sigma$, where $c_s$ is the sound speed, $\kappa$ is the epicyclic frequency ($\sim$ the rotation frequency
+A a disk's susceptibility to GIs is parameterized by the Toomre $Q$-parameter \citep{toomre1981};  $Q = c_s \kappa / \pi G \Sigma$, where $c_s$ is the sound speed, $\kappa$ is the epicyclic frequency ($\sim$ the rotation frequency
 $\Omega$ in a nearly Keplerian disk), and $\Sigma$ is the disk surface mass density.
@@
 spiral structures to develop over a broad range of radii \citep{laughlin1998,nelson1998,nelson2000a,pickett1998, pickett2000,pickett2003}.
+{\bf SM: not sure of the meaning here.
+A low Q is what causes GIs to develop over a broad range of radii.
+Perhaps you mean the long range nature can cause coherent spiral structures to form with large radial extent?}
 
-Numerous studies have shown that thermal physics can control GI amplitudes by a {\it balance} of heating and cooling \citep[for example]{tomley1991,tomley1994,pickett1998,pickett2000, pickett2003,gammie2001,boss2002,rice2003b,mejia2005,boley2006,boley2007, stamatellos2008,cossins2009}.
+Numerous studies have shown that thermal physics control GI amplitudes by a {\it balance} of heating and cooling \citep[for example][]{tomley1991,tomley1994,pickett1998,pickett2000, pickett2003,gammie2001,boss2002,rice2003b,mejia2005,boley2006,boley2007, stamatellos2008,cossins2009}.
 Heating by GIs tends to increase $c_s$, thus increasing $Q$.
-If disk heating persists, the instability will shut off once $Q$ becomes sufficiently large.
-However, if the disk cools radiatively, then heating and cooling can reach a balance at nonlinear wave amplitude.
+If disk heating persists, the instability will be suppressed once $Q$ becomes sufficiently large.
+However, heating and cooling can reach a balance at nonlinear wave amplitude.
 GI-activity can sustain this balance  at a relatively constant,
-but unstable, value of $Q$ \citep{paczynski1978, lin1981,goldreich1965}.  
-For sufficiently low cooling rates, the disk can exist in a state of quasi-equilibrium, where cooling is 
+but unstable, value of $Q$ \citep{paczynski1978, lin1981,goldreich1965},and 
+the disk can exist in a state of quasi-equilibrium. In this state cooling is 
 balanced by heating arising from the inward flow of matter and shocks induced by the GIs
@@
 Fully nonlinear hydrodynamic simulations of low-$Q$ disks show that multiple
-modes can become unstable in the linear regime \citep[e.g.]{nelson1998, pickett1998} {\bf additional refs?}.
+modes can become unstable in the linear regime \citep[e.g.][]{nelson1998, pickett1998} {\bf additional refs?}.
 Even if the initial growth is dominated by one mode, numerous modes usually appear in the nonlinear regime.
 As a consequence, if a disk achieves a quasi-steady balance between heating and cooling, then there tends to be power at all $m$-values resolved by a given numerical method \citep{mejia2005,boley2006}.
-In this  {\it asymptotic state}, GI-active disks develop overlapping density structures of different strengths, geometries and coherence,producing strongly nonlinear turbulence that pervades the entire disk \citep[e.g.]{pickett2003, mejia2005,boley2006}.
+In this  {\it asymptotic state}, GI-active disks develop overlapping density structures of different strengths, geometries and coherence, producing strongly nonlinear turbulence that pervades the entire disk \citep[e.g.][]{pickett2003, mejia2005,boley2006}.
 \citet{gammie2001} refers to this state as {\it gravitoturbulence}. Gravitoturbulence provides a possible bridge between the detailed physics of GIs and the viscous transport parameter $\alpha$.
@@
 accretion disks around black holes, this heuristic approach has proved fruitful in many astrophysical situations by allowing for analytic solutions and relatively easy numerical modeling over time-spans inaccessible to detailed modeling.
-In the $\alpha$ disk formalism, when the accretion rate $\dot{M}$ is constant in both time and radius, $\dot{M} = 3\pi\alpha c_s^2 \Sigma / \Omega$ where $c_s$ is the sound speed, $\Sigma$ the surface density, and $\Omega$ the rotational frequency.
-The $\alpha$ formalism makes no assumption about the nature of the angular momentum transport, only that {\it heating and cooling are dominated by local processes}; the validity of an $\alpha$-disk picture requires self-consistency of local energy dissipation.
+In the $\alpha$ disk formalism, when the accretion rate $\dot{M}$ is constant in both time and radius, $\dot{M} = 3\pi\alpha c_s \Sigma H$ where $c_s$ is the sound speed, $\Sigma$ the surface density, and $H$ the disk scale height.
+The $\alpha$ formalism makes no assumption about the nature of the angular momentum transport, only that {\it heating and cooling are dominated by local processes}{\bf SM: This is not my understanding of the $\alpha$ formalism, I don't think it says anything about cooling.
+It also means that ang mom transport happens locally and that this causes local heating}; the validity of an $\alpha$-disk picture requires self-consistency of local energy dissipation.
 
-Dissipative locality in GI-active disks has not been rigorously tested and remains an open question. 
+Dissipative locality {\bf SM: need to explain what is meant by this more precisely} in GI-active disks has not been rigorously tested and remains an open question. 
 Several studies support the idea that GI transport in real disks is, in many important respects, an intrinsically global phenomenon and cannot be properly treated by a local $\alpha$-like prescription \citep{laughlin1996, balbus1999, lodato2005, mejia2005, boley2006, cai2008}.
 This almost certainly applies to the embedded phase of both high and low-mass stars when infall from the collapsing protostellar cloud feeds mass into the disk at a rapid
-rate \citep[e.g.]{yorke1993, laughlin1994, yorke1999,vorobyov2005,vorobyov2006}, and disks observed in the embedded phase can be quite massive relative to their central stars \citep[e.g.]{osorio2003}.
-However, despite the long-distance nature of gravitational torques, theoretical and numerical results suggest that in protostellar disks where heating and cooling are in balance it may be valid to represent GI-induced transport by an {\it effective viscosity}, a local process \citep[e.g.]{gammie2001, rice2005, lodato2004}.
+rate \citep[e.g.]{yorke1993, laughlin1994, yorke1999,vorobyov2005,vorobyov2006}, and disks observed in the embedded phase can be quite massive relative to their central stars \citep[e.g.][]{osorio2003}.
+However, despite the long-distance nature of gravitational torques, theoretical and numerical results suggest that in protostellar disks where heating and cooling are in balance it may be valid to represent GI-induced transport by an {\it effective viscosity}, a local process \citep[e.g.][]{gammie2001, rice2005, lodato2004}.
 Indeed, \citet{gammie2001} derived an effective alpha for the case of a thin 2D gravitationally unstable disk where the heating due to local turbulent dissipation caused by GIs was balanced by the local cooling rate.
 Modifications to this derivation have since been proposed by {\bf (Tom: include other derivations and refs)}.
-Recent simulations \citep[e.g.]{cossins2009, vorobyov2010} suggest that, at least in the case of imposed cooling, $\alpha_{eff}$ roughly converges to the value predicted by \citeauthor{gammie2001}.
+Recent simulations \citep[e.g.]{cossins2009, vorobyov2010} suggest that, at least in the case of imposed cooling, $\alpha_{\rm{eff}}$ roughly converges to the value predicted by \citeauthor{gammie2001}.
 
@@
 Models have $t_{cool} = 2$ ORP.
-{\bf (Scott: Please check these numbers)}
 The disk has an initial surface density $\Sigma \sim r^{ -1}$, and initial inner and outer radii of 2.3 and 40 AU, respectively.
@@
 Here  the density and angular momentum distributions are iteratively solved, using the specified equation of state,  until convergence is achieved.
-This  equilibrium disk was then seeded with small random perturbations to allow GIs to grow.
-The resultant marginally unstable disk served as the $t = 0$ load for the hydro code.
+This  equilibrium disk was then seeded with small 0.01\% $\delta\rho/\rho$ random perturbations to allow GIs to grow.
+The resultant marginally unstable disk served as the initial disk for the simulation.
 
-The initially unstable disk model passed through several phases of evolution \citep[see also]{pickett2003, mejia2005}.
-The initial load is unstable to the growth of axisymmetric structures.
+The initially unstable disk model passed through several phases of evolution \citep[see also][]{pickett2003, mejia2005}.
+The initial disk is unstable to the growth of axisymmetric structures.
 During this axisymmetric phase,
-which lasts several ORPs, the disk shrinks slightly, in an axisymmetric manner, until it becomes unstable to nonaxisymmetric perturbations.
+which lasts several ORPs, the disk cools and contracts, the contraction is small radially but dramatic in the vertical direction.
+Around 3-4 ORPs the instabilities begin grow to non-linear amplitudes.
 When this condition is met, the disk undergoes a strong burst of GIs that predominantly manifest themselves in one or two discrete global spiral modes.
-The disk expands violently during the burst phase, producing a significant rearrangement of the disk's mass distribution on a timescale of one or
-two ORPs.
+The disk expands violently during the burst phase, producing a significant rearrangement of the disk's mass distribution on a timescale of a few ORPs.
 This is then followed by a period of several ORPs where heating temporarily washes out some of the nonaxisymmetry in the disk.
@@
 
-Simulations were run with four different angular resolutions, with grid sizes in $(r,\phi,z)$ of  $(j_{max}, k_{max}, l_{max}) =$ (512,64,64), (512,128,64), (512,256,64), and (512,512,64).
-Because of the large computational overhead associated with starting each simulation from$t = 0$, the $l_{max} =$ 128 simulation was run through the axisymmetric, burst and transition phases.
+Simulations were run with four different angular resolutions, with grid sizes in $(r,\phi,z)$ of  $(j_{max}, k_{max}, l_{max}) =$ (512,64,64), (512,64,128), (512,64,256), and (512,64,512).
+Because of the large computational overhead associated with starting each simulation from $t = 0$, the $l_{max} =$ 128 simulation was run through the axisymmetric, burst and transition phases.
 Simulations using $l_{max} =$ 64, 256, and 512 all begin by interpolating the $l_{max} =$ 128 simulation to higher or lower azimuthal resolution at 9.6 ORPs, near the end of the transition
@@
 Figure \ref{Final_Q} shows $Q$, as a function of radius, at $t \sim$ 18 ORP, while table \ref{tbl:ams} lists $Q_{avg}$, the value of $Q$ of for the asymptotic disk, time-averaged of 12 -- 19 ORP and spatially averaged over 10 -- 40 AU, as a function of radius for each resolution,
-{\bf (Scott: Figure 1 needs to be redone - I have extracted the 
-plot from your chapter 3 and it only shows the 128 model.  We should show all four resolutions
-or ditch this figure)}.
 
@@
 where
-\renewcommand{\theequation}{\arabic{equation}\alph{subequation}}
-\setcounter{subequation}{2}
-\begin{equation}
-a_m = \int \rho \cos(m\phi)\varpi {\rm d}\varpi {\rm d}z {\rm d}\phi,
-\end{equation}
-\addtocounter{equation}{-1}
-\addtocounter{subequation}{1}
-\begin{equation}
-b_m =\int \rho \sin(m\phi)\varpi {\rm d}\varpi {\rm d}z {\rm d}\phi.
-\end{equation}
-\renewcommand{\theequation}{\arabic{equation}}
+\begin{subequations}
+\begin{align}
+a_m &= \int \rho \cos(m\phi)\varpi {\rm d}\varpi {\rm d}z {\rm d}\phi,\\
+b_m &=\int \rho \sin(m\phi)\varpi {\rm d}\varpi {\rm d}z {\rm d}\phi.
+\end{align}
+\end{subequations}
 
@@
 In particular, $\langle A_{tot} \rangle$ will represent the power over the radial range 10 -- 40 AU, integrated from $ m = 2$ to $l_{max}/2$ and averaged over the time range 12 -- 19 ORPs.
-The density distribution cannot be resolved in azimuth on angular scales less than $2\pi/l_{max}$, therefore $A_m$ is limited to modes $m \le l_{max}/2$.
+The Fourier amplitudes cannot be accurately measured at resolutions smaller than $l_{max}/2$ \citep{shannon1984}, therefore $A_m$ is limited to modes $m \le l_{max}/2$.
 Because the central star is artificially fixed to the grid center, the $m = 1$ mode cannot be be realistically treated; it is excluded from $\langle A_{tot} \rangle$.
@@
 
-Previous studies have shown that disks in the asymptotic state have power at all resolvable $m$-values  \citep[e.g.]{mejia2005,boley2006}.
+Previous studies have shown that disks in the asymptotic state have power at all resolvable $m$-values  \citep[e.g.][]{mejia2005,boley2006}.
 Table \ref{tbl:ams} lists $\langle A_{tot} \rangle$ along with $\langle A_{2-7} \rangle  /  \langle A_{tot} \rangle$,  the fraction of the total power falling into low-order ($m = 2$ -- 7) modes, for each resolution.
 While the total power is approximately independent of resolution, $\langle A_{tot} \rangle = 1.90 \pm 0.07$,
-the modal distribution of this power shows marked differences.
+the distribution of this power shows marked differences.
 In particular, the fractional power in lower order modes decreases monotonically with increasing resolution.
@@
 
-The shift of power from lower to higher order modes as the resolution of the computational grid is increased arises from the fact that, while $\langle A_{tot} \rangle$ is approximately independent of $l_{max}$, the number of degrees of freedom available for this power, $l_{max}/2$, increases linearly with the resolution.
+The shift of power from lower to higher order modes as the resolution of the computational grid is increased arises from the fact that, while $\langle A_{tot} \rangle$ is approximately independent of $l_{max}$, {\bf SM: Need some explanation as to why this is the case.} the number of degrees of freedom available for this power, $l_{max}/2$, increases linearly with the resolution.
 Thus power is naturally spread from lower order to higher order modes as the available $m$-values increases.
-Indeed, even if the disk is dominated by one or two low-order modes at the time of outburst, this power cascades to higher-order modes \citep[for example]{laughlin1997, laughlin1998,laughlin1996}.
+Indeed, even if the disk is dominated by one or two low-order modes at the time of outburst, this power cascades to higher-order modes \citep[for example][]{laughlin1997, laughlin1998,laughlin1996}.
 
@@
 is significantly less than the change going from $l_{max} = 64$ to 128, with a strong suggestion that this diminution will not proceed much further with higher azimuthal resolution.
-Indeed, a  Richardson extrapolation {\bf (add reference)} of the tabulated values of  $\langle A_{2-7} \rangle / \langle A_{tot} \rangle$ suggests a limiting value of  $\langle A_{2-7} \rangle / \langle A_{tot} \rangle \sim$ 0.43 for increasingly higher resolutions.
-This is displayed graphically in Figure 4, where the fractional power in lower order modes is plotted against the  normalized grid
+Indeed, a  Richardson extrapolation {\bf (add reference) SM: suggest Press et.
+al} \citep{press1992} of the tabulated values of  $\langle A_{2-7} \rangle / \langle A_{tot} \rangle$ suggests a limiting value of  $\langle A_{2-7} \rangle / \langle A_{tot} \rangle \sim$ 0.43 for increasingly higher resolutions.
+This is displayed graphically in figure 4, where the fractional power in lower order modes is plotted against the  normalized grid
 spacing for each value of $l_{max}$. The grid spacing  is normalized such that spacing of the $l_{max}$ = 512 grid
@@
 More than half of the modal power in a converged disk resides in higher-order modes, modes that help define a state of gravitoturbulence.
-The fact that the $l_{max} =$ 512 simulation has $\langle A_{2-7} \rangle / \langle A_{tot} \rangle$ near the convergence value lends some confidence in the use of the $l_{max} =$ 512 model in examining the convergence of $\alpha$.
+{\bf SM: I'm not sure this is accurate, gravitoturbulence isn't really turbulence (at least as it's normally defined) and the high order modes aren't really turbulence either} The fact that the $l_{max} =$ 512 simulation has $\langle A_{2-7} \rangle / \langle A_{tot} \rangle$ near the convergence value lends some confidence in the use of the $l_{max} =$ 512 model for the studie of effective $\alpha$.
 
@@
 The high-order modes tend to produce pockets of local GI activity, which can cancel each other
-over the whole disk, leading to the expectation that  gravitational torques in the disk will decrease as $l_{max}$ is increased.
+over the whole disk, {\bf SM: I think there's a sentence missing here explaining that higher resolution leads to more of the amplitude being concentrated in high order modes, while the total amplitude stays the same.} leading to the expectation that  gravitational torques in the disk will decrease as $l_{max}$ is increased.
 This can be tested by examining the torque contributions from  various Fourier modes.
@@
 where $a_{\phi m} =\slantfrac{1}{\pi} \int \rho \cos(m\phi){\rm d}\phi$ and $b_{\phi m} =\slantfrac{1}{\pi} \int \rho \sin(m\phi){\rm d}\phi$, and only the gravitational potential produced by the mass distribution given by $\rho_m$ is included in $\Phi$.
-{\bf (Scott: Equation (6) is the form you use in your chapter 3, but it
-does not appear to be correct; it does not yield a correct value for $m=0$)}
 
-Figure 5 displays $\langle {\rm \bf C}_{Ztot} \rangle$, the time-averaged gravitational torque summed over all disk modes (excluding $m=1$), along with the torques summed over low-order modes, $\langle {\rm \bf C}_{Z(2-n)} \rangle$, where $n = $ 2, 3, 4, and 5, as a function of radius and azimuthal resolution.
-{\bf (Note to all: the version of Figure 5 included with this draft  shows the torques
+Figure 5 {\bf SM: Two things about this figure.
+Are you sure you removed the m=1?
+You would have had to have written your own software or significantly changed what I gave you to do this.
+Also I think the amplitude of the torques are not right if we're really considering a 1$M_\odot$ star.
+These numbers look like the ones for a 0.5$M_\odot$ star} displays $\langle {\rm \bf C}_{Ztot} \rangle$, the time-averaged gravitational torque summed over all disk modes (excluding $m=1$), along with the torques summed over low-order modes, $\langle {\rm \bf C}_{Z(2-n)} \rangle$, where $n = $ 2, 3, 4, and 5, as a function of radius and azimuthal resolution.
+{\bf (Note to all: the version of figure 5 included with this draft  shows the torques
 time-averaged over for four different time-frames.  I did this because I wanted to see if the disk was 
@@
 yet in an asymptotic state - or further from it than at later times - I'm not sure this is
-a significant difference.  Looking forward to Figure 7, it can be argued that the
+a significant difference.  Looking forward to figure 7, it can be argued that the
 disk is reasonably settled, in a statistical sense, shortly after 12 ORP.
@@
 In this case the scale is doubled.
-Two features are immediately apparent: 1) the total gravitational torque is smaller for higher resolution simulations, and 2) the total torque is dominated by low-order modes.
+Two features are immediately apparent: 1) the total gravitational torque is smaller for higher resolution simulations, and 2) the total torque is dominated by low-order modes for all azimuthal resolutions.
 Both reflect the fact that a while higher fraction of the total power in larger $l_{max}$ simulations is in higher order modes, these modes make only minimal contributions to the gravitational torque.
@@
 \begin{equation}
-\alpha_{eff}(\varpi) = \left| \frac{{\rm d} \ln \Omega}{{\rm d} \ln \varpi} \right|^{-1} 
+\alpha_{\rm{eff}}(\varpi) = \left| \frac{{\rm d} \ln \Omega}{{\rm d} \ln \varpi} \right|^{-1} 
 \frac {\langle T_{\varpi\phi}^{grav}\rangle + \langle T_{\varpi\phi}^{Reyn}\rangle}
@@
 The denominator $\Sigma c_c^2$ was evaluated as $\int_z \rho c_c^2 {\rm d}z$.
-{\bf (Scott: please provide some text explaining why this choice was made for
-${\Sigma c_s^2},$)}
+{\bf SM: I need to rewrite this entire paragraph}
 
@@
 \end{equation}
-{\bf (There is a problem with this equation)}
-Figure 6 shows the effective $\alpha$s for the four different grid resolutions, time averaged from 12 to 19 ORPs.
-For comparison, the predicted values by \citet{gammie2001} for the strongly self-gravitating limit (lower curve) and the non-self-gravitating limit (upper curve) are shown.
-Note that predictions from are based upon a thin
+Figure 6 shows the effective $\alpha$s for the four different grid resolutions, time averaged from 12 to 19 ORPs.{\bf SM: is this 18 or 19 ORPs?} For comparison, the predicted values by \citet{gammie2001} for the strongly self-gravitating limit (lower curve) and the non-self-gravitating limit (upper curve) are shown.
+Note that predictions from {\bf ??
+is this \citep{gammie2001}?} are based upon a thin
 disk approximation, where the local balance of heating and cooling dominates the energetics.
@@
 As the azimuthal resolution increases, the effective $\alpha$ in the GI active region decreases.
-This is due to the decrease in the total gravitational torque, as illustrated in Figure 5.
+This is due to the decrease in the total gravitational torque, as illustrated in figure 5.
 This decrease in gravitational torque is due, in turn, to the shift of non-axisymmetric power from low-order global modes, which produce large torques, to high-order local modes, which can cancel each other out when integrated over the whole disk.
@@
 Although the average $\alpha$ values have not converged to a single value for my highest two resolutions, the percentage change continues to decrease.
-In addition, Figure 6 shows that the $l_{max} =$ 256
+In addition, figure 6 shows that the $l_{max} =$ 256
 and 512 simulations have qualitatively similar time averaged profiles over a large range of disk radii.
@@
 Although the effective $\alpha$ curves do correspond to the Gammie prediction, the results from these simulations differ from Gammie's in one significant way.
-The local shearing box simulations performed by \citet{gammie2001} show that the total effective $\alpha$ is composed of a gravitational stress component and a Reynolds stress component that are nearly equal (see Gammie (2001) Figure 3).
+The local shearing box simulations performed by \citet{gammie2001} show that the total effective $\alpha$ is composed of a gravitational stress component and a Reynolds stress component that are nearly equal (see Gammie (2001) figure 3).
 However, all of the analyses
@@
 % \tablecomments{
-% Table 1. The $\alpha$ of Figure 1, averaged over the radial region from 10 -- 40 AU, tabulated for different 
+% Table 1. The $\alpha$ of figure 1, averaged over the radial region from 10 -- 40 AU, tabulated for different 
 % azimuthal resolutions, along with amplitudes of Fourier components for non-axisymmetric density structures. 
@@
 
-\begin{deluxetable}{cccccc}\label{tbl:ams}
+\begin{deluxetable}{cccccc}
 \tablecolumns{5} 
@@
 }
+\label{tbl:ams}
 \end{deluxetable}
@@
 \begin{figure}\label{Final_Q}
-%\plotone{Final_Q_Values.eps}
-\caption{Final Q values - {\bf \it NEEDS REPLOTTING}
-}
+\plotone{figures/qcompareres.eps}
+\caption{Toomre Q distribution versus radius for each of the four resolutions at t $\sim$ 18 ORPs}
 \end{figure}
@@
 \begin{figure}\label{fig:DensityPlots}
-%\plotone{Figure_1.eps}
+%\plotone{figures/Figure_1.eps}
 \caption
@@
 \begin{figure}\label{fig:Am_vs_log_m}
-%\plotone{Figure_3.eps}
+%\plotone{figures/Figure_3.eps}
 \caption
@@
 \begin{figure}
-%\plotone{Power_Extrapolation.eps}
+%\plotone{figures/Power_Extrapolation.eps}
 \caption
@@
 \begin{figure}
-%\plotone{Torque_Array.eps}
+%\plotone{figures/Torque_Array.eps}
 \caption
@@
 %\begin{figure}
-%\plotone{Mode_CascadeRatios.pdf}
+%\plotone{figures/Mode_CascadeRatios.pdf}
 %\caption
@@
 \label{fig:alpha_v_radius}
-%\plotone{Figure_2.eps}
+%\plotone{figures/Figure_2.eps}
 \caption
@@
 \begin{figure}
-%\plotone{Alpha_variations_stack512.eps}
+%\plotone{figures/Alpha_variations_stack512.eps}
 \caption

### 2010-12-08 Paper_1.tex
Rearranged figure labels.
	modified:   Paper_1.tex
	renamed:    figures/qcompareres.eps -> figures/Figure1.eps
	new file:   figures/Figure2.eps
	new file:   figures/Figure3.eps
	new file:   figures/Figure4.eps
	new file:   figures/Figure5.eps
	new file:   figures/Figure6.eps
	new file:   figures/Figure7.eps
Tom's edits
	modified:   Paper_1.tex

@@
 
-\title{Convergence Studies and Effective Alphas for Disks with Gravitational Instabilities.
+\title{Convergence Studies on Mass Transport in Disks with Gravitational Instabilities.
  \\
@@
 
-Using 3D hydrodynamic simulations of a cooling protostellar disk undergoing gravitational instabilities (GIs), we compute the effective \citet{shakura1973} $\alpha$ due to gravitational stresses and compare them to predictions from an analytic local theory for thin disks by \citet{gammie2001}.
-Our goal is to determine how accurately a locally defined $\alpha$ can characterize mass and angular momentum transport by GIs in disks.
-We examine disks with an imposed constant global cooling time \citep{mejia2005}.
-Models were computed with grids of different spatial resolution to investigate how the computed $\alpha$ is affected by numerical resolution.
+We conducted a 3D hydrodynamic convergence study of a protostellar disk
+subject to a constant global cooling time 
+and to gravitational instabilities (GIs). 
+The disk was followed during its asymptotic state, where
+heating and cooling are in balance.
+Four simulations were conducted, identical except for the number of azimuthal grid 
+points, $l_{max} = $ 64, 128, 256 and 512, used in the calculations. 
+The strengths of the azimuthal modes, gravitational torques, and the effective \citet{shakura1973} $\alpha$ 
+arising from gravitational stresses were determined for each resolution.  These
+were used to examine the resolution-dependence of each quantity
+and to evaluate both the converged modal power distribution and the effective value of $\alpha$
+as $l_{max} \rightarrow \infty$.   Power was found in all modes and 
+the power summed over all modes was essentially independent of resolution.
+The modal distribution of power, however, was significantly affected by resolution.  
+The size of the azimuthal grid, $l_{max}$,
+limits the number of modes to $l_{max}/2$.  Higher resolution simulations therefore
+have more modes available to them.  Power is spread across all available modes.  As a consequence,
+higher resolution simulations have a larger fraction of their total power in high-order modes.
+High-order modes
+manifest themselves more locally than low-order modes, therefore 
+greater resolution leads to smaller gravitational stresses.
+As a result, higher resolution simulations experience 
+weaker gravitational torques than lower resolution simulations.   The effective  $\alpha$
+also depends upon the magnitude of the stresses.  Thus
+$\alpha_{eff}$ is also a function of grid resolution, decreasing as  $l_{max}$ increases.
+The converged $\alpha_{eff}$ is consistent with
+predictions from an analytic local theory for thin disks by \citet{gammie2001}, 
+but only over many dynamic times when averaged over a substantial volume of the disk.
+Results obtained in the $l_{max} = 512$ simulations differ with converged values
+by less than 20\%.
 
@@
 
-While detailed three-dimensional hydrodynamic modeling of protostellar disks can provide considerable insight into the physical and thermal states of these disks, computing resources limit such studies to snapshots spanning timescales much shorter than a disk's evolutionary lifetime.
-In contrast, approaches that replace detailed physics with statistical parameters that encapsulate this physics allow for calculations covering much longer timeframes.
-Both approaches can play important roles in broadening our understanding of these complex nonlinear systems, provided a connection between the detailed physics and statistical parameters can be demonstrated and understood.
-In what follows, we examine the applicability of an $\alpha$-disk formulation \citep{shakura1973} to  the
+Detailed three-dimensional hydrodynamic modeling of protostellar disks can provide considerable insight into the physical and thermal states of these disks.
+However,  computing resources limit these studies to snapshots spanning timescales much shorter than a disk's evolutionary lifetime.
+The level of spatial resolution (and hence temporal resolution, due to the Courant condition) used in calculations can play a deterministic role in interpretations of models; while lower resolutions allow more extensive spatial and temporal studies, this benefit can come at the sacrifice of important physics.
+Convergence studies provide a means
+of deriving important outcomes without the expense of very high resolution simulations.
+
+In contrast to sophisticated calculations that include detailed physics, 
+approaches exist that replace detailed physics with statistical parameters that encapsulate this physics.  This
+allows for calculations covering much longer timeframes.  Both detailed and 
+parametric
+approaches can play important roles in broadening our understanding of complex nonlinear systems, provided a connection between the detailed physics and statistical parameters can be demonstrated and understood.
+In what follows, we perform convergence
+tests on hydrodynamical disk simulations to examine the applicability of an $\alpha$-disk 
+formulation \citep{shakura1973} to  the
 evolution of protostellar disks subject to gravitational instabilities.
 
-Gravitational instabilities (GIs) play an important, {\bf SM: I'm not sure this is true in general.
-Perhaps reword to ``GIs can play an important role...and may dominate the structure for certain phases of evolution''}and at times dominant, role in driving the structural and thermal
-evolution of  protostellar disks \citep[for a review, see][]{durisen2005}.
-%In particular, torques arising from GIs provide a very effective means of redistributing matter and 
-%angular momentum within these disks 
-%\citep{Lod04, Lod05, Vor05, Vor06, Bol09}.  
+%Here $\alpha$ is a dimensionless parameter that
+%provides a measure of viscosity, without providing a detailed specification of the viscous mechanism.
+
+Gravitational instabilities (GIs) can play an important, and at times dominant, role in driving the structural 
+evolution of  protostellar disks \citep[for reviews, see][]{durisen2005, durisen2010}. {\bf (We need the Durisen 2010 (2011?)
+review article reference)}
 Thermal processes play the primary role in regulating the amplitude and outcome of these instabilities \citep{pickett1998,pickett2000,mejia2005,nelson1998,nelson2000a}.
-A a disk's susceptibility to GIs is parameterized by the Toomre $Q$-parameter \citep{toomre1981};  $Q = c_s \kappa / \pi G \Sigma$, where $c_s$ is the sound speed, $\kappa$ is the epicyclic frequency ($\sim$ the rotation frequency
+A disk's susceptibility to GIs can be parameterized by the Toomre $Q$-parameter \citep{toomre1981};  $Q = c_s \kappa / \pi G \Sigma$, where $c_s$ is the sound speed, $\kappa$ is the epicyclic frequency ($\sim$ the rotation frequency
 $\Omega$ in a nearly Keplerian disk), and $\Sigma$ is the disk surface mass density.
 For $Q \al$  1.5 to 1.7,  small density perturbations in a disk grow exponentially on a time scale comparable to the rotation period.
-These perturbations manifest themselves as multi-arm spirals with a predominantly trailing pattern that transports angular momentum outward by gravitational torques \citep[e.g.]{larson1984,boss1984,durisen1986,papaloizou1991, laughlin1994,nelson1998,pickett1998}.
-The long range nature of these torques cause nonlinear
-spiral structures to develop over a broad range of radii \citep{laughlin1998,nelson1998,nelson2000a,pickett1998, pickett2000,pickett2003}.
-{\bf SM: not sure of the meaning here.
-A low Q is what causes GIs to develop over a broad range of radii.
-Perhaps you mean the long range nature can cause coherent spiral structures to form with large radial extent?}
+These perturbations manifest themselves as multi-arm spirals with a predominantly trailing pattern that transports angular momentum outward by gravitational torques \citep{larson1984,boss1984,durisen1986,papaloizou1991, laughlin1994,nelson1998,pickett1998}.
 
-Numerous studies have shown that thermal physics control GI amplitudes by a {\it balance} of heating and cooling \citep[for example][]{tomley1991,tomley1994,pickett1998,pickett2000, pickett2003,gammie2001,boss2002,rice2003b,mejia2005,boley2006,boley2007, stamatellos2008,cossins2009}.
+Numerous studies have shown that thermal physics control GI amplitudes by a {\it balance} of heating and cooling \citep[e.g.,][]{tomley1991,tomley1994,pickett1998,pickett2000, pickett2003,gammie2001,boss2002,rice2003b,mejia2005,boley2006,boley2007, stamatellos2008,cossins2009}.
 Heating by GIs tends to increase $c_s$, thus increasing $Q$.
@@
 GI-activity can sustain this balance  at a relatively constant,
-but unstable, value of $Q$ \citep{paczynski1978, lin1981,goldreich1965},and 
-the disk can exist in a state of quasi-equilibrium. In this state cooling is 
+but unstable, value of $Q$ \citep{paczynski1978, lin1981,goldreich1965}, and 
+the disk can exist in a state of quasi-equilibrium. In this state, cooling is 
 balanced by heating arising from the inward flow of matter and shocks induced by the GIs
-\citep{gammie2001, lodato2004, rice2005, cossins2009, vorobyov2010}.
+\citep{gammie2001, lodato2004, rice2005, boley2006, cossins2009, vorobyov2010}.
 
 Fully nonlinear hydrodynamic simulations of low-$Q$ disks show that multiple
-modes can become unstable in the linear regime \citep[e.g.][]{nelson1998, pickett1998} {\bf additional refs?}.
+modes can become unstable in the linear regime \citep{nelson1998, pickett1998, lodato2004,boley2006,cossins2009}.
 Even if the initial growth is dominated by one mode, numerous modes usually appear in the nonlinear regime.
-As a consequence, if a disk achieves a quasi-steady balance between heating and cooling, then there tends to be power at all $m$-values resolved by a given numerical method \citep{mejia2005,boley2006}.
-In this  {\it asymptotic state}, GI-active disks develop overlapping density structures of different strengths, geometries and coherence, producing strongly nonlinear turbulence that pervades the entire disk \citep[e.g.][]{pickett2003, mejia2005,boley2006}.
-\citet{gammie2001} refers to this state as {\it gravitoturbulence}. Gravitoturbulence provides a possible bridge between the detailed physics of GIs and the viscous transport parameter $\alpha$.
+As a consequence, if a disk achieves a quasi-steady balance between heating and cooling, referred to as the {\it asymptotic state},
+then power will exist at all $m$-values resolved by a given numerical method \citep{mejia2005,boley2006}.  
+In the asymptotic state, GI-active disks develop overlapping density structures of different strengths, geometries and coherence, producing strongly nonlinear turbulence that pervades the entire disk \citep[e.g.][]{pickett2003, mejia2005,boley2006}.
+\citet{gammie2001} refers to this state as {\it gravitoturbulence}.
+Gravitoturbulence provides a possible bridge between the detailed physics of GIs and the viscous transport parameter $\alpha$.
+Here convergence studies are important because
+higher order modes dissipate energy on shorter wavelengths than lower order modes.  The mix of low and high order modes available to a disk is set by the resolution of the simulation.
 
-\citet{shakura1973} first parameterized turbulent transport in a steady-state disk with the  parameter $\alpha$, where the source of the turbulence is unspecified.
-Although their original work was directed towards X-ray bright
-accretion disks around black holes, this heuristic approach has proved fruitful in many astrophysical situations by allowing for analytic solutions and relatively easy numerical modeling over time-spans inaccessible to detailed modeling.
-In the $\alpha$ disk formalism, when the accretion rate $\dot{M}$ is constant in both time and radius, $\dot{M} = 3\pi\alpha c_s \Sigma H$ where $c_s$ is the sound speed, $\Sigma$ the surface density, and $H$ the disk scale height.
-The $\alpha$ formalism makes no assumption about the nature of the angular momentum transport, only that {\it heating and cooling are dominated by local processes}{\bf SM: This is not my understanding of the $\alpha$ formalism, I don't think it says anything about cooling.
-It also means that ang mom transport happens locally and that this causes local heating}; the validity of an $\alpha$-disk picture requires self-consistency of local energy dissipation.
+\citet{shakura1973} proposed turbulence as the primary source of effective viscosity in gas disks
+and parameterized turbulent transport in a steady-state disk with the  parameter $\alpha$.  The source of the 
+turbulence in their $\alpha$-disk formulation is unspecified. Although their original work was directed towards X-ray bright
+accretion disks around black holes, this heuristic approach has proven fruitful in many astrophysical situations by allowing for analytic solutions and relatively easy numerical modeling over time-spans inaccessible to detailed modeling.
+In the
+$\alpha$-disk formalism, when the accretion rate $\dot{M}$ is constant in both time and radius, $\dot{M} = 3\pi\alpha c_s \Sigma H$ where $\Sigma$ is the surface density and $H$ represents
+the disk scale height. The $\alpha$ formalism makes no assumption about the nature of the angular momentum transport, only that {\it angular momentum transport
+and heating are dominated by local processes}. Lin and Pringle (1987, 1990) were among the first to suggest
+that GI-induced transport could be described within a viscous
+framework.  However, the  validity of an $\alpha$-disk picture requires self-consistency of local energy dissipation. 
+The locality of dissipation in GI-active disks has not been rigorously tested and remains an open question.
 
-Dissipative locality {\bf SM: need to explain what is meant by this more precisely} in GI-active disks has not been rigorously tested and remains an open question. 
 Several studies support the idea that GI transport in real disks is, in many important respects, an intrinsically global phenomenon and cannot be properly treated by a local $\alpha$-like prescription \citep{laughlin1996, balbus1999, lodato2005, mejia2005, boley2006, cai2008}.
 This almost certainly applies to the embedded phase of both high and low-mass stars when infall from the collapsing protostellar cloud feeds mass into the disk at a rapid
-rate \citep[e.g.]{yorke1993, laughlin1994, yorke1999,vorobyov2005,vorobyov2006}, and disks observed in the embedded phase can be quite massive relative to their central stars \citep[e.g.][]{osorio2003}.
-However, despite the long-distance nature of gravitational torques, theoretical and numerical results suggest that in protostellar disks where heating and cooling are in balance it may be valid to represent GI-induced transport by an {\it effective viscosity}, a local process \citep[e.g.][]{gammie2001, rice2005, lodato2004}.
-Indeed, \citet{gammie2001} derived an effective alpha for the case of a thin 2D gravitationally unstable disk where the heating due to local turbulent dissipation caused by GIs was balanced by the local cooling rate.
-Modifications to this derivation have since been proposed by {\bf (Tom: include other derivations and refs)}.
-Recent simulations \citep[e.g.]{cossins2009, vorobyov2010} suggest that, at least in the case of imposed cooling, $\alpha_{\rm{eff}}$ roughly converges to the value predicted by \citeauthor{gammie2001}.
+rate \citep[e.g.]{yorke1993, laughlin1994, yorke1999,vorobyov2005,vorobyov2006}.  
+However, despite the long-distance nature of gravitational torques, theoretical and numerical results suggest that in protostellar disks where heating and cooling are in balance, i.e., disks subject to gravitoturbulence,
+ it may be valid to represent GI-induced transport by an {\it effective viscosity}, a local process \citep{gammie2001, lodato2004, rice2005, cossins2009, vorobyov2010}.  In particular, 
+an $\alpha$-viscosity may be consistent
+with transport in GI acitve disks for disk to star mass ratio $M_d/M_* \le 0.2$ -- 0.3; GI-active disks
+in systems with larger $M_d/M_*$ are dominated by lower-order modes that act in a more global sense and thus cannot be well represented by an $\alpha$-approach
+\citep{lodato2004, cossins2009, vorobyov2010}.
 
-{\bf (Tom: rework and finish this paragraph, or toss it)}
-Three possibilities exist: 1) GI-transport is predominately a long-range effect and thus not 
-amenable to an $\alpha$-disk representation, 2) GI-transport can be well represented as
-a short-range effect and thus is directly amenable to an $\alpha$-disk representation,
-If long-range energy transport and/or angular momentum transport are negligible,  or if
-the long-range transport could be shown to represent a known fraction of the 
-***. 
-Identifying the principal angular momentum transport mechanism
-and how it behaves is critical to understanding whether disks can be described by an alpha disk model.
+\citet{gammie2001} derived an effective alpha for the case of a thin 2D gravitationally unstable disk where the heating due to local turbulent dissipation caused by GIs was balanced by the local cooling rate.
+Assuming a razor thin gravitationally unstable disk where the heating due to local turbulent dissipation caused by
+GIs was balanced by a local cooling rate of $t_{cool}$.
+Gammie found
+$\alpha_{eff} = [\gamma(\gamma - 1) \slantfrac{9}{4}\Omega t_{cool}]^{-1}$,
+where $\gamma$ is the two dimensional adiabatic index and $t_{cool}$ is a cooling parameter. 
+The two dimensional index relates
+to the three dimensional index, $\Gamma$, by 
+$\gamma = (3\Gamma - 1) / (\Gamma + 1)$
+for a non-self-gravitating
+disk and
+$\gamma = 3 - 2/\Gamma $
+for a strongly self gravitating disk.   \cite{lodato2004} derived essentially the same result for their 3D study.
+Motivated by
+the \cite{laughlin1996}, \cite{gammie2001}, and \cite{lodato2004, lodato2005} results 
+that demonstrated differences in how low and high order modes manifest themselves, Kratter et al. (2008) 
+proposed $\alpha_{eff} = (\alpha_{short}^2 + \alpha_{long}^2)^{1/2}$, where $\alpha_{short}$ and $\alpha_{long}$
+are functions of $Q$ and $M_d/M_*$ representing 
+the different wavelength regimes expected to dominate for different values of the mass ratio and cooling rate. 
+{\bf (Scott: need to add kratter2008 to the bib file.  The reference is Kratter, K.M., Matzner, C. D., Krumholz, M.R., 
+2008. ApJ 681, 375)}
+Fully 3D simulations \citep[e.g.][]{lodato2004, cossins2009} suggest that, at least in the case of imposed cooling, $\alpha_{\rm{eff}}$ roughly converges to the value predicted by \citeauthor{gammie2001}.
 
-{\bf (Tom: add text concerning need for convergence approach)}
-Here we report the results of a convergence study designed to determine whether GI-active disks, chaotic by nature, achieve statistical equilibrium; if so, to what level do these disks approximate the gravitoturbulence of \citet{gammie2001} and with what distribution of $\alpha_{eff}$; and to characterize how the GI cascades from low-order to high-order modes.
+We address these, and similar questions, using a grid-based, finite-difference 3D hydrodynamics code
+with self gravity.
+For computational grids of a fixed {\it physical} size,
+grids with more elements provide finer meshes, and thus a more accurate solution. 
+However, increased accuracy is offset by temporal sacrifices.
+The number of equations solved at each timestep 
+increases as the grid size is increased.  At the same time, the Courant condition 
+forces shorter timesteps.  A mesh convergence study 
+provides a common approach to balancing accuracy and computing resources.
+In the following sections we follow the evolution of a disk subject to GIs
+using four different
+resolutions in the azimuthal direction.  The number of azimuthal grid points used in simulations
+may be particularly important, because higher azimuthal resolution 
+allows GI power to spread to higher-order modes 
+that behave more locally. We test this by examining the power in various 
+modes to see how they are affected by the choice of grid.
+In the process, 
+we report the results of a convergence study designed to determine whether GI-active disks, chaotic by nature, achieve statistical equilibrium; if so, to what level do these disks approximate the gravitoturbulence of \citet{gammie2001} and with what distribution of $\alpha_{eff}$; and to characterize how the GI cascades from low-order to high-order modes.
 To this end, we follow the evolution of a GI-active disk, subject to constant cooling, in its asymptotic state, where cooling and heating are essentially balanced.
@@
 
-{\bf (Tom: add paragraph providing outline of paper)}
+The balance of this paper is organized as follows.  Section 2 provides the details of the numerical approach and defines the models.  An analysis of the modal structure of the simulations 
+is found in \S3.1, the gravitational torques arising from these modes follows in \S3.2.  Effective converged alphas are 
+discussed in \S3.3.  Section 3.4 deals with the degree to which selected grid sizes lead to results consistent with
+convergence.  Fragmentation is discussed in \S3.5.  A summary of results is found in \S4.
 
-%Nonlinear evolutions of disks undergoing GIs, even local thin-disk treatments, 
-%\citep[e.g.,][]{Joh03} 
-%are sufficiently complex that they must be done numerically,  
-%typically using a finite element approach, where
-%a set of coupled equations is solved simultaneously on a computational grid.  
-%Computational constraints prohibit a full examination of GI unstable disks over 
-%the life of the disk.
-%Indeed, existing computational resources limit 
-%detailed three-dimensional 
-%simulations of disk evolution to tens of
-%outer disk rotations.  
 %We cannot yet follow the entire dynamical evolution of a 
@@
 %can extrapolate the consequences over much longer times (e.g., Boley et al 2006).
-
-%For 
-%grids of a fixed {\it physical} size,
-%grids with more elements provide finer meshes, which typically produce a more accurate solution. 
-%However, this increased accuracy if offset by temporal sacrifices.
-%The number of equations to be solved at each timestep 
-%increases as the grid size is increased.  At the same time, the Courant condition 
-%forces these timesteps to be shorter in duration.  A mesh convergence study 
-%provides a common approach to balancing accuracy and computing resources.
-%In the following sections we follow the evolution of a disk subject to GUs
-%using four different
-%resolutions in the azimuthal direction.  The number of azimuthal grid points used in simulations
-%may be particularly important, because higher resolution 
-%in azimuth allows GI power to spread to higher-order modes 
-%that behave more locally. We test this by examining the power in various 
-%modes to see how they are affected by the choice of grid.
 
@@
 
-Figure \ref{Final_Q} shows $Q$, as a function of radius, at $t \sim$ 18 ORP, while table \ref{tbl:ams} lists $Q_{avg}$, the value of $Q$ of for the asymptotic disk, time-averaged of 12 -- 19 ORP and spatially averaged over 10 -- 40 AU, as a function of radius for each resolution,
+Figure 
+%\ref{qcompareres}
+\ref{fig:Final_Q} 
+shows $Q$, as a function of radius, at $t \sim$ 18 ORP.
+Table \ref{tbl:ams} lists $Q_{avg}$, the value of $Q$ of for the asymptotic disk, time-averaged of 12 -- 19 ORP and spatially averaged over 10 -- 40 AU, as a function of radius for each resolution.
+From $\sim$ 12 -- 50 AU, $Q \sim 1.0$ -- 1.4 and thus the disk is subject to the growth of nonaxisymmetric structures over this full range.
+Interior to $\sim$ 27 AU, the value of $Q$ shows no dependence on
+the azimuthal resolution.  From 27 -- 50 AU, lower resolution simulations display lower values of $Q$.
 
@@
 Figure \ref{fig:DensityPlots}, shows midplane and radial densities  at $t \sim 18$ ORPs for the $l_{max} =$ 64, 128, 256, and 512 simulations.
+Two features are readily apparent; disks run at higher resolution are thicker than lower resolution disks {\bf (anyone want to comment on this?)} and the character of density perturbations differ with resolution.
 The lower resolution (smaller $l_{max}$) simulations display longer wavelength spiral structures, with larger amplitudes and more coherence than those seen in higher resolution simulations.
@@
 
-Here  $\rho_o$ is the axisymmetric (zero-order)  component of the density.
+Here  $\rho_o$ is the axisymmetric component of the density.
 In what follows, $A_m$ is averaged over time to suppress fluctuations on the dynamic time scale due to the chaotic nature of GIs in the asymptotic  phase.
@@
 The Fourier amplitudes cannot be accurately measured at resolutions smaller than $l_{max}/2$ \citep{shannon1984}, therefore $A_m$ is limited to modes $m \le l_{max}/2$.
-Because the central star is artificially fixed to the grid center, the $m = 1$ mode cannot be be realistically treated; it is excluded from $\langle A_{tot} \rangle$.
+Because the central star is artificially fixed to the grid center, the dynamics of the $m=1$ mode may not
+be accurately treated. For this reason, it is excluded from $\langle A_{tot} \rangle$.
 
@@
 
-Previous studies have shown that disks in the asymptotic state have power at all resolvable $m$-values  \citep[e.g.][]{mejia2005,boley2006}.
+Previous studies have shown that disks in the asymptotic state have power at all resolvable $m$-values  \citep{lodato2004,mejia2005,boley2006,cossins2009}.
 Table \ref{tbl:ams} lists $\langle A_{tot} \rangle$ along with $\langle A_{2-7} \rangle  /  \langle A_{tot} \rangle$,  the fraction of the total power falling into low-order ($m = 2$ -- 7) modes, for each resolution.
@@
 
-The shift of power from lower to higher order modes as the resolution of the computational grid is increased arises from the fact that, while $\langle A_{tot} \rangle$ is approximately independent of $l_{max}$, {\bf SM: Need some explanation as to why this is the case.} the number of degrees of freedom available for this power, $l_{max}/2$, increases linearly with the resolution.
+The shift of power from lower to higher order modes as the resolution of the computational grid is increased arises from the fact that, while $\langle A_{tot} \rangle$ is approximately independent of $l_{max}$,  the number of degrees of freedom available for this power, $l_{max}/2$, increases linearly with the resolution.
 Thus power is naturally spread from lower order to higher order modes as the available $m$-values increases.
@@
 is significantly less than the change going from $l_{max} = 64$ to 128, with a strong suggestion that this diminution will not proceed much further with higher azimuthal resolution.
-Indeed, a  Richardson extrapolation {\bf (add reference) SM: suggest Press et.
-al} \citep{press1992} of the tabulated values of  $\langle A_{2-7} \rangle / \langle A_{tot} \rangle$ suggests a limiting value of  $\langle A_{2-7} \rangle / \langle A_{tot} \rangle \sim$ 0.43 for increasingly higher resolutions.
-This is displayed graphically in figure 4, where the fractional power in lower order modes is plotted against the  normalized grid
+Indeed, a  Richardson extrapolation \citep{press1992} of the tabulated values of  $\langle A_{2-7} \rangle / \langle A_{tot} \rangle$ suggests a limiting value of  $\langle A_{2-7} \rangle / \langle A_{tot} \rangle \sim$ 0.42 for increasingly higher resolutions.
+This is displayed graphically in Figure \ref{fig:Power_Extrap}, where the fractional power in lower order modes is plotted against the  normalized grid
 spacing for each value of $l_{max}$. The grid spacing  is normalized such that spacing of the $l_{max}$ = 512 grid
@@
 Convergence occurs as the normalized spacing goes to zero.
-The Richardson extrapolation value  of $\langle A_{2-7}\rangle  / \langle A_{tot}\rangle  = 0.43$ is plotted at a normalized grid spacing  equal to zero.
-More than half of the modal power in a converged disk resides in higher-order modes, modes that help define a state of gravitoturbulence.
-{\bf SM: I'm not sure this is accurate, gravitoturbulence isn't really turbulence (at least as it's normally defined) and the high order modes aren't really turbulence either} The fact that the $l_{max} =$ 512 simulation has $\langle A_{2-7} \rangle / \langle A_{tot} \rangle$ near the convergence value lends some confidence in the use of the $l_{max} =$ 512 model for the studie of effective $\alpha$.
+The Richardson extrapolation value  of $\langle A_{2-7}\rangle  / \langle A_{tot}\rangle  = 0.42$ is plotted at a normalized grid spacing  equal to zero.
+The majority of the modal power in a converged disk resides in higher-order modes.
+%These modes dissipate on shorter wavelengths than low-order modes and hence act more locally.
 
@@
 
-Coherent density structures in a disk produce gravitational torques.
-Low-order modes in $\Delta\rho / \rho$  have longer wavelengths and, if coherent in radius, have longer lever arms to produce torques.
-On the other hand, high-order modes have relatively short wavelengths which, particularly if they lack coherence, produce more localized effects.
-The high-order modes tend to produce pockets of local GI activity, which can cancel each other
-over the whole disk, {\bf SM: I think there's a sentence missing here explaining that higher resolution leads to more of the amplitude being concentrated in high order modes, while the total amplitude stays the same.} leading to the expectation that  gravitational torques in the disk will decrease as $l_{max}$ is increased.
-This can be tested by examining the torque contributions from  various Fourier modes.
+Low-order spiral modes in $\Delta\rho / \rho$  have longer wavelengths that, if coherent in radius, have longer lever 
+arms to produce torques, while high-order modes have relatively short wavelengths which, particularly if they 
+lack coherence, produce more localized effects. High-order modes tend to produce pockets of local GI activity, 
+which can cancel each other over the whole disk.  The finding that the fractional power in high-order modes increases
+with resolution, while the total power stays the same, leads to
+ the expectation that  gravitational torques in the disk will decrease as $l_{max}$ is increased. This can be tested by examining the torque contributions from  various Fourier modes.
 
@@
 
-Figure 5 {\bf SM: Two things about this figure.
-Are you sure you removed the m=1?
-You would have had to have written your own software or significantly changed what I gave you to do this.
-Also I think the amplitude of the torques are not right if we're really considering a 1$M_\odot$ star.
-These numbers look like the ones for a 0.5$M_\odot$ star} displays $\langle {\rm \bf C}_{Ztot} \rangle$, the time-averaged gravitational torque summed over all disk modes (excluding $m=1$), along with the torques summed over low-order modes, $\langle {\rm \bf C}_{Z(2-n)} \rangle$, where $n = $ 2, 3, 4, and 5, as a function of radius and azimuthal resolution.
-{\bf (Note to all: the version of figure 5 included with this draft  shows the torques
-time-averaged over for four different time-frames.  I did this because I wanted to see if the disk was 
-"settling down" with time.  I think the figure suggests  that the 12-14 ORP disk is not
-yet in an asymptotic state - or further from it than at later times - I'm not sure this is
-a significant difference.  Looking forward to figure 7, it can be argued that the
-disk is reasonably settled, in a statistical sense, shortly after 12 ORP.
+Figure 5  displays 
+ time-averaged torques summed over a number of low-order modes, $\sum_1^n \langle {\rm \bf C}_{Z(n)} \rangle$, $n = $ 1, 2, 3, 4, and 5, and $\langle {\rm \bf C}_{Ztot} \rangle$,
+the total time-averaged gravitational torque arising from modes 1 thru $l_{max}/2$, for four different time-frames,
+as a function of radius and azimuthal resolution. 
+The scale of the $y$-axis in each subplot is the same, with the exception of the $l_{max} = 64$ simulation, 
+where the scale is doubled.  
+Because the star's position is artificially fixed at the center of the computational grid, 
+torques arising from the $m=1$ mode,
+typically 5-10\% of the total torque when averaged over the radial range of the figure,
+should not be considered realistic.  
+Michael \& Durisen (2010) found that freeing the star makes only 10-20\%
+differences,  thus the difference between torques arising from the $m=1$ mode with the star fixed are roughly half that found when the star is free to move {\bf (Scott or Dick: please rework this sentence)}.
+
+We note two results of Figure 5: 1) the total gravitational torque is smaller for higher resolution simulations, and 2) the total torque is dominated by low-order modes for all azimuthal resolutions.
+Both reflect the fact that while a higher fraction of the total power in larger $l_{max}$ simulations is in higher order modes, these modes make only minimal contributions to the gravitational torque.
+As resolution increases, non-axisymmetric power is shifted from low-order global modes, which produce relatively large gravitational torques, to high-order local modes, which tend to cancel each other when integrated over the whole disk.
+
+{\bf (Note to all: Figure 5 shows torques
+time-averaged over for four different time-frames.  I created multiple frames because I wanted to see if torques were 
+diminishing with time, as might be expected if the disk has not achieved its asymptotic state.  
 Dick has told me he only believes the 12-18 ORP result - Dick, I hope 
-I'm not misquoting you).  
+I'm not misquoting you. Should we go with multiple panels representing different time-frames or 
+go solely with the 12 - 18 orp panel?
 }
-The scale of the $y$-axis in each subplot is the same, with the exception of the $l_{max} = 64$ simulation. 
-In this case the scale is doubled.
-Two features are immediately apparent: 1) the total gravitational torque is smaller for higher resolution simulations, and 2) the total torque is dominated by low-order modes for all azimuthal resolutions.
-Both reflect the fact that a while higher fraction of the total power in larger $l_{max}$ simulations is in higher order modes, these modes make only minimal contributions to the gravitational torque.
-As resolution increases, non-axisymmetric power is shifted from low-order global modes, which produce relatively large gravitational torques, to high-order local modes, which tend to cancel each other out when integrated over the whole disk.
 
@@
 Therefore, in order to compare our results with predicted and published results, we must extract two-dimensional information from our three-dimensional disks.
-Following \citet{lodato2004},
+Following \cite{gammie2001} and \citet{lodato2004},
 \begin{equation}
+\label{alpha1}
 \alpha_{\rm{eff}}(\varpi) = \left| \frac{{\rm d} \ln \Omega}{{\rm d} \ln \varpi} \right|^{-1} 
 \frac {\langle T_{\varpi\phi}^{grav}\rangle + \langle T_{\varpi\phi}^{Reyn}\rangle}
-{\Sigma c_s^2},
+{\Sigma c_s^2 },
 \end{equation}
@@
 (or ``Newton'') and hydrodynamic (or ``Reynolds'') stress tensors, respectively, $\Sigma$ the surface density,
-and $c_s$ the azimuthally averaged midplane sound speed.
-Brackets surrounding the stress terms indicate that these are integrated over the $z$-direction.
-The denominator $\Sigma c_c^2$ was evaluated as $\int_z \rho c_c^2 {\rm d}z$.
-{\bf SM: I need to rewrite this entire paragraph}
+and $c_s$ the azimuthally averaged midplane sound speed.  Brackets indicate quantities integrated over the $z$-direction. 
+{\bf (All: In the previous section, brackets were used to denote time averages (torques),  while here brackets are used to denote spatial
+averages.  I welcome suggestions for recasting the notation in either case)}
+The denominator $\Sigma c_c^2$ was evaluated as $\int_z \rho c_c^2 {\rm d}z$. {\bf SM: I need to rewrite this entire paragraph}
 
-Numerical uncertainties render the Reynolds stress difficult to determine.
-Fortunately, this difficulty is mitigated by the fact that previous studies have shown the magnitude of the Reynolds stress to be small relative to the gravitational stress (references?).
-Thus it can be safely neglected in Equation (7).
+%Gammie found
+%\begin{equation}
+%\alpha_{eff} = [\gamma(\gamma - 1) \slantfrac{9}{4}\Omega t_{cool}]^{-1}
+%\end{equation}
+%where $\gamma$ is the two dimensional adiabatic index and $t_{cool}$ is a cooling parameter. 
+%The two dimensional index relates
+%to the three dimensional index, $\Gamma$, by 
+%$\gamma = (3\Gamma - 1) / (\Gamma + 1)$
+%for a non-self-gravitating
+%disk and
+%$\gamma = 3 - 2/\Gamma $
+%for a strongly self gravitating disk.
 
-The gravitational torque from equation \ref{torque1} can be used to determine the
-gravitational stress by
+Numerical uncertainties render Reynolds stresses difficult to determine accurately. 
+This difficulty is mitigated by the fact that 
+previous studies have found Reynolds stresses to be small relative to gravitational stresses -- see, 
+for example, Figure 5 of \cite{lodato2004} and Figure 10 of \cite{boley2006} -- suggesting that this term
+can be safely neglected in the evaluation of $\alpha_{eff}$.   For this reason, 
+here we omit
+$\langle T_{\varpi\phi}^{Reyn}\rangle$ from the calculation of $\alpha_{eff}$. 
+We note that \cite{gammie2001}  found
+Reynolds and gravitational stresses contributed equally to $\alpha_{eff}$ in his study.  However,
+his was a local, shearing-sheet, 2D study.  As such, it was unable to capture global effects, such
+as those seen in the cited 3D studies.
+
+The Newton stresses of equation \ref{alpha1} can be evaluated using the gravitational torque of equation \ref{torque1}, i.e.,
 \begin{equation}
+\label{newtonstress}
 \begin{split}
@@
 \end{equation}
-Figure 6 shows the effective $\alpha$s for the four different grid resolutions, time averaged from 12 to 19 ORPs.{\bf SM: is this 18 or 19 ORPs?} For comparison, the predicted values by \citet{gammie2001} for the strongly self-gravitating limit (lower curve) and the non-self-gravitating limit (upper curve) are shown.
-Note that predictions from {\bf ??
-is this \citep{gammie2001}?} are based upon a thin
-disk approximation, where the local balance of heating and cooling dominates the energetics.
+Figure 6 shows $\alpha_{eff}$, time-averaged from 12 to 18 ORPs, calculated using equations \ref{alpha1} and \ref{newtonstress},
+for the four grid resolutions.  For comparison, 
+$\alpha_{eff}$ predicted by \citet{gammie2001} for disks where the local balance of heating and cooling dominates the energetics.
+The lower curve of Gammie applies for the strongly self-gravitating limit while the upper curve
+is for the non-self-gravitating limit.   The $l_{max} = 512$ curve is consistent with these two limits.  However, 
+the  $l_{max} = 512$ curve does not represent a converged model.
+To determine whether the converged value of $\alpha_{eff}$ is consistent with predictions, we examine 
+the spatially (10-40 AU) and temporally (12-19 ORP) averaged  $\alpha_{avg}$,
+provided in Table 1, for each value of  $l_{max}$.    A Richardson extrapolation applied to these values  
+converges to $\alpha_{avg} = 0.017$ (over the same radial  and time range)
+as the grid spacing goes to zero.   This falls
+between the $\alpha_{avg}$ values for the Gammie curves 0.014 (strongly self-gravitating) and 0.027 (non-self-gravitating),
+with better agreement towards the strongly self-gravitating limit.
+
+At first glance, these results suggest strong agreement with Gammie and thus strong support for dissipative
+locality.  However, a cautionary note is in order.  At any given radius, 
+stresses, and hence $\alpha_{eff}$, are highly variable with time.
+Figure 7 shows the time-variability of $\alpha_{eff}$, 
+normalized to its average value between at $t = 14$ and 18 ORP, at seven radial locations.   
+Also shown is the orbital period at each radius, provided by the length of the line at the end of each subplot. 
+Fluctuations in $\alpha_{eff}$
+on the order of 100\% or more are seen at all radii
+on  time scales range from fractions to integral multiples of the dynamic (orbital) time.  The spatial averaging
+of Figure 6 washes out this local variabilty.
+Thus,
+even in this highly GI-active region, the predictions of Gammie are only valid over many dynamic times 
+when averaged over a substantial volume of the disk.
+
+\subsection{Convergence and $l_{max}$}
+
+The results above permit an assessment of appropriate azimuthal grids for the
+problem under consideration and similar problems.  The $l_{max} = 512$ simulation yields a value 
+for $\alpha_{avg}$ of 0.020.  This  overestimates the converged value of $\alpha_{avg}
+= 0.017$ by $\sim 18$\%.  The $l_{max} = 512$ simulation also overestimates the fractional
+power in the high-order ($m = 2$ -- 7) modes by essentially the same amount.
+Both the gravitational torques and effective alphas are functions of the
+gravitational stresses within the disk.  If the Reynolds stresses are, in fact, negligible relative to the gravitational stresses,
+then the errors in the calculated torques and effective alphas that stem from limited azimuthal resolution 
+should be the same.   In this case, the  $l_{max} = 512$ simulation should also be expected to yield
+gravitational torques that are $\sim 18$\% larger than the converged torques.  For comparison, the
+errors in the $l_{max} = 256$ simulations would exceed 50\%.   For many (most?) problems of interest,
+results obtained using $l_{max} = 512$ should provide reasonable agreement with results obtained 
+with higher resolution.
+
+\subsection{Fragmentation and Effective Alphas}
+
+\noindent
+{\bf (Aaron, this section is for you to write or toss, your decision.  The current text should be thought of as a filler only)}
 
@@
 Lodato \& Rice (2005) postulated that a disk in a state of gravitoturbulence would have locally balanced heating and cooling rates.
-If the heating were due to local dissipation, then $\alpha$ and $t_{cool}$ are related by
+If the heating is due to local dissipation, then $\alpha$ and $t_{cool}$ are related by
 \begin{equation}
-\alpha = \left| \frac{{\rm d}\ln \Omega}{{\rm d} \ln R} \right|^{-2}
+\alpha = \left| \frac{{\rm d}\ln \Omega}{{\rm d} \ln \varpi} \right|^{-2}
 \frac{1}{\gamma(\gamma - 1) t_{cool}\Omega}.
 \end{equation}
-\citet{rice2005} concluded that for a given $t_{cool}$  there exists some maximum stress quantified by $\alpha_{max}$, which if exceeded would cause a disk to fragment.
-The horizontal dot dashed line in figure 6 indicates the $\alpha_{max} \approx 0.06$ found by \citeauthor{rice2005}.
-\citet{clarke2007} have revised this limit to $\alpha_{max} = 0.12$ for cases where $t_{cool}$  varies with time.
-
-Figure 7 shows the time-variation of $\alpha$, normalized to its value at $t = 18$ ORP), as a function of radius and  time.
-This plot suggests that the disk has not yet achieved statistical equilibrium until $\sim $14 ORP {\bf (Everyone:  What do you think?
-We are using averages over12-18 ORP,
-are we comfortable with those limits?)}
-
-\begin{center}
-\line(1,0){250}
-\end{center}
-
-{\bf From this point on the draft is unfinished; the text that follows is extracted from 
-Scott's Chapter 3, with little to no modification, so for now consider it as filler.   
-I am still working on the rest of this section, which I expect to be the concluding section
-of the paper (excluding a summary).  Rather than further delay getting
-the draft to you, I am sending it as is.   I will be adding text and incORPorating results from a 
-few recent papers and will send the rest soon.}
-
-\begin{center}
-\line(1,0){250}
-\end{center}
-
-As the azimuthal resolution increases, the effective $\alpha$ in the GI active region decreases.
-This is due to the decrease in the total gravitational torque, as illustrated in figure 5.
-This decrease in gravitational torque is due, in turn, to the shift of non-axisymmetric power from low-order global modes, which produce large torques, to high-order local modes, which can cancel each other out when integrated over the whole disk.
-In fact, the decrease in the ratio of low-order power to the total power $\langle A_{2-7}\rangle / \langle A_\Sigma \rangle$ closely tracks the decrease in the time averaged $\alpha$ averaged from 10 to 40 AU, which can be seen in table \ref{tbl:ams} as the $\alpha_{avg}$ value.
-Although the average $\alpha$ values have not converged to a single value for my highest two resolutions, the percentage change continues to decrease.
-In addition, figure 6 shows that the $l_{max} =$ 256
-and 512 simulations have qualitatively similar time averaged profiles over a large range of disk radii.
-These profiles also fall within the predicted Gammie curves, with tendency toward the lower (strongly self-gravitating) curve.
-Indeed the $\alpha_{avg}$ value of 0.020 for the $l_{max} = $ 512 simulation is squarely between the $\alpha_{avg}$ values for the Gammie curves 0.014 (strongly self-gravitating) and 0.027 (non-self-gravitating).
-
-Although the effective $\alpha$ curves do correspond to the Gammie prediction, the results from these simulations differ from Gammie's in one significant way.
-The local shearing box simulations performed by \citet{gammie2001} show that the total effective $\alpha$ is composed of a gravitational stress component and a Reynolds stress component that are nearly equal (see Gammie (2001) figure 3).
-However, all of the analyses
-presented here only consider the gravitational stress. When I attempt to measure
-the Reynolds stress, I find it is quite small compared to the gravitational stress. As
-noted previously, the Reynolds stress is very hard to determine in global simulations.
-However, the prediction made by Gammie relies only on the idea that heating from
-GI activity balances cooling. So the exact mix of stresses contributing to the total
-amount of heating could be different depending on the type of simulation, the physics
-considered in the simulation, and the simulation parameters.
+\citet{rice2005} concluded that for a given $t_{cool}$  there exists some maximum stress, quantified by $\alpha_{max}$, which if exceeded would cause a disk to fragment.
+Their predicted value $\alpha_{max} \approx 0.06$ is shown as a horizontal dotted line in Figure 6.
+Note that both the $l_{max} = $ 64 and 128 simulations exceed this limiting value over some radial range, yet neither displayed fragmentation.
+For cases where $t_{cool}$  varies with time, \citet{clarke2007} revised this limit to $\alpha_{max} = 0.12$.
 
@@
 
-When considering the effect of azimuthal resolution on the asymptotic behavior
-of a GI unstable disk I found that without sufficient azimuthal resolution of at least
-$l_{max} =$ 512 gravitational torques and effective gravitational $\alpha$
-will be underestimated
-for disk with a constant global cooling rate. I also found that, when properly resolved,
-more non-axisymmetric power propagates to higher order modes, and the amplitude
-of low-order modes decreases. This results in a decrease of the gravitational torque
-and effective gravitational $\alpha$. For $l_{max} =$ 512 the effective gravitational 
-$\alpha$ measured
-is in very good agreement with the prediction of Gammie (2001). Gammie's 
-prediction follows from an argument of balance of GI heating and cooling, resulting in
-``gravitoturbulence''. However, his detailed simulations found that the torques and
-heating in the disk were due in equal parts to gravitational stresses and Reynolds
-stresses, this is in contrast to my measurements which only consider gravitational
-stresses.
+\noindent
+{\bf (All: My style of writing papers is to write the summary after everything is completed.  Therefore, I have left this
+section blank.)}
 
 \acknowledgements
+Please add any acknowledgements here.
 
@@
 
-\begin{figure}\label{Final_Q}
-\plotone{figures/qcompareres.eps}
+\begin{figure}
+%\plotone{figures/qcompareres.eps}
+\plotone{figures/qcompareres.pdf}
 \caption{Toomre Q distribution versus radius for each of the four resolutions at t $\sim$ 18 ORPs}
+\label{fig:Final_Q}
 \end{figure}
 
-\begin{figure}\label{fig:DensityPlots}
+\begin{figure}
 %\plotone{figures/Figure_1.eps}
+\plotone{figures/Figure_1.pdf}
 \caption
@@
 }
+\label{fig:DensityPlots}
 \end{figure}
@@
 
-\begin{figure}\label{fig:Am_vs_log_m}
+\begin{figure}
 %\plotone{figures/Figure_3.eps}
+\plotone{figures/Figure_3.pdf}
 \caption
@@
 }
+\label{fig:Am_vs_log_m}
 \end{figure}
@@
 %\plotone{figures/Power_Extrapolation.eps}
+\plotone{figures/Power_Extrapolation.pdf}
 \caption
@@
 Also plotted is the Richardson Extrapolation for the asymptotic value  
-$\langle A_{2-7} \rangle / \langle A_{tot} \rangle = 0.43$ as the grid spacing goes to zero.  
+$\langle A_{2-7} \rangle / \langle A_{tot} \rangle = 0.42$ as the grid spacing goes to zero.  
 }
+\label{fig:Power_Extrap}
 \end{figure}
@@
 %\plotone{figures/Torque_Array.eps}
+\plotone{figures/Torque_Array.pdf}
 \caption
 {
-Torque Array}
+Time-averaged torques summed over a number of low-order modes, 
+$\sum_1^n \langle {\rm \bf C}_{Z(n)} \rangle$, where $n = $1,  2, 3, 4, and 5, along
+with the total time-averaged gravitational torque,
+$\langle {\rm \bf C}_{Ztot} \rangle = \sum_1^n \langle {\rm \bf C}_{Z(n)} \rangle$, where $n =  l_{max}/2$,
+as a function of radius and azimuthal resolution. 
+The scale of the $y$-axis in each subplot is the same, with the exception of the $l_{max} = 64$ simulation, 
+where the scale is doubled.  
+Note that the total gravitational torque is smaller for simulations run at higher azimuthal resolutions, and the total torque is dominated by low-order modes for all azimuthal resolutions. 
+}
+\label{fig:Torque_array}
 \end{figure}
@@
 \begin{figure}
+%\plotone{figures/Figure_2.eps}
+\plotone{figures/Figure_2.pdf}
+\caption
+{Effective $\alpha$'s, time-averaged between 12 and 18 ORPs, for simulations performed with four different azimuthal grid resolutions.
+Dashed lines show values of $\alpha$ predicted by  \citet{gammie2001}  for the
+strongly self-gravitating limit (lower line) and the non-self-gravitating limit
+(upper line).
+The dotted line at $\log(\alpha) = -1.22$ represents the maximum stress a disk subject to constant cooling
+can maintain without fragmenting, as  proposed by  \cite{rice2005}.
+}
 \label{fig:alpha_v_radius}
-%\plotone{figures/Figure_2.eps}
-\caption
-{Effective $\alpha$'s for the $t_{cool} =$ 2 ORP, $\gamma = 5/3$, $\Sigma \sim r^{ -1}$ disk, averaged between 12 and 18 ORPs, for four different different azimuthal grid resolutions, given by the number of azimuthal grid-points Lmax.
-The parallel dashed lines show the predictions of Gammie (2001) for $\alpha$ (upper: strongly self-gravitating limit, lower: non-self-gravitating limit), while the dashed line at $\log(\alpha) = -1.22$ represents the maximum stress a disk can maintain without fragmenting (Rice et al.
-2005) .
-}
 \end{figure}
@@
 %\plotone{figures/Alpha_variations_stack512.eps}
+\plotone{figures/Alpha_variations_stack512.pdf}
 \caption
-{Alpha variations.}
+{
+Time variability of $\alpha$ as a function of radius.
+Here $\alpha$ is normalized to its averaged value between at $t = 14$ and 18 ORP.
+The orbital period at each radius is provided by the length of the bracketed line shown above its radius label.
+}
+\label{fig:alpha_stack}
 \end{figure}

### 2010-12-08 Paper_1.tex
Tom's changes w/ figures working.
	modified:   Paper_1.tex

@@
 
-\begin{figure}
-%\plotone{figures/qcompareres.eps}
-\plotone{figures/qcompareres.pdf}
+\begin{figure}\label{Final_Q}
+\plotone{figures/Figure1.eps}
 \caption{Toomre Q distribution versus radius for each of the four resolutions at t $\sim$ 18 ORPs}
-\label{fig:Final_Q}
 \end{figure}
 
-\begin{figure}
-%\plotone{figures/Figure_1.eps}
-\plotone{figures/Figure_1.pdf}
+\begin{figure}\label{fig:DensityPlots}
+%\plotone{figures/Figure2.eps}
 \caption
@@
 }
-\label{fig:DensityPlots}
 \end{figure}
@@
 
-\begin{figure}
-%\plotone{figures/Figure_3.eps}
-\plotone{figures/Figure_3.pdf}
+\begin{figure}\label{fig:Am_vs_log_m}
+%\plotone{figures/Figure3.eps}
 \caption
@@
 }
-\label{fig:Am_vs_log_m}
 \end{figure}
@@
 \begin{figure}
-%\plotone{figures/Power_Extrapolation.eps}
-\plotone{figures/Power_Extrapolation.pdf}
+%\plotone{figures/Figure4.eps}
 \caption
@@
 Also plotted is the Richardson Extrapolation for the asymptotic value  
-$\langle A_{2-7} \rangle / \langle A_{tot} \rangle = 0.42$ as the grid spacing goes to zero.  
+$\langle A_{2-7} \rangle / \langle A_{tot} \rangle = 0.43$ as the grid spacing goes to zero.  
 }
-\label{fig:Power_Extrap}
 \end{figure}
@@
 \begin{figure}
-%\plotone{figures/Torque_Array.eps}
-\plotone{figures/Torque_Array.pdf}
+%\plotone{figures/Figure5.eps}
 \caption
 {
-Time-averaged torques summed over a number of low-order modes, 
-$\sum_1^n \langle {\rm \bf C}_{Z(n)} \rangle$, where $n = $1,  2, 3, 4, and 5, along
-with the total time-averaged gravitational torque,
-$\langle {\rm \bf C}_{Ztot} \rangle = \sum_1^n \langle {\rm \bf C}_{Z(n)} \rangle$, where $n =  l_{max}/2$,
-as a function of radius and azimuthal resolution. 
-The scale of the $y$-axis in each subplot is the same, with the exception of the $l_{max} = 64$ simulation, 
-where the scale is doubled.  
-Note that the total gravitational torque is smaller for simulations run at higher azimuthal resolutions, and the total torque is dominated by low-order modes for all azimuthal resolutions. 
-}
-\label{fig:Torque_array}
+Torque Array}
 \end{figure}
@@
 \begin{figure}
-%\plotone{figures/Figure_2.eps}
-\plotone{figures/Figure_2.pdf}
+\label{fig:alpha_v_radius}
+%\plotone{figures/Figure6.eps}
 \caption
-{Effective $\alpha$'s, time-averaged between 12 and 18 ORPs, for simulations performed with four different azimuthal grid resolutions.
-Dashed lines show values of $\alpha$ predicted by  \citet{gammie2001}  for the
-strongly self-gravitating limit (lower line) and the non-self-gravitating limit
-(upper line).
-The dotted line at $\log(\alpha) = -1.22$ represents the maximum stress a disk subject to constant cooling
-can maintain without fragmenting, as  proposed by  \cite{rice2005}.
+{Effective $\alpha$'s for the $t_{cool} =$ 2 ORP, $\gamma = 5/3$, $\Sigma \sim r^{ -1}$ disk, averaged between 12 and 18 ORPs, for four different different azimuthal grid resolutions, given by the number of azimuthal grid-points Lmax.
+The parallel dashed lines show the predictions of Gammie (2001) for $\alpha$ (upper: strongly self-gravitating limit, lower: non-self-gravitating limit), while the dashed line at $\log(\alpha) = -1.22$ represents the maximum stress a disk can maintain without fragmenting (Rice et al.
+2005) .
 }
-\label{fig:alpha_v_radius}
 \end{figure}
@@
 \begin{figure}
-%\plotone{figures/Alpha_variations_stack512.eps}
-\plotone{figures/Alpha_variations_stack512.pdf}
+%\plotone{figures/Figure7.eps}
 \caption
-{
-Time variability of $\alpha$ as a function of radius.
-Here $\alpha$ is normalized to its averaged value between at $t = 14$ and 18 ORP.
-The orbital period at each radius is provided by the length of the bracketed line shown above its radius label.
-}
-\label{fig:alpha_stack}
+{Alpha variations.}
 \end{figure}

### 2010-12-14 Paper_1.tex
Edited through end of section 2.
	modified:   Paper_1.tex
	modified:   general.bib

@@
-\documentclass[manuscript]{aastex} % one-column, double-spaced document
-%\documentclass[preprint2]{aastex}  %% double-column, single-spaced document
-% \documentclass[12pt,preprint]{aastex}
-% \documentclass[aaspp4]{aastex}
-% \documentclass[aas2pp4]{aastex}
-% \documentclass[aj_pt4]{aastex}
-% \documentclass[apjpt4]{aastex}
-% \documentstyle[12pt,aasms4]{article}
-% \documentclass[12pt]{article}
-%\documentstyle[12pt]{article}
-% \tighten
-%
-%
+\documentclass[manuscript]{aastex}
+
 \newcommand{\ie}{{\it i.e.}}
@@
 
-We conducted a 3D hydrodynamic convergence study of a protostellar disk
-subject to a constant global cooling time 
-and to gravitational instabilities (GIs). 
-The disk was followed during its asymptotic state, where
-heating and cooling are in balance.
+We conducted a convergence study of the azimuthal resolution in 3D hydrodynamic simulations of a protostellar disk susceptible to gravitational instabilities (GIs). 
+The disk was primarily studied during its asymptotic state, when
+heating and cooling are in balance. The cooling was treated as a constant global cooling time.
 Four simulations were conducted, identical except for the number of azimuthal grid 
@@
 The strengths of the azimuthal modes, gravitational torques, and the effective \citet{shakura1973} $\alpha$ 
-arising from gravitational stresses were determined for each resolution.  These
-were used to examine the resolution-dependence of each quantity
-and to evaluate both the converged modal power distribution and the effective value of $\alpha$
-as $l_{max} \rightarrow \infty$.   Power was found in all modes and 
-the power summed over all modes was essentially independent of resolution.
-The modal distribution of power, however, was significantly affected by resolution.  
+arising from gravitational stresses were determined for each resolution.  We examined the resolution-dependence of each quantity
+and evaluated both the converged modal amplitude distribution and the effective value of $\alpha$
+as $l_{max} \rightarrow \infty$.   Non-axisymmetric amplitude was found in all modes and 
+the amplitude summed over all modes was essentially independent of resolution.
+The modal distribution of amplitudes, however, was significantly affected by resolution.  
 The size of the azimuthal grid, $l_{max}$,
-limits the number of modes to $l_{max}/2$.  Higher resolution simulations therefore
-have more modes available to them.  Power is spread across all available modes.  As a consequence,
-higher resolution simulations have a larger fraction of their total power in high-order modes.
+limits the number of modes that can be accurately measured to $l_{max}/2$.  Higher resolution simulations therefore
+have more modes available to them.  Non-axisymmetric structure is present in all available modes.  As a consequence,
+higher resolution simulations have a larger fraction of their total amplitude in high-order modes.
 High-order modes
-manifest themselves more locally than low-order modes, therefore 
-greater resolution leads to smaller gravitational stresses.
+are more confined to act on local structure than low-order modes, therefore as the resolution increases and low-order amplitudes decrease the total gravitational stresses decreases as well.
 As a result, higher resolution simulations experience 
@@
 
-%\keywords{ }
-
-%%%%%% INTRODUCTION %%%%%%%%%%%%%%%
 \section{Introduction}
 
-Detailed three-dimensional hydrodynamic modeling of protostellar disks can provide considerable insight into the physical and thermal states of these disks.
+Detailed three-dimensional hydrodynamic modeling of protostellar disks has provided considerable insight into the physical and thermal states of these disks.
 However,  computing resources limit these studies to snapshots spanning timescales much shorter than a disk's evolutionary lifetime.
@@
 Here convergence studies are important because
-higher order modes dissipate energy on shorter wavelengths than lower order modes.  The mix of low and high order modes available to a disk is set by the resolution of the simulation.
+higher order modes dissipate energy on shorter wavelengths than lower order modes.
+The mix of low and high order modes available to a disk is set by the resolution of the simulation.
+{\bf Multiple issues with this paragraph.
+Didn't want to rewrite unilaterally.
+1) In the first couple sentences the description of mode growth needs some tweaking, many modes appear in the linear regime, and although we have seen amplitude at all m values I don't know that it's a given.
+2) I wouldn't describe the structures formed by GI activity as ``strongly nonlinear turbulence''.
+3) In reviewing Gammie 2001 it seems that he is actually referring to a turbulent cascade (a fact I hadn't fully appreciated until now) I think it's pretty clear that the effective alpha we see is, in fact, not due to a turbulent cascade.
+i.e.
+alpha is driven by low order coherent modes.
+We should discuss this at some point}
 
@@
 
-Several studies support the idea that GI transport in real disks is, in many important respects, an intrinsically global phenomenon and cannot be properly treated by a local $\alpha$-like prescription \citep{laughlin1996, balbus1999, lodato2005, mejia2005, boley2006, cai2008}.
+Several studies support the idea that angular momentum transport by GIs in real disks is, in many important respects, an intrinsically global phenomenon and cannot be properly treated by a local $\alpha$-like prescription \citep{laughlin1996, balbus1999, lodato2005, mejia2005, boley2006, cai2008}.
 This almost certainly applies to the embedded phase of both high and low-mass stars when infall from the collapsing protostellar cloud feeds mass into the disk at a rapid
@@
 
-\citet{gammie2001} derived an effective alpha for the case of a thin 2D gravitationally unstable disk where the heating due to local turbulent dissipation caused by GIs was balanced by the local cooling rate.
-Assuming a razor thin gravitationally unstable disk where the heating due to local turbulent dissipation caused by
-GIs was balanced by a local cooling rate of $t_{cool}$.
+\citet{gammie2001} derived an effective alpha for the case of a thin 2D gravitationally unstable disk where the heating caused by GIs was balanced by the local cooling rate.  
+Assuming a razor thin gravitationally unstable disk where the heating caused by
+GIs was balanced by a local cooling rate of $t_{cool}\Omega$.
 Gammie found
@@
 $\gamma = 3 - 2/\Gamma $
-for a strongly self gravitating disk.   \cite{lodato2004} derived essentially the same result for their 3D study.
+for a strongly self gravitating disk.   \cite{lodato2004} derived a similar result when considering three dimensional protoplanetary disks.
 Motivated by
-the \cite{laughlin1996}, \cite{gammie2001}, and \cite{lodato2004, lodato2005} results 
-that demonstrated differences in how low and high order modes manifest themselves, Kratter et al. (2008) 
-proposed $\alpha_{eff} = (\alpha_{short}^2 + \alpha_{long}^2)^{1/2}$, where $\alpha_{short}$ and $\alpha_{long}$
-are functions of $Q$ and $M_d/M_*$ representing 
+the results from \cite{laughlin1996}, \cite{gammie2001}, and \cite{lodato2004, lodato2005}, which demonstrated differences in the manifestation of high and low-order, \citet{kratter2008} 
+proposed $\alpha_{eff} = (\alpha_{short}^2 + \alpha_{long}^2)^{1/2}$. Here $\alpha_{short}$ and $\alpha_{long}$
+are functions of $Q$ and $M_d/\left( M_d+M_*\right)$ representing 
 the different wavelength regimes expected to dominate for different values of the mass ratio and cooling rate. 
-{\bf (Scott: need to add kratter2008 to the bib file.  The reference is Kratter, K.M., Matzner, C. D., Krumholz, M.R., 
-2008. ApJ 681, 375)}
 Fully 3D simulations \citep[e.g.][]{lodato2004, cossins2009} suggest that, at least in the case of imposed cooling, $\alpha_{\rm{eff}}$ roughly converges to the value predicted by \citeauthor{gammie2001}.
@@
 grids with more elements provide finer meshes, and thus a more accurate solution. 
-However, increased accuracy is offset by temporal sacrifices.
+However, increased accuracy is offset by computational penalties.
 The number of equations solved at each timestep 
@@
 may be particularly important, because higher azimuthal resolution 
-allows GI power to spread to higher-order modes 
-that behave more locally. We test this by examining the power in various 
+allows non-axisymmetric amplitude to spread to higher-order modes 
+that behave more locally. We test this by examining the amplitude in various 
 modes to see how they are affected by the choice of grid.
 In the process, 
-we report the results of a convergence study designed to determine whether GI-active disks, chaotic by nature, achieve statistical equilibrium; if so, to what level do these disks approximate the gravitoturbulence of \citet{gammie2001} and with what distribution of $\alpha_{eff}$; and to characterize how the GI cascades from low-order to high-order modes.
+we report the results of a convergence study designed to determine whether GI-active disks, chaotic by nature, achieve statistical equilibrium If this is the case, we investigate how closely these disks approximate the gravitoturbulence of \citet{gammie2001} and examine the distribution of $\alpha_{eff}$.
 To this end, we follow the evolution of a GI-active disk, subject to constant cooling, in its asymptotic state, where cooling and heating are essentially balanced.
@@
 The initially unstable disk model passed through several phases of evolution \citep[see also][]{pickett2003, mejia2005}.
-The initial disk is unstable to the growth of axisymmetric structures.
-During this axisymmetric phase,
+The initial disk is unstable to the growth of non-axisymmetric structures.
+During the initial axisymmetric phase,
 which lasts several ORPs, the disk cools and contracts, the contraction is small radially but dramatic in the vertical direction.
@@
 This is then followed by a period of several ORPs where heating temporarily washes out some of the nonaxisymmetry in the disk.
-Finally the disk settles into a quasi-steady, long-lived asymptotic state of sustained GI-activity over a large part of the disk, with an overall balance of heating and cooling.
+Finally, following this transition phase, the disk settles into a quasi-steady, long-lived asymptotic state of sustained GI-activity over a large part of the disk, with an overall balance of heating and cooling.
 
 Simulations were run with four different angular resolutions, with grid sizes in $(r,\phi,z)$ of  $(j_{max}, k_{max}, l_{max}) =$ (512,64,64), (512,64,128), (512,64,256), and (512,64,512).
-Because of the large computational overhead associated with starting each simulation from $t = 0$, the $l_{max} =$ 128 simulation was run through the axisymmetric, burst and transition phases.
+Because of the computational overhead associated with running each simulation starting from $t = 0$, the $l_{max} =$ 128 simulation was run through the axisymmetric, burst and transition phases.
 Simulations using $l_{max} =$ 64, 256, and 512 all begin by interpolating the $l_{max} =$ 128 simulation to higher or lower azimuthal resolution at 9.6 ORPs, near the end of the transition
 phase.
-Although this strategy saves a large amount of computational time, it limits the scope of the analysis to the asymptotic phase and it introduces a perturbation to the disk which takes several ORPs to subside.
-By 12 ORPs, the perturbation has dissipated and all four disks have transitioned to the asymptotic state.
+Although this strategy saves a large amount of computational time, it limits the scope of the analysis to the asymptotic phase.
+By 12 ORPs, all four disks have transitioned to the asymptotic state.
 All simulations were
@@
 
-Figure 
-%\ref{qcompareres}
-\ref{fig:Final_Q} 
-shows $Q$, as a function of radius, at $t \sim$ 18 ORP.
+Figure \ref{fig:Final_Q} shows $Q$, as a function of radius, at $t \sim$ 18 ORP.
 Table \ref{tbl:ams} lists $Q_{avg}$, the value of $Q$ of for the asymptotic disk, time-averaged of 12 -- 19 ORP and spatially averaged over 10 -- 40 AU, as a function of radius for each resolution.
@@
 
-% \begin{deluxetable}{cccc} 
-% \tablecolumns{4} 
-% \tablewidth{0pc} 
-% \tablecaption{Effective Alphas} 
-% \tablehead{ 
-% \colhead{}    &  \multicolumn{3}{c}{Non-shell Stars} &   \colhead{}   & 
-%\multicolumn{3}{c}{Shell Stars} \\ 
-%\cline{2-4} \cline{6-8} \\ 
-%\cline{2-8} \\ 
-% \colhead{Azimuthal Zones} & \colhead{$\alpha$}   & \colhead{$A_{tot}$}    
-% & \colhead{$A_{2-5} / A_{tot}$} \\ 
-%\colhead{} & \colhead{} &\colhead{} &  \colhead{angle} & \colhead{$10^{-28}$ ergs /}  \\
-%\colhead{} & \colhead{} &\colhead{} &  \colhead{angle} 
-%&  \multicolumn{2}{c}{$(10^{-28}$ ergs / cm$^3$ s sr)}  \\
-% } 
-% \startdata
-% 64	 &    0.0587	 &        2.02 	 & 0.752	      \\
-% 128	   &    0.0402	  &       1.99  	 & 0.608      \\
-% 256	  &     0.0237	 &        1.87	 & 0.535     \\
-% 512	  &     0.0241	 &        2.28	 & 0.467     \\
-% \enddata 
-% \tablecomments{
-% Table 1. The $\alpha$ of figure 1, averaged over the radial region from 10 -- 40 AU, tabulated for different 
-% azimuthal resolutions, along with amplitudes of Fourier components for non-axisymmetric density structures. 
-% The radially averaged $\alpha$ predicted by Gammie (2001) are tablulated for comparison.
-% }
-% \end{deluxetable}
-
 \begin{deluxetable}{cccccc}
@@
 
-\begin{figure}\label{Final_Q}
+\begin{figure}
 \plotone{figures/Figure1.eps}
 \caption{Toomre Q distribution versus radius for each of the four resolutions at t $\sim$ 18 ORPs}
+\label{fig:Final_Q}
 \end{figure}
 
-\begin{figure}\label{fig:DensityPlots}
-%\plotone{figures/Figure2.eps}
+\begin{figure}
+\plotone{figures/Figure2.eps}
 \caption
@@
 
-\begin{figure}\label{fig:Am_vs_log_m}
-%\plotone{figures/Figure3.eps}
+\begin{figure}
+\plotone{figures/Figure3.eps}
 \caption
@@
 \begin{figure}
-%\plotone{figures/Figure4.eps}
+\plotone{figures/Figure4.eps}
 \caption
@@
 \begin{figure}
-%\plotone{figures/Figure5.eps}
+\plotone{figures/Figure5.eps}
 \caption
@@
 
-%\begin{figure}
-%\plotone{figures/Mode_CascadeRatios.pdf}
-%\caption
-%{
-%Mode Cascade Ratios .  Note that by the end of the time frame shown, there is rough parity between 
-%torques arising from tower order ($m = 2$ -- 5) mades and torques from higher order modes.}
-%\end{figure}
-
 \begin{figure}
 \label{fig:alpha_v_radius}
-%\plotone{figures/Figure6.eps}
+\plotone{figures/Figure6.eps}
 \caption
@@
 \begin{figure}
-%\plotone{figures/Figure7.eps}
+\plotone{figures/Figure7.eps}
 \caption

### 2010-12-22 Paper_1.tex
Scott's first round of revisions.
	modified:   Paper_1.tex
	modified:   general.bib

@@
 also depends upon the magnitude of the stresses.  Thus
-$\alpha_{eff}$ is also a function of grid resolution, decreasing as  $l_{max}$ increases.
-The converged $\alpha_{eff}$ is consistent with
+$\alpha_{\rm{eff}}$ is also a function of grid resolution, decreasing as  $l_{max}$ increases.
+The converged $\alpha_{\rm{eff}}$ is consistent with
 predictions from an analytic local theory for thin disks by \citet{gammie2001}, 
@@
 alpha is driven by low order coherent modes.
-We should discuss this at some point}
+We should discuss this at some point SM}
 
@@
 the disk scale height. The $\alpha$ formalism makes no assumption about the nature of the angular momentum transport, only that {\it angular momentum transport
-and heating are dominated by local processes}. Lin and Pringle (1987, 1990) were among the first to suggest
+and heating are dominated by local processes}. \citet{lin1987,lin1990} were among the first to suggest
 that GI-induced transport could be described within a viscous
@@
 an $\alpha$-viscosity may be consistent
-with transport in GI acitve disks for disk to star mass ratio $M_d/M_* \le 0.2$ -- 0.3; GI-active disks
+with transport in GI active disks for disk to star mass ratio $M_d/M_* \le 0.2$ -- 0.3; GI-active disks
 in systems with larger $M_d/M_*$ are dominated by lower-order modes that act in a more global sense and thus cannot be well represented by an $\alpha$-approach
@@
 Gammie found
-$\alpha_{eff} = [\gamma(\gamma - 1) \slantfrac{9}{4}\Omega t_{cool}]^{-1}$,
+\begin{equation}
+\alpha_{\rm{eff}} = [\gamma(\gamma - 1) \slantfrac{9}{4}\Omega t_{cool}]^{-1},
+\label{eq:gammie}
+\end{equation}
 where $\gamma$ is the two dimensional adiabatic index and $t_{cool}$ is a cooling parameter. 
@@
 the results from \cite{laughlin1996}, \cite{gammie2001}, and \cite{lodato2004, lodato2005}, which demonstrated differences in the manifestation of high and low-order, \citet{kratter2008} 
-proposed $\alpha_{eff} = (\alpha_{short}^2 + \alpha_{long}^2)^{1/2}$. Here $\alpha_{short}$ and $\alpha_{long}$
+proposed $\alpha_{\rm{eff}} = (\alpha_{short}^2 + \alpha_{long}^2)^{1/2}$. Here $\alpha_{short}$ and $\alpha_{long}$
 are functions of $Q$ and $M_d/\left( M_d+M_*\right)$ representing 
 the different wavelength regimes expected to dominate for different values of the mass ratio and cooling rate. 
-Fully 3D simulations \citep[e.g.][]{lodato2004, cossins2009} suggest that, at least in the case of imposed cooling, $\alpha_{\rm{eff}}$ roughly converges to the value predicted by \citeauthor{gammie2001}.
+Fully 3D simulations \citep[e.g.][]{lodato2004, cossins2009} suggest that, at least in the case of imposed cooling, $\alpha_{\rm{\rm{eff}}}$ roughly converges to the value predicted by \citeauthor{gammie2001}.
 
@@
 However, increased accuracy is offset by computational penalties.
-The number of equations solved at each timestep 
+The number of equations solved at each time step 
 increases as the grid size is increased.  At the same time, the Courant condition 
-forces shorter timesteps.  A mesh convergence study 
+forces shorter time steps.  A mesh convergence study 
 provides a common approach to balancing accuracy and computing resources.
@@
 In the process, 
-we report the results of a convergence study designed to determine whether GI-active disks, chaotic by nature, achieve statistical equilibrium If this is the case, we investigate how closely these disks approximate the gravitoturbulence of \citet{gammie2001} and examine the distribution of $\alpha_{eff}$.
+we report the results of a convergence study designed to determine whether GI-active disks, chaotic by nature, achieve statistical equilibrium If this is the case, we investigate how closely these disks approximate the gravitoturbulence of \citet{gammie2001} and examine the distribution of $\alpha_{\rm{eff}}$.
 To this end, we follow the evolution of a GI-active disk, subject to constant cooling, in its asymptotic state, where cooling and heating are essentially balanced.
@@
 The disk expands violently during the burst phase, producing a significant rearrangement of the disk's mass distribution on a timescale of a few ORPs.
-This is then followed by a period of several ORPs where heating temporarily washes out some of the nonaxisymmetry in the disk.
+This is then followed by a period of several ORPs where heating temporarily washes out some of the non-axisymmetry in the disk.
 Finally, following this transition phase, the disk settles into a quasi-steady, long-lived asymptotic state of sustained GI-activity over a large part of the disk, with an overall balance of heating and cooling.
@@
 
-Figure \ref{fig:Final_Q} shows $Q$, as a function of radius, at $t \sim$ 18 ORP.
-Table \ref{tbl:ams} lists $Q_{avg}$, the value of $Q$ of for the asymptotic disk, time-averaged of 12 -- 19 ORP and spatially averaged over 10 -- 40 AU, as a function of radius for each resolution.
-From $\sim$ 12 -- 50 AU, $Q \sim 1.0$ -- 1.4 and thus the disk is subject to the growth of nonaxisymmetric structures over this full range.
+Figure \ref{fig:Final_Q} shows the Toomre $Q$, as a function of radius, for all four simulations at $t \sim$ 18 ORP.
+Table \ref{tbl:ams} lists $Q_{avg}$, the value of $Q$ of each of the asymptotic disks, time-averaged from 12 -- 19 ORP and spatially averaged over 10 -- 40 AU.
+From $\sim$ 12 -- 50 AU, $Q \sim 1.0$ -- 1.4 and thus the disk is subject GIs over this full range.
 Interior to $\sim$ 27 AU, the value of $Q$ shows no dependence on
-the azimuthal resolution.  From 27 -- 50 AU, lower resolution simulations display lower values of $Q$.
+the azimuthal resolution.
+From 27 -- 50 AU, lower resolution simulations display lower values of $Q$.
+{\bf I suppose 27 AU was chosen by eye?
+Or was there some other analysis.
+I vote to toss these last two sentences as I'm not convinced w/o a time averaged quantity.
+Unless this observation is used in further reasoning in the discussion...SM}
 
@@
 In the asymptotic state, GI-active disks develop complex density structures arising from the superposition of
-all modes present in the disk.
-Since the number of modes accessible to the disk is a function of the azimuthal resolution, we expect that higher resolution simulations will manifest more power in higher-order modes
-that behave more locally. We test this by examining the power in various modes to see how they are affected by the choice of grid.
+modes in the disk.
+Since the number of modes accessible to the disk depends on the azimuthal resolution, we expect that higher resolution simulations will exhibit non-axisymmetric amplitude in higher-order modes inaccessible at lower resolution, and these higher-order modes will
+behave more locally. We test this by examining the amplitude in various modes to see how they are affected by the choice of grid.
 
@@
 
-Figure \ref{fig:DensityPlots}, shows midplane and radial densities  at $t \sim 18$ ORPs for the $l_{max} =$ 64, 128, 256, and 512 simulations.
-Two features are readily apparent; disks run at higher resolution are thicker than lower resolution disks {\bf (anyone want to comment on this?)} and the character of density perturbations differ with resolution.
+Figure \ref{fig:DensityPlots}, shows mid-plane and radial densities  at $t \sim 18$ ORPs for the $l_{max} =$ 64, 128, 256, and 512 simulations.
+It is readily apparent that the character of the non-axisymmetric density structures differ with
+resolution.
+One can clearly see that the lower resolution simulations, i.e.
+$l_{max} = 64$ and 128, have primarily low-order structures with $m = 2$ and 3 dominating.
+In contrast, the higher resolution simulations have much more fine structure, the modes are much more washed out, and the low-order modes no longer dominate.
 The lower resolution (smaller $l_{max}$) simulations display longer wavelength spiral structures, with larger amplitudes and more coherence than those seen in higher resolution simulations.
-Higher angular resolution allows GI power to spread to higher-order azimuthal modes, as demonstrated by a Fourier decomposition of the density.
+Higher angular resolution allows non-axisymmetric structures to grow in higher-order azimuthal modes, as demonstrated by a Fourier decomposition of the density.
 The strengths of these modes are given by their global Fourier amplitudes,
@@
 Table \ref{tbl:ams} lists $\langle A_{tot} \rangle$ along with $\langle A_{2-7} \rangle  /  \langle A_{tot} \rangle$,  the fraction of the total power falling into low-order ($m = 2$ -- 7) modes, for each resolution.
+We identify $m=2-7$ as the low-order modes that are considered to be global modes, again
+$m=1$ is excluded due to the fixed central star. This determination was made by considering the radial range over 
+which the mode is most effective in transporting angular momentum, i.e. from the inner Linblad resonance to the outer 
+Linblad resonance, and comparing it to the disk scale height. These quantities are roughly equal for $m=7$, for lower 
+order modes the radial range exceeds the scale height.
 While the total power is approximately independent of resolution, $\langle A_{tot} \rangle = 1.90 \pm 0.07$, 
@@
 The shift of power from lower to higher order modes as the resolution of the computational grid is increased arises from the fact that, while $\langle A_{tot} \rangle$ is approximately independent of $l_{max}$,  the number of degrees of freedom available for this power, $l_{max}/2$, increases linearly with the resolution.
-Thus power is naturally spread from lower order to higher order modes as the available $m$-values increases.
+Thus amplitude is naturally spread from lower order to higher order modes as the available $m$-values increase.
 Indeed, even if the disk is dominated by one or two low-order modes at the time of outburst, this power cascades to higher-order modes \citep[for example][]{laughlin1997, laughlin1998,laughlin1996}.
+{\bf Haven't had a chance to check these refs but I'm reluctant to call what we see a cascade of power.
+I suggest striking this sentence.
+SM}
 
-Given the resolution dependence of how power is distributed,  we must examine what resolution is sufficient to properly capture the evolution of these disks.
+Given how the non-axisymmetric amplitude distribution depends on the azimuthal resolution, we must examine what resolution is sufficient to properly capture the evolution of these disks.
 Table \ref{tbl:ams} shows that while the fractional power in lower order modes decreases with each doubling of $l_{max}$, the amount of change in going from $l_{max} = 256$ to 512
@@
 Indeed, a  Richardson extrapolation \citep{press1992} of the tabulated values of  $\langle A_{2-7} \rangle / \langle A_{tot} \rangle$ suggests a limiting value of  $\langle A_{2-7} \rangle / \langle A_{tot} \rangle \sim$ 0.42 for increasingly higher resolutions.
-This is displayed graphically in Figure \ref{fig:Power_Extrap}, where the fractional power in lower order modes is plotted against the  normalized grid
+This is displayed graphically in figure \ref{fig:Power_Extrap}, where the fractional power in lower order modes is plotted against the  normalized grid
 spacing for each value of $l_{max}$. The grid spacing  is normalized such that spacing of the $l_{max}$ = 512 grid
@@
 The Richardson extrapolation value  of $\langle A_{2-7}\rangle  / \langle A_{tot}\rangle  = 0.42$ is plotted at a normalized grid spacing  equal to zero.
-The majority of the modal power in a converged disk resides in higher-order modes.
+%The majority of the modal power in a converged disk resides in higher-order modes. 
 %These modes dissipate on shorter wavelengths than low-order modes and hence act more locally.
@@
 arms to produce torques, while high-order modes have relatively short wavelengths which, particularly if they 
-lack coherence, produce more localized effects. High-order modes tend to produce pockets of local GI activity, 
-which can cancel each other over the whole disk.  The finding that the fractional power in high-order modes increases
+lack coherence can cancel each other out, and produce more localized effects. 
+The finding that the fractional power in high-order modes increases
 with resolution, while the total power stays the same, leads to
- the expectation that  gravitational torques in the disk will decrease as $l_{max}$ is increased. This can be tested by examining the torque contributions from  various Fourier modes.
+ the expectation that the gravitational torques in the disk will decrease as $l_{max}$ is increased. This can be tested by examining the torque contributions from  various Fourier modes.
 
@@
 as only this component drives mass and angular momentum transfer.
-The torque contribution from each mode $m$ can be calculated by replacing $\rho$ in Equation (5) with the density distribution reconstructed from a single Fourier component
+The torque contribution from each mode $m$ can be calculated by replacing $\rho$ in equation \eqref{torque1} with the density distribution reconstructed from a single Fourier component
 \begin{equation}
@@
 
-Figure 5  displays 
+Figure \ref{fig:torquearray} displays 
  time-averaged torques summed over a number of low-order modes, $\sum_1^n \langle {\rm \bf C}_{Z(n)} \rangle$, $n = $ 1, 2, 3, 4, and 5, and $\langle {\rm \bf C}_{Ztot} \rangle$,
@@
 should not be considered realistic.  
-Michael \& Durisen (2010) found that freeing the star makes only 10-20\%
-differences,  thus the difference between torques arising from the $m=1$ mode with the star fixed are roughly half that found when the star is free to move {\bf (Scott or Dick: please rework this sentence)}.
+\citet{michael2010} found that freeing the star changes the torques by 10-20\% at most, 
+thus the difference between torques arising from the $m=1$ mode with the star fixed and free to move are small compared to the differences found when comparing azimuthal resolution.
 
-We note two results of Figure 5: 1) the total gravitational torque is smaller for higher resolution simulations, and 2) the total torque is dominated by low-order modes for all azimuthal resolutions.
+We note two results of figure \ref{fig:torquearray}: 1) the total gravitational torque is smaller for higher resolution simulations, and 2) the total torque is dominated by low-order modes for all azimuthal resolutions.
 Both reflect the fact that while a higher fraction of the total power in larger $l_{max}$ simulations is in higher order modes, these modes make only minimal contributions to the gravitational torque.
-As resolution increases, non-axisymmetric power is shifted from low-order global modes, which produce relatively large gravitational torques, to high-order local modes, which tend to cancel each other when integrated over the whole disk.
+As resolution increases, non-axisymmetric amplitude is shifted from low-order global modes, which produce relatively large gravitational torques, to high-order local modes, which tend to cancel each other when integrated over the whole disk.
 
@@
 
+{\bf I still think the scale in figure 5 is off by a factor of 2 probably due to the fact that you've used a 0.5 $M_\odot$ star. I also vote for a single 12-18 ORP panel. SM}
+
 \subsection{Effective Alphas}
 
-An $\alpha$-framework treats disks as essentially two-dimensional objects subject to local energy dissipation.
+%An $\alpha$-framework treats disks as essentially two-dimensional objects subject to local energy dissipation.
 Therefore, in order to compare our results with predicted and published results, we must extract two-dimensional information from our three-dimensional disks.
+
+{\bf I've made some significant changes to this text from Tom's original. Specifically in the first, fourth, and fifth paragraphs. Please look over and make sure this fits in properly. SM}
+
+After obtaining the gravitational stresses in the disk, one can compute an effective $\alpha$. 
 Following \cite{gammie2001} and \citet{lodato2004}, 
@@
 \alpha_{\rm{eff}}(\varpi) = \left| \frac{{\rm d} \ln \Omega}{{\rm d} \ln \varpi} \right|^{-1} 
-\frac {\langle T_{\varpi\phi}^{grav}\rangle + \langle T_{\varpi\phi}^{Reyn}\rangle}
+\frac {T_{\varpi\phi}^{grav} + T_{\varpi\phi}^{Reyn}}
 {\Sigma c_s^2 },
 \end{equation}
-where $\Omega$ is the azimuthally averaged rotation speed, $T_{\varpi\phi}^{grav}$ and $T_{\varpi\phi}^{Reyn}$ represent the azimuthally averaged $\varpi,\phi$-component of the specific (per unit mass) gravitational
-(or ``Newton'') and hydrodynamic (or ``Reynolds'') stress tensors, respectively, $\Sigma$ the surface density,
-and $c_s$ the azimuthally averaged midplane sound speed.  Brackets indicate quantities integrated over the $z$-direction. 
-{\bf (All: In the previous section, brackets were used to denote time averages (torques),  while here brackets are used to denote spatial
-averages.  I welcome suggestions for recasting the notation in either case)}
-The denominator $\Sigma c_c^2$ was evaluated as $\int_z \rho c_c^2 {\rm d}z$. {\bf SM: I need to rewrite this entire paragraph}
+where $\Omega$ is the azimuthally averaged rotation speed, $T_{\varpi\phi}^{grav}$ and $T_{\varpi\phi}^{Reyn}$ represent the  gravitational and Reynolds stress tensors, respectively, $\Sigma$ the surface density,
+and $c_s$ the sound speed.
+In comparison to \citeauthor{gammie2001}'s 2D models and \citeauthor{lodato2004}'s highly flattened smoothed particle hydrodynamics models, the models presented here have a significant vertical extent.
+Because of this, the sound speed can vary dramatically from the mid-plane to the low density regions, so it was unclear as to the best way to determine a representative sound speed.
+In order to include contributions to the sound speed from the entire vertical extent, the denominator $\Sigma c_s^2$ was evaluated as $\int_z \rho c_s^2 {\rm d}z$.
 
@@
 previous studies have found Reynolds stresses to be small relative to gravitational stresses -- see, 
-for example, Figure 5 of \cite{lodato2004} and Figure 10 of \cite{boley2006} -- suggesting that this term
-can be safely neglected in the evaluation of $\alpha_{eff}$.   For this reason, 
+for example, figure 5 of \cite{lodato2004} and figure 10 of \cite{boley2006} -- suggesting that this term
+can be safely neglected in the evaluation of $\alpha_{\rm{eff}}$.   For this reason, 
 here we omit
-$\langle T_{\varpi\phi}^{Reyn}\rangle$ from the calculation of $\alpha_{eff}$. 
+$T_{\varpi\phi}^{Reyn}$ from the calculation of $\alpha_{\rm{eff}}$. 
 We note that \cite{gammie2001}  found
-Reynolds and gravitational stresses contributed equally to $\alpha_{eff}$ in his study.  However,
+Reynolds and gravitational stresses contributed equally to $\alpha_{\rm{eff}}$ in his study.  However,
 his was a local, shearing-sheet, 2D study.  As such, it was unable to capture global effects, such
@@
 
-The Newton stresses of equation \ref{alpha1} can be evaluated using the gravitational torque of equation \ref{torque1}, i.e.,
+The Newton stresses of equation \eqref{alpha1} can be evaluated using the gravitational torque of equation \eqref{torque1}, i.e.,
 \begin{equation}
@@
 \end{equation}
-Figure 6 shows $\alpha_{eff}$, time-averaged from 12 to 18 ORPs, calculated using equations \ref{alpha1} and \ref{newtonstress},
+Figure \ref{fig:alpha_v_radius} shows $\alpha_{\rm{eff}}$, time-averaged from 12 to 18 ORPs, calculated using equations \eqref{alpha1} and \eqref{newtonstress},
 for the four grid resolutions.  For comparison, 
-$\alpha_{eff}$ predicted by \citet{gammie2001} for disks where the local balance of heating and cooling dominates the energetics.
-The lower curve of Gammie applies for the strongly self-gravitating limit while the upper curve
-is for the non-self-gravitating limit.   The $l_{max} = 512$ curve is consistent with these two limits.  However, 
+$\alpha_{\rm{eff}}$ predicted by equation \eqref{eq:gammie} from \citet{gammie2001} are plotted for disks.
+The lower  Gammie curve applies for the strongly self-gravitating limit while the upper curve
+represents the non-self-gravitating limit.   The $l_{max} = 512$ curve is consistent with these two limits.  However, 
 the  $l_{max} = 512$ curve does not represent a converged model.
-To determine whether the converged value of $\alpha_{eff}$ is consistent with predictions, we examine 
-the spatially (10-40 AU) and temporally (12-19 ORP) averaged  $\alpha_{avg}$,
-provided in Table 1, for each value of  $l_{max}$.    A Richardson extrapolation applied to these values  
+To determine whether the converged value of $\alpha_{\rm{eff}}$ is consistent with predictions, we examine 
+the spatially (10 -- 40 AU) and temporally (12 -- 19 ORP) averaged  $\alpha_{avg}$,
+provided in table \ref{tbl:ams}, for each value of  $l_{max}$.    A Richardson extrapolation applied to these values  
 converges to $\alpha_{avg} = 0.017$ (over the same radial  and time range)
-as the grid spacing goes to zero.   This falls
-between the $\alpha_{avg}$ values for the Gammie curves 0.014 (strongly self-gravitating) and 0.027 (non-self-gravitating),
-with better agreement towards the strongly self-gravitating limit.
+as the grid spacing goes to zero.   This falls squarely
+between the $\alpha_{avg}$ values for the Gammie curves 0.014 (strongly self-gravitating) and 0.027 (non-self-gravitating).
 
 At first glance, these results suggest strong agreement with Gammie and thus strong support for dissipative
-locality.  However, a cautionary note is in order.  At any given radius, 
-stresses, and hence $\alpha_{eff}$, are highly variable with time.
-Figure 7 shows the time-variability of $\alpha_{eff}$, 
+locality and GIs acting as a local mechanism overall.
+However, several cautionary notes are in order.
+It is important to remember that the bulk of the gravitational torque, and therefore the $\alpha_{\rm{eff}}$, is generated by low order modes that are coherent over tens of AU.
+As table \ref{tbl:ams} shows, even at the highest resolution, and lowest relative strength of low-order modes, the $A_{2-7}$ modes account for half of the total amplitude.
+In order to accurately measure and characterize the contribution from these modes, one must consider a significant fraction of the disk.
+Furthermore, simulations of disks with a large radial extent (e.g.
+100s of AU \citep{boley2009}) show coherent low order modes over a significant fraction of the disk.
+In order to determine the radial extent and strength of such modes one should consider the entire range of radii over which the disk is susceptible to instabilities i.e.
+$Q \lesssim 1.5$.
+
+In order to predict an $\alpha_{\rm{eff}}$, \citeauthor{gammie2001}'s formula, presented in equation \eqref{eq:gammie}, also requires knowledge of the cooling rate.
+This is generally not known {\it a priori} for a given disk.
+Moreover, the cooling rate may not be easily parametrized with a local or global prescription over the entire disk.
+Simulations with realistic radiative physics \citep{boley2006} have shown that cooling rates can vary substantially as a function of both radius and time in a GI active disk.
+In fact, it is these areas where the cooling rates change, leading to more or less GI activity, where many interesting phenomena may occur.
+
+In addition, at any given radius, 
+stresses, and hence $\alpha_{\rm{eff}}$, are highly variable with time.
+Figure \ref{fig:alphavar} shows the time-variability of $\alpha_{\rm{eff}}$, 
 normalized to its average value between at $t = 14$ and 18 ORP, at seven radial locations.   
 Also shown is the orbital period at each radius, provided by the length of the line at the end of each subplot. 
-Fluctuations in $\alpha_{eff}$
+Fluctuations in $\alpha_{\rm{eff}}$
 on the order of 100\% or more are seen at all radii
-on  time scales range from fractions to integral multiples of the dynamic (orbital) time.  The spatial averaging
-of Figure 6 washes out this local variabilty.
+on  time scales range from fractions to integral multiples of the dynamic (orbital) time.  The temporal averaging
+of figure \ref{fig:alpha_v_radius} washes out this local variability.
 Thus,
@@
 power in the high-order ($m = 2$ -- 7) modes by essentially the same amount.
+However, it should be noted that the fractional difference between successive doubling of the resolution has decreased for $l_{max} = 512$.
+From $l_{max} =64$ to 128 the fractional difference in $\alpha_{avg}$ is 36\%.
+This decreases to 26\% for the jump from $l_{max} = 256$ to 512.
+Furthermore, the converged value represents infinite azimuthal resolution, which is, of course, impossible to achieve in a fixed grid scheme.
+It is unclear as to what variation of $\alpha$ between two values of $l_{max}$ would represent convergence as $\alpha$ itself is a dynamically varying quantity, see figure \ref{fig:alphavar}.
+
+Another consideration is the effect of $r$ and $z$ resolutions on the outcome of the simulation.
+In this study we have chosen to focus on azimuthal resolution because it has the greatest effect on resolving the GI structures responsible for producing the gravitational stresses.
+However, as one increases the azimuthal resolution the radius at which the cells have roughly equal sides increases as well.
+Ideally, this should occur near the radial range where GIs are most active.
+For $l_{max} = 512$, $\delta r = \delta \phi r$ at $\approx 13.5$ AU, which is near the radius where GIs are active.
+
 Both the gravitational torques and effective alphas are functions of the
@@
 gravitational torques that are $\sim 18$\% larger than the converged torques.  For comparison, the
-errors in the $l_{max} = 256$ simulations would exceed 50\%.   For many (most?) problems of interest,
+errors in the $l_{max} = 256$ simulations would exceed 50\%. For many problems of interest,
 results obtained using $l_{max} = 512$ should provide reasonable agreement with results obtained 
 with higher resolution.
+
+It is also worth noting that the overestimation of the effective $\alpha$ is more than likely dependent on the total strength of non-axisymmetric amplitude $\langle A_{tot} \rangle$, which may be smaller depending on the physics of a simulation.
+For example, \citet{boley2006} found weaker overall GI activity when considering realistic radiative physics.
+A similar study of the effect of azimuthal resolution in disks cooled with a realistic scheme will be presented in a follow-up paper.
+
+{\bf Substantially changed this section, might still need some work. SM}
 
@@
 
-In an $\alpha$ disk formalism the mechanism giving rise to the transport of angular momentum also acts to dissipate energy.
+%In an $\alpha$ disk formalism the mechanism giving rise to the transport of angular momentum also acts to dissipate energy.
 Lodato \& Rice (2005) postulated that a disk in a state of gravitoturbulence would have locally balanced heating and cooling rates.
 If the heating is due to local dissipation, then $\alpha$ and $t_{cool}$ are related by
-\begin{equation}
-\alpha = \left| \frac{{\rm d}\ln \Omega}{{\rm d} \ln \varpi} \right|^{-2}
-\frac{1}{\gamma(\gamma - 1) t_{cool}\Omega}.
-\end{equation}
-\citet{rice2005} concluded that for a given $t_{cool}$  there exists some maximum stress, quantified by $\alpha_{max}$, which if exceeded would cause a disk to fragment.
-Their predicted value $\alpha_{max} \approx 0.06$ is shown as a horizontal dotted line in Figure 6.
+%\begin{equation}
+%\alpha = \left| \frac{{\rm d}\ln \Omega}{{\rm d} \ln \varpi} \right|^{-2}
+%\frac{1}{\gamma(\gamma - 1) t_{cool}\Omega}.
+%\end{equation}
+%\citet{rice2005} concluded that for a given $t_{cool}$  there exists some maximum stress, quantified by $\alpha_{max}$, which if exceeded would cause a disk to fragment.
+Their predicted value $\alpha_{max} \approx 0.06$ is shown as a horizontal dotted line in figure \ref{fig:alpha_v_radius}.
 Note that both the $l_{max} = $ 64 and 128 simulations exceed this limiting value over some radial range, yet neither displayed fragmentation.
@@
 \enddata 
-\tablecomments{
+\caption{
 All quantities are time-averaged over 12 -- 19 ORP.
 $Q_{avg}$ and $\alpha_{avg}$ are also spatially
-averaged over 10 - 40 AU.
+averaged over 10 -- 40 AU.
 }
@@
 \plotone{figures/Figure1.eps}
-\caption{Toomre Q distribution versus radius for each of the four resolutions at t $\sim$ 18 ORPs}
+\caption{Azimuthally averaged mid-plane Toomre Q distribution versus radius for each of the four resolutions at t $\sim$ 18 ORPs}
 \label{fig:Final_Q}
@@
 {
-Midplane disk densities at $t \sim 18$ ORPs for simulations with 64, 128, 256 and 512 azimuthal grid elements.
+Mid-plane disk densities at $t \sim 18$ ORPs for simulations with 64, 128, 256 and 512 azimuthal grid elements.
 The simulations differ only in the number of  azimuthal zones used in the calculations.
@@
 }
+\label{fig:Power_Extrap}
 \end{figure}
@@
 Torque Array}
+\label{fig:torquearray}
 \end{figure}
@@
 \caption
-{Effective $\alpha$'s for the $t_{cool} =$ 2 ORP, $\gamma = 5/3$, $\Sigma \sim r^{ -1}$ disk, averaged between 12 and 18 ORPs, for four different different azimuthal grid resolutions, given by the number of azimuthal grid-points Lmax.
-The parallel dashed lines show the predictions of Gammie (2001) for $\alpha$ (upper: strongly self-gravitating limit, lower: non-self-gravitating limit), while the dashed line at $\log(\alpha) = -1.22$ represents the maximum stress a disk can maintain without fragmenting (Rice et al.
-2005) .
+{Effective $\alpha$'s for the $t_{cool} =$ 2 ORP, $\gamma = 5/3$, $\Sigma \sim r^{ -1}$ disk, averaged between 12 and 18 ORPs, for four different different azimuthal grid resolutions, given by the number of azimuthal grid-points $l_{max}$.
+The dashed lines show the predictions of Gammie (2001) for $\alpha$ (upper: strongly self-gravitating limit, lower: non-self-gravitating limit), while the dashed line at $\log(\alpha) = -1.22$ represents the maximum stress a disk can maintain without fragmenting (Rice et al.
+2005).
+{\bf suggest removing the Rice line if it is not discussed in the paper.}
 }
@@
 {Alpha variations.}
+\label{fig:alphavar}
 \end{figure}

### 2011-01-05 Paper_1.tex
Tom's changes...
	modified:   Paper_1.tex
	modified:   figures/Figure1.eps
	modified:   figures/Figure2.eps
	modified:   figures/Figure3.eps
	modified:   figures/Figure4.eps
	modified:   figures/Figure5.eps
	modified:   figures/Figure6.eps
	modified:   figures/Figure7.eps

@@
 The mix of low and high order modes available to a disk is set by the resolution of the simulation.
-{\bf Multiple issues with this paragraph.
+{\bf (SM Comments: Multiple issues with this paragraph.
 Didn't want to rewrite unilaterally.
@@
 alpha is driven by low order coherent modes.
-We should discuss this at some point SM}
+We should discuss this at some point), (TSC thoughts: I feel comfortable with the text as written.
+As per Scott's points 1 and 2, I think the works referenced are consistent with the text.
+As for point 3, we do see a transfer of power from lower orders to higher orders as
+time progresses, just as in Gammie.
+This is a valid usage of the term cascade.
+Gammie use of the word ``cascade'' is predicated on the same type of analysis conducted here.
+I will leave the paragraph as is and let Aaron and Dick comment.)}
 
@@
 Motivated by
-the results from \cite{laughlin1996}, \cite{gammie2001}, and \cite{lodato2004, lodato2005}, which demonstrated differences in the manifestation of high and low-order, \citet{kratter2008} 
+the results of \cite{laughlin1996}, \cite{gammie2001}, and \cite{lodato2004, lodato2005}, which demonstrated differences in the manifestation of high and low-order azimuthal modes, \citet{kratter2008}
 proposed $\alpha_{\rm{eff}} = (\alpha_{short}^2 + \alpha_{long}^2)^{1/2}$. Here $\alpha_{short}$ and $\alpha_{long}$
-are functions of $Q$ and $M_d/\left( M_d+M_*\right)$ representing 
+are functions of $Q$ and the fractional mass of the disk, $M_d/\left( M_d+M_*\right)$, representing 
 the different wavelength regimes expected to dominate for different values of the mass ratio and cooling rate. 
@@
 In the following sections we follow the evolution of a disk subject to GIs
-using four different
-resolutions in the azimuthal direction.  The number of azimuthal grid points used in simulations
+using grids with four different
+resolutions in the azimuthal direction.  The number of azimuthal elements used in simulations
 may be particularly important, because higher azimuthal resolution 
-allows non-axisymmetric amplitude to spread to higher-order modes 
+allows power in non-axisymmetric density structures to spread to higher-order modes 
 that behave more locally. We test this by examining the amplitude in various 
@@
 Table \ref{tbl:ams} lists $Q_{avg}$, the value of $Q$ of each of the asymptotic disks, time-averaged from 12 -- 19 ORP and spatially averaged over 10 -- 40 AU.
-From $\sim$ 12 -- 50 AU, $Q \sim 1.0$ -- 1.4 and thus the disk is subject GIs over this full range.
-Interior to $\sim$ 27 AU, the value of $Q$ shows no dependence on
-the azimuthal resolution.
-From 27 -- 50 AU, lower resolution simulations display lower values of $Q$.
-{\bf I suppose 27 AU was chosen by eye?
-Or was there some other analysis.
-I vote to toss these last two sentences as I'm not convinced w/o a time averaged quantity.
-Unless this observation is used in further reasoning in the discussion...SM}
+From $\sim$ 12 -- 50 AU, $Q \sim 1.0$ -- 1.4 and thus the disk is subject to GIs over this full range.
 
@@
 
-{\bf (Note to all: Throughout the paper I use integrals in places where we 
+{\bf (Note from TSC: Throughout the paper I use integrals in places where we 
 formally have used summations.  I note that this is a common practice, though
@@
 Indeed, even if the disk is dominated by one or two low-order modes at the time of outburst, this power cascades to higher-order modes \citep[for example][]{laughlin1997, laughlin1998,laughlin1996}.
-{\bf Haven't had a chance to check these refs but I'm reluctant to call what we see a cascade of power.
-I suggest striking this sentence.
-SM}
+{\bf (SM note: Haven't had a chance to check these refs but I'm reluctant to call what we see a cascade of power.
+I suggest striking this sentence.), (TSC response: I have indicated my thoughts on this in the introductory section.
+Dick \& Aaron, chip in here.)}
 
@@
 Dick has told me he only believes the 12-18 ORP result - Dick, I hope 
-I'm not misquoting you. Should we go with multiple panels representing different time-frames or 
-go solely with the 12 - 18 orp panel?
+I'm not misquoting you.  I like the multiple panels because it seems to show that the disk has not yet achieved 
+its asymptotic state.  Comparing the 12-14, 14-16, and 16-18 orp panels for
+the 512 simulation shows that the total torque decreases with time.  The 12-18 orp frame is heavily weighted
+by the 12-14 orp window.  Also, the fractional power in the $m =$ 2 and 3 modes diminishes with time, while the 12-18 orp
+frame is heavily influenced by the 12-14 orp window.  Should we go with multiple panels representing different time-frames or 
+go solely with the 12 - 18 orp panel?  Scott votes for the 12-18 panel by itself, I argue for the four frames.)
 }
-
-{\bf I still think the scale in figure 5 is off by a factor of 2 probably due to the fact that you've used a 0.5 $M_\odot$ star. I also vote for a single 12-18 ORP panel. SM}
 
@@
 
-{\bf I've made some significant changes to this text from Tom's original. Specifically in the first, fourth, and fifth paragraphs. Please look over and make sure this fits in properly. SM}
+\noindent {\bf (I've made some significant changes to this text from Tom's original.
+Specifically in the first, fourth, and fifth paragraphs.
+Please look over and make sure this fits in properly.
+SM), (TSC - I am passing Scott's changes along intact.)}
 
@@
 In comparison to \citeauthor{gammie2001}'s 2D models and \citeauthor{lodato2004}'s highly flattened smoothed particle hydrodynamics models, the models presented here have a significant vertical extent.
-Because of this, the sound speed can vary dramatically from the mid-plane to the low density regions, so it was unclear as to the best way to determine a representative sound speed.
+Because of this, the sound speed can vary dramatically from the mid-plane to the low density regions, so it was unclear as to the best way to determine a representative sound speed {\bf (we should drop the last phrase/sentence - TSC)}.
 In order to include contributions to the sound speed from the entire vertical extent, the denominator $\Sigma c_s^2$ was evaluated as $\int_z \rho c_s^2 {\rm d}z$.
@@
 However, several cautionary notes are in order.
-It is important to remember that the bulk of the gravitational torque, and therefore the $\alpha_{\rm{eff}}$, is generated by low order modes that are coherent over tens of AU.
-As table \ref{tbl:ams} shows, even at the highest resolution, and lowest relative strength of low-order modes, the $A_{2-7}$ modes account for half of the total amplitude.
+It is important to remember that the bulk of the gravitational torque, and therefore the $\alpha_{\rm{eff}}$, is generated by low order modes that are coherent over tens of AU {\bf (TSC: Scott, what is the basis for saying these are coherent over this radial range?)}.
+As table \ref{tbl:ams} shows {\bf (there is a problem with the table label and reference - it provides the wrong table number for the table caption.
+It originally gave the wrong number here (it gave table 3), but moving the location of the label statement to its current location within the table environment fixed the problem here, but not in the table caption.
+Anyone know how to get this to work?
+TSC)}, even at the highest resolution, and lowest relative strength of low-order modes, the $A_{2-7}$ modes account for half of the total amplitude {\bf (TSC: but essentially all of the torque - see Fig.
+5.
+The question becomes, at what range of $m$ can modes begin to be considered ``local''?
+The results of Lodato \& Clarke 2004 suggests that $m = 5$ modes are
+reasonably local.
+)}.
 In order to accurately measure and characterize the contribution from these modes, one must consider a significant fraction of the disk.
@@
 on the order of 100\% or more are seen at all radii
-on  time scales range from fractions to integral multiples of the dynamic (orbital) time.  The temporal averaging
+on  time scales ranging from fractions to integral multiples of the dynamic (orbital) time.  The temporal averaging
 of figure \ref{fig:alpha_v_radius} washes out this local variability.
@@
 \begin{deluxetable}{cccccc}
+
  \tablecolumns{5}
@@
 }
+
 \startdata
@@
 \enddata  
+\label{tbl:ams}
 \caption{
@@
 }
-\label{tbl:ams}
+
 \end{deluxetable}
@@
 \begin{figure}
-\label{fig:alpha_v_radius}
 \plotone{figures/Figure6.eps}
@@
 }
+\label{fig:alpha_v_radius}
 \end{figure}

### 2011-06-08 Paper_1.tex
Tom's submitted version w/ pared down abstract.
	modified:   Paper_1.tex
	modified:   figures/Figure1.eps
	modified:   figures/Figure3.eps
	modified:   figures/Figure5.eps
	modified:   figures/Figure7.eps
	modified:   general.bib

--- 2010-12-22 Paper_1.tex
@@
 \usepackage{amssymb}
+\usepackage{ulem}
+\graphicspath{{figures/}}
+
+\definecolor{magenta}{rgb}{1.0,0.0,0.85}
+
+\newcommand{\ACBc}[1]{{\color{red} #1}}
+\newcommand{\TSCc}[1]{{\color{magenta} #1}}
+\newcommand{\SMc}[1]{{\color{blue} #1}}
 
@@
 
-\title{Convergence Studies on Mass Transport in Disks with Gravitational Instabilities.
+\title{Convergence Studies of Mass Transport in Disks with Gravitational Instabilities.
  \\
-I. The Constant Cooling Case}
+I. The Constant Cooling Time Case}
 \author{Scott Michael}
@@
 
-\author{Aaron Boley}
+\author{Aaron C. Boley}
 \affil{Department of Astronomy, University of Florida, Gainesville, FL 32611}
@@
 
-We conducted a convergence study of the azimuthal resolution in 3D hydrodynamic simulations of a protostellar disk susceptible to gravitational instabilities (GIs). 
-The disk was primarily studied during its asymptotic state, when
-heating and cooling are in balance. The cooling was treated as a constant global cooling time.
-Four simulations were conducted, identical except for the number of azimuthal grid 
-points, $l_{max} = $ 64, 128, 256 and 512, used in the calculations. 
-The strengths of the azimuthal modes, gravitational torques, and the effective \citet{shakura1973} $\alpha$ 
-arising from gravitational stresses were determined for each resolution.  We examined the resolution-dependence of each quantity
-and evaluated both the converged modal amplitude distribution and the effective value of $\alpha$
-as $l_{max} \rightarrow \infty$.   Non-axisymmetric amplitude was found in all modes and 
-the amplitude summed over all modes was essentially independent of resolution.
-The modal distribution of amplitudes, however, was significantly affected by resolution.  
-The size of the azimuthal grid, $l_{max}$,
-limits the number of modes that can be accurately measured to $l_{max}/2$.  Higher resolution simulations therefore
-have more modes available to them.  Non-axisymmetric structure is present in all available modes.  As a consequence,
-higher resolution simulations have a larger fraction of their total amplitude in high-order modes.
-High-order modes
-are more confined to act on local structure than low-order modes, therefore as the resolution increases and low-order amplitudes decrease the total gravitational stresses decreases as well.
-As a result, higher resolution simulations experience 
-weaker gravitational torques than lower resolution simulations.   The effective  $\alpha$
-also depends upon the magnitude of the stresses.  Thus
-$\alpha_{\rm{eff}}$ is also a function of grid resolution, decreasing as  $l_{max}$ increases.
-The converged $\alpha_{\rm{eff}}$ is consistent with
-predictions from an analytic local theory for thin disks by \citet{gammie2001}, 
-but only over many dynamic times when averaged over a substantial volume of the disk.
-Results obtained in the $l_{max} = 512$ simulations differ with converged values
-by less than 20\%.
+We conduct a convergence study of a protostellar disk, subject to a constant 
+global cooling time and susceptible to gravitational instabilities (GIs), at
+a time when heating and cooling are roughly balanced. Our goal is to 
+determine the gravitational torques produced by GIs, the level to which transport can
+be represented by a simple alpha-disk formulation, and to examine 
+fragmentation criteria. Four simulations are conducted, identical except for the number
+of azimuthal computational grid points used. A Fourier decomposition of 
+nonaxisymmetric density structures in  $\cos(m\phi)$, $\sin(m\phi)$ is performed to evaluate 
+the amplitudes $A_m$ of these structures. The $A_m$, gravitational torques, and
+the effective \citet{shakura1973} $\alpha$  arising from gravitational stresses
+are determined for each resolution. We find nonzero $A_m$ for all $m$-values and
+that $A_m$  summed over all $m$ is essentially independent of resolution. Because
+the number of measurable $m$-values is limited to half the number of azimuthal
+grid points, higher-resolution simulations have a larger fraction of their total
+amplitude in higher-order structures. These structures act more locally than
+lower-order structures. Therefore, as the resolution increases the total 
+gravitational stress decreases as well, leading higher-resolution simulations to experience
+weaker average gravitational torques than lower-resolution simulations. The 
+effective $\alpha$ also depends upon the magnitude of the stresses, thus
+$\alpha_{eff}$ also decreases with increasing resolution. Our converged
+$\alpha_{eff}$ is consistent with predictions from
+an analytic local theory for thin disks by Gammie (2001), but only over many
+dynamic times when averaged over a substantial volume of the disk.
 
@@
 Detailed three-dimensional hydrodynamic modeling of protostellar disks has provided considerable insight into the physical and thermal states of these disks.
-However,  computing resources limit these studies to snapshots spanning timescales much shorter than a disk's evolutionary lifetime.
-The level of spatial resolution (and hence temporal resolution, due to the Courant condition) used in calculations can play a deterministic role in interpretations of models; while lower resolutions allow more extensive spatial and temporal studies, this benefit can come at the sacrifice of important physics.
+However,  computing resources limit these studies to durations spanning timescales much shorter than a disk's evolutionary lifetime.
+The level of spatial resolution, and hence temporal resolution due to the Courant condition, can play a deterministic role in interpretations of calculations; while lower resolutions allow more extensive spatial and temporal studies, this benefit can come at the sacrifice of important physics.
 Convergence studies provide a means
-of deriving important outcomes without the expense of very high resolution simulations.
+of deriving important outcomes while assessing the resolution needed to determining these outcomes.
 
@@
 approaches exist that replace detailed physics with statistical parameters that encapsulate this physics.  This
-allows for calculations covering much longer timeframes.  Both detailed and 
+allows for calculations covering much longer time frames.  Both detailed and 
 parametric
@@
 
-%Here $\alpha$ is a dimensionless parameter that
-%provides a measure of viscosity, without providing a detailed specification of the viscous mechanism.
-
 Gravitational instabilities (GIs) can play an important, and at times dominant, role in driving the structural 
-evolution of  protostellar disks \citep[for reviews, see][]{durisen2005, durisen2010}. {\bf (We need the Durisen 2010 (2011?)
-review article reference)}
-Thermal processes play the primary role in regulating the amplitude and outcome of these instabilities \citep{pickett1998,pickett2000,mejia2005,nelson1998,nelson2000a}.
+evolution of  protostellar disks \citep[for reviews, see][]{durisen2005, durisen2007}.
+Thermal processes play the primary role in regulating the amplitude and outcome of these instabilities \citep{pickett1998,pickett2000,pickett2003,nelson1998,nelson2000a, mejia2005}.
 A disk's susceptibility to GIs can be parameterized by the Toomre $Q$-parameter \citep{toomre1981};  $Q = c_s \kappa / \pi G \Sigma$, where $c_s$ is the sound speed, $\kappa$ is the epicyclic frequency ($\sim$ the rotation frequency
 $\Omega$ in a nearly Keplerian disk), and $\Sigma$ is the disk surface mass density.
-For $Q \al$  1.5 to 1.7,  small density perturbations in a disk grow exponentially on a time scale comparable to the rotation period.
+For $Q \al$ 1.7,  small density perturbations in a disk grow exponentially on a timescale comparable to the rotation period \citep{durisen2007}.
 These perturbations manifest themselves as multi-arm spirals with a predominantly trailing pattern that transports angular momentum outward by gravitational torques \citep{larson1984,boss1984,durisen1986,papaloizou1991, laughlin1994,nelson1998,pickett1998}.
@@
 
+%\ACBc{[ACB: I've been thinking about the whole heating-by-pdV-to-stabilize-disks thingy.  I am not so sure it is true! Let's 
+%consider a purely isentropic compression.  In this case, $Q= (\gamma K)^{1/2}\Sigma^{\left(2-\gamma\right)/2}\kappa/(\pi G)$. 
+%A compression STILL destabilizes the disk.  One needs to change $K$. Have I set up a straw man by restricting myself to 
+%thinking of an isentropic compression? I do not think so because pdV should not change entropy. ]}
+
+We note here a caution on semantics.  GIs in disks commonly manifest themselves  as non-axisymmetric density structures whose strengths can be characterized 
+by a Fourier decomposition of the density in terms of $\sin m\phi$ and $\cos m\phi$, where 
+$m$ is the order of the Fourier term and $\phi$ is the usual 
+cylindrical coordinate.  While discrete coherent spiral waves seen in these  disks may well be correctly thought of as eigenmodes, particularly 
+at low $m$-values, the analysis necessary to confirm that these structures are, in fact, truly modes is generally lacking.
+Power in a specific $m$-value does not imply the existence of that mode, nor does it necessarily represent the strength of a mode that truly exists.
+Indeed, a disk with a {\it single} $m = 2$ eigenmode growing to nonlinear amplitudes will exhibit power at all even values of $m$.
+A disk with {\it two} nonlinear eigenmodes of
+$m = 2$ and 3 will exhibit power at all $m$-values.  
+For this reason, in the balance of this paper we will avoid referring to the Fourier terms as modes, but rather will refer to them as Fourier structures or components.
+The term {\it mode} will be reserved for those cases where eigenmodes are known, or strongly suspected, to exist.
+
 Fully nonlinear hydrodynamic simulations of low-$Q$ disks show that multiple
@@
 then power will exist at all $m$-values resolved by a given numerical method \citep{mejia2005,boley2006}.  
-In the asymptotic state, GI-active disks develop overlapping density structures of different strengths, geometries and coherence, producing strongly nonlinear turbulence that pervades the entire disk \citep[e.g.][]{pickett2003, mejia2005,boley2006}.
-\citet{gammie2001} refers to this state as {\it gravitoturbulence}.
-Gravitoturbulence provides a possible bridge between the detailed physics of GIs and the viscous transport parameter $\alpha$.
-Here convergence studies are important because
-higher order modes dissipate energy on shorter wavelengths than lower order modes.
-The mix of low and high order modes available to a disk is set by the resolution of the simulation.
-{\bf Multiple issues with this paragraph.
-Didn't want to rewrite unilaterally.
-1) In the first couple sentences the description of mode growth needs some tweaking, many modes appear in the linear regime, and although we have seen amplitude at all m values I don't know that it's a given.
-2) I wouldn't describe the structures formed by GI activity as ``strongly nonlinear turbulence''.
-3) In reviewing Gammie 2001 it seems that he is actually referring to a turbulent cascade (a fact I hadn't fully appreciated until now) I think it's pretty clear that the effective alpha we see is, in fact, not due to a turbulent cascade.
-i.e.
-alpha is driven by low order coherent modes.
-We should discuss this at some point SM}
+In the asymptotic state, GI-active disks develop overlapping density structures of different strengths, geometries and coherence that pervade the entire disk \citep[e.g.,][]{pickett2003, mejia2005,boley2006}.
+%\citet{gammie2001} refers to this state as {\it gravitoturbulence}. Gravitoturbulence 
+These structures provide a possible bridge between the detailed physics of GIs and the viscous transport parameter $\alpha$
+\citep{gammie2001}.  Here convergence studies are important because 
+higher-order modes dissipate energy on shorter wavelengths than lower-order modes; the mix of low and high-order azimuthal density structures
+available to a disk is set by the resolution of the simulation.
 
 \citet{shakura1973} proposed turbulence as the primary source of effective viscosity in gas disks
-and parameterized turbulent transport in a steady-state disk with the  parameter $\alpha$.  The source of the 
+and parameterized turbulent transport in a steady-state disk with the  parameter $\alpha$.  In this case,  the kinematic
+viscosity is defined by $\nu = \alpha c_s H$, where  $c_s$ is the sound speed and $H$ is the disk height.
+The source of the 
 turbulence in their $\alpha$-disk formulation is unspecified. Although their original work was directed towards X-ray bright
 accretion disks around black holes, this heuristic approach has proven fruitful in many astrophysical situations by allowing for analytic solutions and relatively easy numerical modeling over time-spans inaccessible to detailed modeling.
-In the
-$\alpha$-disk formalism, when the accretion rate $\dot{M}$ is constant in both time and radius, $\dot{M} = 3\pi\alpha c_s \Sigma H$ where $\Sigma$ is the surface density and $H$ represents
-the disk scale height. The $\alpha$ formalism makes no assumption about the nature of the angular momentum transport, only that {\it angular momentum transport
+%In the  $\alpha$-disk formalism, when the accretion rate $\dot{M}$ is constant in both time and radius, 
+%$\dot{M} = 3\pi\alpha c_s \Sigma H$ where $\Sigma$ is the surface density and $H$ represents
+%the disk scale height. 
+The $\alpha$ formalism makes no assumption about the nature of the angular momentum transport, only that {\it angular momentum transport
 and heating are dominated by local processes}. \citet{lin1987,lin1990} were among the first to suggest
@@
 
-Several studies support the idea that angular momentum transport by GIs in real disks is, in many important respects, an intrinsically global phenomenon and cannot be properly treated by a local $\alpha$-like prescription \citep{laughlin1996, balbus1999, lodato2005, mejia2005, boley2006, cai2008}.
+In fact, several studies support the idea that angular momentum transport by GIs in real disks is, in many important respects, an intrinsically global phenomenon and cannot be properly treated by a local $\alpha$-like prescription \citep{laughlin1996, balbus1999, lodato2005, mejia2005, boley2006, cai2008}.
 This almost certainly applies to the embedded phase of both high and low-mass stars when infall from the collapsing protostellar cloud feeds mass into the disk at a rapid
-rate \citep[e.g.]{yorke1993, laughlin1994, yorke1999,vorobyov2005,vorobyov2006}.  
-However, despite the long-distance nature of gravitational torques, theoretical and numerical results suggest that in protostellar disks where heating and cooling are in balance, i.e., disks subject to gravitoturbulence,
- it may be valid to represent GI-induced transport by an {\it effective viscosity}, a local process \citep{gammie2001, lodato2004, rice2005, cossins2009, vorobyov2010}.  In particular, 
-an $\alpha$-viscosity may be consistent
-with transport in GI active disks for disk to star mass ratio $M_d/M_* \le 0.2$ -- 0.3; GI-active disks
+rate \citep[e.g.,][]{yorke1993, laughlin1994, yorke1999,vorobyov2005,vorobyov2006}.  
+However, despite the long-distance nature of gravitational torques, theoretical and numerical results suggest that in protostellar disks where heating and cooling are in rough balance, 
+ it may be valid to represent GI-induced transport by an {\it effective viscosity}, as if it were a local process \citep{gammie2001, lodato2004, rice2005, cossins2009, vorobyov2010}.
+In particular, multiple studies suggest that
+an effective $\alpha$-viscosity may be consistent
+with transport in GI-active disks for a disk to star mass ratio $M_d/M_* \le 0.2$ -- 0.3; GI-active disks
 in systems with larger $M_d/M_*$ are dominated by lower-order modes that act in a more global sense and thus cannot be well represented by an $\alpha$-approach
 \citep{lodato2004, cossins2009, vorobyov2010}. 
+We note, however, that the these studies all assumed
+a {\it local cooling time}, $t_{cool}\Omega =$ constant, 
+where $t_{cool}$ is defined by the rate at which cooling decreases the disk's internal energy density
+$\epsilon$ by $d\epsilon/dt = \epsilon / t_{cool}$.  In contrast, we and our collaborators use either a {\it global cooling time},
+$t_{cool} = $ constant \citep[e.g.,][this paper]{pickett2003, mejia2005}, or use realistic opacities \citep[e.g.,][]{mejiaphd2004, cai2006, cai2008, boley2006, boley2007,  boley2009} which lead to something in between $t_{cool} = $ constant  and
+$t_{cool}\Omega =$ constant.
 
-\citet{gammie2001} derived an effective alpha for the case of a thin 2D gravitationally unstable disk where the heating caused by GIs was balanced by the local cooling rate.  
-Assuming a razor thin gravitationally unstable disk where the heating caused by
-GIs was balanced by a local cooling rate of $t_{cool}\Omega$.
-Gammie found
+\citet{gammie2001} derived an effective alpha for the case of a thin gravitationally unstable disk where the heating caused by GIs was balanced by the local cooling rate.  
+Assuming a local cooling time of $t_{cool}\Omega =$ constant,  Gammie found
 \begin{equation}
@@
 \end{equation}
-where $\gamma$ is the two dimensional adiabatic index and $t_{cool}$ is a cooling parameter. 
-The two dimensional index relates
-to the three dimensional index, $\Gamma$, by 
+where $\gamma$ is the two-dimensional adiabatic index. 
+The two-dimensional index relates
+to the three-dimensional adiabatic index, $\Gamma$, by 
 $\gamma = (3\Gamma - 1) / (\Gamma + 1)$
@@
 $\gamma = 3 - 2/\Gamma $
-for a strongly self gravitating disk.   \cite{lodato2004} derived a similar result when considering three dimensional protoplanetary disks.
-Motivated by
-the results from \cite{laughlin1996}, \cite{gammie2001}, and \cite{lodato2004, lodato2005}, which demonstrated differences in the manifestation of high and low-order, \citet{kratter2008} 
-proposed $\alpha_{\rm{eff}} = (\alpha_{short}^2 + \alpha_{long}^2)^{1/2}$. Here $\alpha_{short}$ and $\alpha_{long}$
-are functions of $Q$ and $M_d/\left( M_d+M_*\right)$ representing 
-the different wavelength regimes expected to dominate for different values of the mass ratio and cooling rate. 
-Fully 3D simulations \citep[e.g.][]{lodato2004, cossins2009} suggest that, at least in the case of imposed cooling, $\alpha_{\rm{\rm{eff}}}$ roughly converges to the value predicted by \citeauthor{gammie2001}.
+for a strongly self gravitating disk.   \cite{lodato2004} derived a similar result when considering three-dimensional protoplanetary disks.
+Fully 3D simulations \citep[e.g.,][]{lodato2004, cossins2009} suggest that, at least in the case of imposed 
+$t_{cool}\Omega = $ constant cooling, $\alpha_{\rm{\rm{eff}}}$ roughly converges to the value predicted by \citeauthor{gammie2001}.
 
-We address these, and similar questions, using a grid-based, finite-difference 3D hydrodynamics code
-with self gravity.
-For computational grids of a fixed {\it physical} size,
-grids with more elements provide finer meshes, and thus a more accurate solution. 
-However, increased accuracy is offset by computational penalties.
-The number of equations solved at each time step 
-increases as the grid size is increased.  At the same time, the Courant condition 
-forces shorter time steps.  A mesh convergence study 
+In what follows, we address these and related questions using a grid-based, finite-difference 3D hydrodynamics code with 
+self-gravity to follow the evolution of a protostellar disk with a constant cooling time and subject to GIs.
+We examine how grid resolution affects
+the amplitude distribution of azimuthal density structures arising from GIs, the gravitational torques arising from these structures,
+the effective $\alpha$, and the extent to which transport can be represented by an $\alpha$-disk formulation.
+%For computational grids of a fixed {\it physical} size,
+%grids with more elements provide finer meshes, and thus a more accurate solution. 
+%However, increased accuracy is offset by computational penalties.
+%The number of equations solved at each time step 
+%increases as the grid size is increased.  At the same time, the Courant condition 
+%forces shorter time steps.  
+A mesh convergence study 
 provides a common approach to balancing accuracy and computing resources.
-In the following sections we follow the evolution of a disk subject to GIs
-using four different
-resolutions in the azimuthal direction.  The number of azimuthal grid points used in simulations
+It permits an assessment of whether or not convergence occurs and what resolution is required.
+%allowing us to determine if convergence occurs and what is the required resol
+In the following sections, we follow the evolution of a disk subject to GIs
+using grids with four different
+resolutions in the azimuthal direction.  The number of azimuthal elements used in simulations
 may be particularly important, because higher azimuthal resolution 
-allows non-axisymmetric amplitude to spread to higher-order modes 
-that behave more locally. We test this by examining the amplitude in various 
-modes to see how they are affected by the choice of grid.
+allows power in non-axisymmetric density structures to spread to higher-order modes 
+that behave more locally. 
+Although full convergence testing eventually requires investigation of the radial and vertical resolution in our
+cylindrical grid, these tests are expensive and will be left to a future paper.  We suspect that the azimuthal
+direction is most important to test first because it is non-axisymmetry that produces the torques leading
+to mass and angular momentum transport.
+We test this by examining the amplitude in various 
+$m$-values to see how they are affected by the choice of grid.
 In the process, 
-we report the results of a convergence study designed to determine whether GI-active disks, chaotic by nature, achieve statistical equilibrium If this is the case, we investigate how closely these disks approximate the gravitoturbulence of \citet{gammie2001} and examine the distribution of $\alpha_{\rm{eff}}$.
-To this end, we follow the evolution of a GI-active disk, subject to constant cooling, in its asymptotic state, where cooling and heating are essentially balanced.
-Simulations are performed using four different resolutions in the azimuthal direction.
+we report the results of a convergence study designed to determine whether GI-active disks, chaotic by nature, achieve statistical equilibrium.
+If this is the case, we investigate how closely these disks approximate equation (1) from
+ \citet{gammie2001} and examine the distribution of $\alpha_{\rm{eff}}$.
 
-The balance of this paper is organized as follows.  Section 2 provides the details of the numerical approach and defines the models.  An analysis of the modal structure of the simulations 
-is found in \S3.1, the gravitational torques arising from these modes follows in \S3.2.  Effective converged alphas are 
-discussed in \S3.3.  Section 3.4 deals with the degree to which selected grid sizes lead to results consistent with
+The balance of this paper is organized as follows.
+Section 2 provides the details of the numerical approach and defines the models.
+An analysis of the azimuthal density structures of the simulations
+is found in \S3.1, the gravitational torques arising from these structures follows in \S3.2.  Effective converged alphas are 
+discussed in \S3.3.  Section 3.4 deals with the degree to which selected grid-sizes lead to results consistent with
 convergence.  Fragmentation is discussed in \S3.5.  A summary of results is found in \S4.
-
-%We cannot yet follow the entire dynamical evolution of a 
-%protoplanetary disk nor
-%model all the radiative and particle processes of interest at once in full 3D. However, 
-%by following GI simulations to their asymptotic states, when these exist, we
-%can extrapolate the consequences over much longer times (e.g., Boley et al 2006).
 
@@
 Simulations were conducted with the three-dimensional hydrodynamics code used by the Indiana University Hydrodynamics Group in several previous studies \citep{pickett1998, pickett2000, mejiaphd2004, mejia2005, boley2006, cai2008}.
-This code uses a second-order, explicit, Eulerian scheme to solve Poisson's equation, an equation of state, and the equations of hydrodynamics in conservative form on a uniform cylindrical grid.
+This code uses a second-order, explicit, Eulerian scheme to solve Poisson's equation, an ideal gas equation of state, and the equations of hydrodynamics in conservative form on a uniform cylindrical grid.
 The code assumes mirror symmetry about the equatorial plane.
@@
 
-We assume an equation of state $P = (\gamma -1)\epsilon$, where $P$ is the pressure, $\epsilon$ is the internal energy density, and $\gamma$ is the ratio of specific heats, given here by $\gamma = 5/3$.
-The model disks are cooled by decreasing their internal energy according to the prescription $d\epsilon/dt = \epsilon / t_{cool}$, where $t_{cool}$ is the global cooling time in Outer Rotation Periods (ORP).
-For the grid resolutions used here, one ORP is defined as the rotational period at radial zone 200.
-
-%\subsection{Models}
+Initial conditions for the disk's structure and thermodynamic state were set using an equilibrium star plus disk model generated by a modified \citet{hachisu1986} self-consistent field relaxation method \citep{pickett1996,pickett2003,mejiaphd2004,mejia2005,cai2006}.
+Here  the density and angular momentum distributions are iteratively solved, using the specified equation of state,  until convergence is achieved.
+This  equilibrium disk was
+given random cell-to-cell $\sim10^{-4}$ density perturbations to seed growth of  GIs.  The resultant marginally unstable disk served as the initial disk for the simulation.
 
@@
 In what follows, quantities specified with real physical units are derived assuming a 1 $M_\odot$ central star and a 0.153  $M_\odot$ disk.
-In this case, the ORP is defined at $r = 33$ AU and one ORP $\approx$ 303 years.
-Models have $t_{cool} = 2$ ORP.
-The disk has an initial surface density $\Sigma \sim r^{ -1}$, and initial inner and outer radii of 2.3 and 40 AU, respectively.
+In this case, the outer rotation period (ORP) is defined at $r = 33$ AU and one ORP $\approx$ 214 years.
+The disk has an initial surface density that approximates $\Sigma \sim r^{ -1}$, with initial inner and outer radii of 2.3 and 40 AU, respectively.
 
-Initial conditions for the disk's structure and thermodynamic state were set using an equilibrium star plus disk model generated using a modified \citet{hachisu1986} self-consistent field (SCF) relaxation method \citep{pickett1996,pickett2003,mejiaphd2004,mejia2005,cai2006}.
-Here  the density and angular momentum distributions are iteratively solved, using the specified equation of state,  until convergence is achieved.
-This  equilibrium disk was then seeded with small 0.01\% $\delta\rho/\rho$ random perturbations to allow GIs to grow.
-The resultant marginally unstable disk served as the initial disk for the simulation.
+ We assume an equation of state $P = (\gamma -1)\epsilon$, where $P$ is the pressure, $\epsilon$ is the internal energy density, and $\gamma$ is the ratio of specific heats, given here by $\gamma = 5/3$.
+The model disks are cooled by decreasing their internal energy according to the prescription $d\epsilon/dt = \epsilon / t_{cool}$, where $t_{cool}$ is the global cooling time, which is set to 2 ORP everywhere.
 
 The initially unstable disk model passed through several phases of evolution \citep[see also][]{pickett2003, mejia2005}.  
-The initial disk is unstable to the growth of non-axisymmetric structures.
 During the initial axisymmetric phase, 
-which lasts several ORPs, the disk cools and contracts, the contraction is small radially but dramatic in the vertical direction.
-Around 3-4 ORPs the instabilities begin grow to non-linear amplitudes.
-When this condition is met, the disk undergoes a strong burst of GIs that predominantly manifest themselves in one or two discrete global spiral modes.
+which lasts several ORPs, the disk cools and contracts.
+The contraction is small radially but dramatic in the vertical direction.
+Around 3 to 4 ORPs, the instabilities begin to grow to nonlinear amplitudes, and the disk
+then undergoes a strong burst of GIs that predominantly manifest themselves in a few discrete global spiral modes.
 The disk expands violently during the burst phase, producing a significant rearrangement of the disk's mass distribution on a timescale of a few ORPs.
 This is then followed by a period of several ORPs where heating temporarily washes out some of the non-axisymmetry in the disk.
-Finally, following this transition phase, the disk settles into a quasi-steady, long-lived asymptotic state of sustained GI-activity over a large part of the disk, with an overall balance of heating and cooling.
+Finally, following this transition phase, the disk settles into a quasi-steady, long-lived asymptotic
+phase of sustained GI-activity over a large part of the disk, with an overall balance of heating and cooling.
 
-Simulations were run with four different angular resolutions, with grid sizes in $(r,\phi,z)$ of  $(j_{max}, k_{max}, l_{max}) =$ (512,64,64), (512,64,128), (512,64,256), and (512,64,512).
+Simulations were run with four different angular resolutions, with grid sizes in cylindrical coordinates $(r,\phi,z)$ of  $(j_{max}, k_{max}, l_{max}) =$ (512,64,64), (512,64,128), (512,64,256), and (512,64,512).
+For reference, the radius $r = 33$ AU at which the ORP is defined is at radial cell $j = 200$.  The radial grid extends to 512
+to accomodate disk expansion during outburst.
 Because of the computational overhead associated with running each simulation starting from $t = 0$, the $l_{max} =$ 128 simulation was run through the axisymmetric, burst and transition phases.
-Simulations using $l_{max} =$ 64, 256, and 512 all begin by interpolating the $l_{max} =$ 128 simulation to higher or lower azimuthal resolution at 9.6 ORPs, near the end of the transition
+Simulations using $l_{max} =$ 64, 256, and 512 all begin by a linear interpolation of the $l_{max} =$ 128 simulation to higher or lower azimuthal resolution at 9.6 ORPs, near the end of the transition
 phase.
 Although this strategy saves a large amount of computational time, it limits the scope of the analysis to the asymptotic phase.
-By 12 ORPs, all four disks have transitioned to the asymptotic state.
-All simulations were
+By 12 ORPs, all four disks appear to have transitioned to the asymptotic state,
+e.g., they all display roughly steady $Q(r)$ profiles starting at $\sim 12$ ORPs.  All simulations were 
 run through $\sim 18$ ORPs.
@@
 Figure \ref{fig:Final_Q} shows the Toomre $Q$, as a function of radius, for all four simulations at $t \sim$ 18 ORP.  
-Table \ref{tbl:ams} lists $Q_{avg}$, the value of $Q$ of each of the asymptotic disks, time-averaged from 12 -- 19 ORP and spatially averaged over 10 -- 40 AU.
-From $\sim$ 12 -- 50 AU, $Q \sim 1.0$ -- 1.4 and thus the disk is subject GIs over this full range.
-Interior to $\sim$ 27 AU, the value of $Q$ shows no dependence on
-the azimuthal resolution.
-From 27 -- 50 AU, lower resolution simulations display lower values of $Q$.
-{\bf I suppose 27 AU was chosen by eye?
-Or was there some other analysis.
-I vote to toss these last two sentences as I'm not convinced w/o a time averaged quantity.
-Unless this observation is used in further reasoning in the discussion...SM}
+%Table \ref{tbl:ams} 
+Table 1 lists $Q_{avg}$, the arithmetic mean of $Q$ for each of the asymptotic disks, time-averaged from 12 -- 18 ORP and spatially averaged over 10 -- 40 AU.
+From $\sim$ 12 -- 50 AU, $Q \sim 1.0$ -- 1.4 and thus the disk is subject to GIs over this full range.
 
@@
 
-In the asymptotic state, GI-active disks develop complex density structures arising from the superposition of
+\subsection{Non-axisymmetric Structure}
+
+In the asymptotic state, GI-active disks develop complex density structures arising from the superposition of non-axisymmetric
 modes in the disk.
-Since the number of modes accessible to the disk depends on the azimuthal resolution, we expect that higher resolution simulations will exhibit non-axisymmetric amplitude in higher-order modes inaccessible at lower resolution, and these higher-order modes will
-behave more locally. We test this by examining the amplitude in various modes to see how they are affected by the choice of grid.
-
-\subsection{Modes}
+Since the number of modes accessible to the disk depends on the azimuthal resolution \citep{shannon1984}, we expect that higher resolution simulations will exhibit non-axisymmetric amplitude in higher-order modes inaccessible at lower resolution, and these higher-order modes will
+behave more locally.
 
@@
 resolution.
-One can clearly see that the lower resolution simulations, i.e.
-$l_{max} = 64$ and 128, have primarily low-order structures with $m = 2$ and 3 dominating.
-In contrast, the higher resolution simulations have much more fine structure, the modes are much more washed out, and the low-order modes no longer dominate.
+One can clearly see that the lower resolution simulations, i.e., $l_{max} = 64$ and 128, have primarily low-order structures with $m = 2$ and 3 dominating.
+In contrast, the higher resolution simulations have much more fine structure,
+and the low-order $m$-values no longer dominate.  
 The lower resolution (smaller $l_{max}$) simulations display longer wavelength spiral structures, with larger amplitudes and more coherence than those seen in higher resolution simulations.
-Higher angular resolution allows non-axisymmetric structures to grow in higher-order azimuthal modes, as demonstrated by a Fourier decomposition of the density.
-The strengths of these modes are given by their global Fourier amplitudes,
+Higher angular resolution allows non-axisymmetric structures to grow in higher-order azimuthal $m$-values, as demonstrated by a Fourier decomposition of the density.
+The strengths of these components are given by their global Fourier amplitudes,
 
@@
 \end{subequations}
+Here  $\rho_o$ is the axisymmetric component of the density.
+In what follows, $A_m$ is averaged over time to suppress fluctuations on the dynamic timescale due to the chaotic nature of GIs in the asymptotic  phase.
+In particular, $\langle A_{tot} \rangle$ will represent the power over the radial range 10 -- 40 AU, summed from $ m = 2$ to $l_{max}/2$ and averaged over the time range 12 -- 18 ORPs.
+The Fourier amplitudes cannot be measured at resolutions smaller than $l_{max}/2$, therefore $A_m$ is limited to modes $m \le l_{max}/2$.
+Because the central star is artificially fixed to the grid center, the $m=1$ mode may not
+be accurately treated. For this reason, it is excluded from $\langle A_{tot} \rangle$.  As discussed below in \S3.2,
+we expect this exclusion to have minimal impact on the results of this work.
 
-Here  $\rho_o$ is the axisymmetric component of the density.
-In what follows, $A_m$ is averaged over time to suppress fluctuations on the dynamic time scale due to the chaotic nature of GIs in the asymptotic  phase.
-In particular, $\langle A_{tot} \rangle$ will represent the power over the radial range 10 -- 40 AU, integrated from $ m = 2$ to $l_{max}/2$ and averaged over the time range 12 -- 19 ORPs.
-The Fourier amplitudes cannot be accurately measured at resolutions smaller than $l_{max}/2$ \citep{shannon1984}, therefore $A_m$ is limited to modes $m \le l_{max}/2$.
-Because the central star is artificially fixed to the grid center, the dynamics of the $m=1$ mode may not
-be accurately treated. For this reason, it is excluded from $\langle A_{tot} \rangle$.
+Previous studies have shown that disks in the asymptotic state have power at all resolvable $m$-values  \citep
+{lodato2004,mejia2005,boley2006,cossins2009}.  This can be seen in
+figure 3, where $\log \langle A_m\rangle$ is displayed as a function of $\log(m)$.
+For all resolutions, components with the highest global amplitudes are those with low $m$-values; $m=2$ through 6 are labeled.
+In what follows, the term ``low-order''  will refer to Fourier components with $m <  8$.
+Table 1 lists $\langle A_{tot} \rangle$ along with $\langle A_{2-7} 
+\rangle  /  \langle A_{tot} \rangle$,  the fraction of the total power falling into low-order ($m = 2$ -- 7) azimuthal structures, 
+for each resolution. 
+We identify $m=2-7$ as representing amplitude in the low-order modes, i.e., coherent structure that is
+global in character.   Again 
+$m=1$ is excluded due to the fixed central star. This determination of $m$-range
+was made by considering the radial range over 
+which the mode is most effective in transporting angular momentum, i.e., from the inner Linblad resonance to the outer 
+Linblad resonance, and comparing it to the disk scale height. These quantities are roughly equal for $m=7$; for 
+lower-order modes the radial range exceeds the scale height. The vertical dotted line in figure 3 corresponds with $m=7$. 
+While the total power is approximately independent of resolution, $\langle A_{tot} \rangle \sim 1.90$, 
+the distribution of this power amongst $m$-values
+shows marked differences. In particular, the fractional power in lower-order $m$-values decreases monotonically with increasing resolution.  Over three quarters of $\langle A_{tot} \rangle$ resides
+in $m =$ 2 -- 7 for the $l_{max} = 64$ simulation, while approximately half resides in these $m$-values when $l_{max} =$ 512.
 
-{\bf (Note to all: Throughout the paper I use integrals in places where we 
-formally have used summations.  I note that this is a common practice, though
-not strictly correct.  I find the use of integrals more elegant, but am willing to
-revert to summations if people push for it.)}
-
-Previous studies have shown that disks in the asymptotic state have power at all resolvable $m$-values  \citep{lodato2004,mejia2005,boley2006,cossins2009}.
-Table \ref{tbl:ams} lists $\langle A_{tot} \rangle$ along with $\langle A_{2-7} \rangle  /  \langle A_{tot} \rangle$,  the fraction of the total power falling into low-order ($m = 2$ -- 7) modes, for each resolution.
-We identify $m=2-7$ as the low-order modes that are considered to be global modes, again
-$m=1$ is excluded due to the fixed central star. This determination was made by considering the radial range over 
-which the mode is most effective in transporting angular momentum, i.e. from the inner Linblad resonance to the outer 
-Linblad resonance, and comparing it to the disk scale height. These quantities are roughly equal for $m=7$, for lower 
-order modes the radial range exceeds the scale height.
-While the total power is approximately independent of resolution, $\langle A_{tot} \rangle = 1.90 \pm 0.07$, 
-the distribution of this power shows marked differences.
-In particular, the fractional power in lower order modes decreases monotonically with increasing resolution.
-Over three quarters of $\langle A_{tot} \rangle$ resides
-in modes $m =$ 2 -- 7 for the $l_{max} = 64$ simulation, while approximately half resides in these modes when $l_{max} =$ 512.
-
-The shift of power from lower to higher order modes as the resolution of the computational grid is increased arises from the fact that, while $\langle A_{tot} \rangle$ is approximately independent of $l_{max}$,  the number of degrees of freedom available for this power, $l_{max}/2$, increases linearly with the resolution.
-Thus amplitude is naturally spread from lower order to higher order modes as the available $m$-values increase.
-Indeed, even if the disk is dominated by one or two low-order modes at the time of outburst, this power cascades to higher-order modes \citep[for example][]{laughlin1997, laughlin1998,laughlin1996}.
-{\bf Haven't had a chance to check these refs but I'm reluctant to call what we see a cascade of power.
-I suggest striking this sentence.
-SM}
+The shift of power from lower-order to higher-order azimuthal structure as the resolution of the computational grid is increased arises from the fact that, while $\langle A_{tot} \rangle$ is approximately independent of $l_{max}$,  the number of degrees of freedom available for this power, $l_{max}/2$, increases linearly with the resolution.
+Thus amplitude is naturally spread from lower-order to higher-order as the available $m$-values increase.
+Indeed, even if the disk is dominated by a few low-order modes at the time of outburst, this power spreads to all
+$m$-values in the nonlinear regime \citep[for example][]{laughlin1997, laughlin1998,laughlin1996}.
 
 Given how the non-axisymmetric amplitude distribution depends on the azimuthal resolution, we must examine what resolution is sufficient to properly capture the evolution of these disks.
-Table \ref{tbl:ams} shows that while the fractional power in lower order modes decreases with each doubling of $l_{max}$, the amount of change in going from $l_{max} = 256$ to 512
+Table 1 shows that while the fractional power in
+lower-order $m$-values
+decreases with each doubling of $l_{max}$, the amount of change in going from $l_{max} = 256$ to 512 
 is significantly less than the change going from $l_{max} = 64$ to 128, with a strong suggestion that this diminution will not proceed much further with higher azimuthal resolution.
-Indeed, a  Richardson extrapolation \citep{press1992} of the tabulated values of  $\langle A_{2-7} \rangle / \langle A_{tot} \rangle$ suggests a limiting value of  $\langle A_{2-7} \rangle / \langle A_{tot} \rangle \sim$ 0.42 for increasingly higher resolutions.
-This is displayed graphically in figure \ref{fig:Power_Extrap}, where the fractional power in lower order modes is plotted against the  normalized grid
+Indeed, a  Richardson extrapolation \citep{press1992} of the tabulated values of  $\langle A_{2-7} \rangle / \langle A_{tot} \rangle$ suggests a limiting value of  $\langle A_{2-7} \rangle / \langle A_{tot} \rangle \sim$ 0.43 for increasingly higher resolutions.
+This is displayed graphically in figure \ref{fig:Power_Extrap}, where the fractional amplitude in lower-orders is plotted against the  normalized grid
 spacing for each value of $l_{max}$. The grid spacing  is normalized such that spacing of the $l_{max}$ = 512 grid
 equals unity. 
-Convergence occurs as the normalized spacing goes to zero.
-The Richardson extrapolation value  of $\langle A_{2-7}\rangle  / \langle A_{tot}\rangle  = 0.42$ is plotted at a normalized grid spacing  equal to zero.
-%The majority of the modal power in a converged disk resides in higher-order modes. 
-%These modes dissipate on shorter wavelengths than low-order modes and hence act more locally.
+Convergence occurs as the normalized spacing goes to zero. The Richardson extrapolation value  
+of $\langle A_{2-7}\rangle  / \langle A_{tot}\rangle  = 0.43$ is plotted at a normalized grid spacing  equal to zero.
+
+%\ACBc{[ACB: This is good, but I think we can do better by: (1) getting a handle on what we think the statistical variation in the %$A$ values is, and (2) using that to get an extrapolated value that has some likelihood value.]} \TSCc{[TSC: I think this would %be overkill.
+It would involve a bit of work and I think we would not gain any additional insight into the problem by doing this %work.
+Added note - During my talk with Aaron, he seemed adamant about this.
+Dealing with point would take a bit
+%of time and effort.  I am not convinced it is worthwhile, but I defer to the rest of the group.]}
 
@@
 
-Low-order spiral modes in $\Delta\rho / \rho$  have longer wavelengths that, if coherent in radius, have longer lever 
+Low-order spiral modes in the density distribution have longer wavelengths that, if coherent in radius, have longer lever 
 arms to produce torques, while high-order modes have relatively short wavelengths which, particularly if they 
-lack coherence can cancel each other out, and produce more localized effects. 
-The finding that the fractional power in high-order modes increases
+lack coherence, can cancel each other out and produce more localized effects. 
+The finding that the fractional power in high-order structure increases
 with resolution, while the total power stays the same, leads to
- the expectation that the gravitational torques in the disk will decrease as $l_{max}$ is increased. This can be tested by examining the torque contributions from  various Fourier modes.
+ the expectation that the gravitational torques in the disk will decrease as $l_{max}$ is increased. This can be tested by examining the torque contributions from  various Fourier components.
 
-The torque {\bf C} on a cylindrical surface of the disk at radius $\varpi$ can be obtained by integrating the stress tensor $T$ over the surface of the cylinder \citep{lyndenbell1972}, i.e.,
+The torque {\bf C} on a cylindrical surface of the disk at radius $\varpi$ can be obtained by integrating the stress tensor $T$ over the surface of the cylinder \citep{lyndenbell1972, boley2006}, i.e.,
 \begin{equation}
@@
 where $a_{\phi m} =\slantfrac{1}{\pi} \int \rho \cos(m\phi){\rm d}\phi$ and $b_{\phi m} =\slantfrac{1}{\pi} \int \rho \sin(m\phi){\rm d}\phi$, and only the gravitational potential produced by the mass distribution given by $\rho_m$ is included in $\Phi$.
+Note that, as defined here, $\rho_m$ is a  function of $\varpi$, $z$ and $\phi$ while $a_{\phi m}$ and
+$b_{\phi m}$ are functions of $\varpi$ and $z$.
 
@@
  time-averaged torques summed over a number of low-order modes, $\sum_1^n \langle {\rm \bf C}_{Z(n)} \rangle$, $n = $ 1, 2, 3, 4, and 5, and $\langle {\rm \bf C}_{Ztot} \rangle$,
-the total time-averaged gravitational torque arising from modes 1 thru $l_{max}/2$, for four different time-frames,
+the total time-averaged gravitational torque arising from modes 1 through $l_{max}/2$
 as a function of radius and azimuthal resolution. 
@@
 should not be considered realistic.  
-\citet{michael2010} found that freeing the star changes the torques by 10-20\% at most, 
+\citet{michael2010} found that freeing the star changes the torques by about 10-20\%, 
 thus the difference between torques arising from the $m=1$ mode with the star fixed and free to move are small compared to the differences found when comparing azimuthal resolution.
 
-We note two results of figure \ref{fig:torquearray}: 1) the total gravitational torque is smaller for higher resolution simulations, and 2) the total torque is dominated by low-order modes for all azimuthal resolutions.
-Both reflect the fact that while a higher fraction of the total power in larger $l_{max}$ simulations is in higher order modes, these modes make only minimal contributions to the gravitational torque.
-As resolution increases, non-axisymmetric amplitude is shifted from low-order global modes, which produce relatively large gravitational torques, to high-order local modes, which tend to cancel each other when integrated over the whole disk.
-
-{\bf (Note to all: Figure 5 shows torques
-time-averaged over for four different time-frames.  I created multiple frames because I wanted to see if torques were 
-diminishing with time, as might be expected if the disk has not achieved its asymptotic state.  
-Dick has told me he only believes the 12-18 ORP result - Dick, I hope 
-I'm not misquoting you. Should we go with multiple panels representing different time-frames or 
-go solely with the 12 - 18 orp panel?
-}
-
-{\bf I still think the scale in figure 5 is off by a factor of 2 probably due to the fact that you've used a 0.5 $M_\odot$ star. I also vote for a single 12-18 ORP panel. SM}
+We note two results in figure \ref{fig:torquearray}: 1) the total gravitational torque is smaller for higher resolution simulations and 2) the total torque is dominated by low-order Fourier components for all azimuthal resolutions.
+Both reflect the fact that while a higher fraction of the total power in larger $l_{max}$ simulations is in higher-order $m$-values, these structures make only minimal contributions to the gravitational torque.
+As resolution increases, non-axisymmetric amplitude is shifted from
+low-order global components, which produce relatively large gravitational torques, to high-order local structures, which tend to cancel each other when integrated over the whole disk.
 
@@
 
-%An $\alpha$-framework treats disks as essentially two-dimensional objects subject to local energy dissipation.
-Therefore, in order to compare our results with predicted and published results, we must extract two-dimensional information from our three-dimensional disks.
-
-{\bf I've made some significant changes to this text from Tom's original. Specifically in the first, fourth, and fifth paragraphs. Please look over and make sure this fits in properly. SM}
-
 After obtaining the gravitational stresses in the disk, one can compute an effective $\alpha$. 
-Following \cite{gammie2001} and \citet{lodato2004}, 
+Following \cite{gammie2001}, \citet{lodato2004}, and \citet{boley2006}, 
 \begin{equation}
@@
 and $c_s$ the sound speed.
-In comparison to \citeauthor{gammie2001}'s 2D models and \citeauthor{lodato2004}'s highly flattened smoothed particle hydrodynamics models, the models presented here have a significant vertical extent.
-Because of this, the sound speed can vary dramatically from the mid-plane to the low density regions, so it was unclear as to the best way to determine a representative sound speed.
+In comparison to \citeauthor{gammie2001}'s 2D models, the models studied here have a significant vertical extent.
+Because of this, the sound speed can vary dramatically from the mid-plane to the low density regions.
 In order to include contributions to the sound speed from the entire vertical extent, the denominator $\Sigma c_s^2$ was evaluated as $\int_z \rho c_s^2 {\rm d}z$.
@@
 %\begin{equation}
-%\alpha_{eff} = [\gamma(\gamma - 1) \slantfrac{9}{4}\Omega t_{cool}]^{-1}
+%\alpha_{\rm{eff}} = [\gamma(\gamma - 1) \slantfrac{9}{4}\Omega t_{cool}]^{-1}
 %\end{equation}
@@
 This difficulty is mitigated by the fact that 
-previous studies have found Reynolds stresses to be small relative to gravitational stresses -- see, 
-for example, figure 5 of \cite{lodato2004} and figure 10 of \cite{boley2006} -- suggesting that this term
-can be safely neglected in the evaluation of $\alpha_{\rm{eff}}$.   For this reason, 
+previous studies have found Reynolds stresses to be small relative to gravitational stresses (see, 
+for example, figure 5 of Lodato \& Rice 2004 and figure 10 of Boley et al. 2006).  This  suggests that the Reynolds stress
+can be neglected in the evaluation of $\alpha_{\rm{eff}}$.   For this reason, 
 here we omit
@@
 
-The Newton stresses of equation \eqref{alpha1} can be evaluated using the gravitational torque of equation \eqref{torque1}, i.e.,
+The gravitational stresses of equation \eqref{alpha1} can be evaluated using the gravitational torque of equation \eqref{torque1}, i.e.,
 \begin{equation}
@@
 To determine whether the converged value of $\alpha_{\rm{eff}}$ is consistent with predictions, we examine 
-the spatially (10 -- 40 AU) and temporally (12 -- 19 ORP) averaged  $\alpha_{avg}$,
-provided in table \ref{tbl:ams}, for each value of  $l_{max}$.    A Richardson extrapolation applied to these values  
+the spatially (10 -- 40 AU) and temporally (12 -- 18 ORP) averaged  $\alpha_{avg}$,
+provided in table 1, for each value of  $l_{max}$.    A Richardson extrapolation applied to these values  
 converges to $\alpha_{avg} = 0.017$ (over the same radial  and time range)
-as the grid spacing goes to zero.   This falls squarely
+as the grid spacing goes to zero.   This falls
 between the $\alpha_{avg}$ values for the Gammie curves 0.014 (strongly self-gravitating) and 0.027 (non-self-gravitating).
@@
 However, several cautionary notes are in order.
-It is important to remember that the bulk of the gravitational torque, and therefore the $\alpha_{\rm{eff}}$, is generated by low order modes that are coherent over tens of AU.
-As table \ref{tbl:ams} shows, even at the highest resolution, and lowest relative strength of low-order modes, the $A_{2-7}$ modes account for half of the total amplitude.
+It is important to remember that the bulk of the gravitational torque, and therefore the $\alpha_{\rm{eff}}$, is generated by low-order
+Fourier components that are coherent over tens of AU.  As table 1 shows, even at the highest resolution, where
+the relative strength of low-order structures is smallest, $\langle A_{2-7} \rangle$ accounts for half of the total amplitude.
 In order to accurately measure and characterize the contribution from these modes, one must consider a significant fraction of the disk.
-Furthermore, simulations of disks with a large radial extent (e.g.
-100s of AU \citep{boley2009}) show coherent low order modes over a significant fraction of the disk.
-In order to determine the radial extent and strength of such modes one should consider the entire range of radii over which the disk is susceptible to instabilities i.e.
-$Q \lesssim 1.5$.
+Even when averaged over 6 ORP, figure 6 shows that $\alpha_{\rm{eff}}$ displays $\sim 100$\% spatial fluctuations
+over a few AU.
+Furthermore, simulations of disks with a larger radial extent, e.g., 100s of AU \citep{boley2009}, show coherent 
+low-order modes over a significant fraction of the disk. 
+In order to determine the radial extent and strength of such modes, one should consider the entire range of radii over which the disk is susceptible to instabilities, i.e., $Q \lesssim 1.5$ (figure 1).
+In this context,
+\cite{balbus1999}
+argue that a self-consistent local-disk picture, as required for a proper $\alpha$-disk formulation,
+is probably not possible for GI-unstable disks,
+except perhaps near the corotation radius of a spiral wave.
+For the lowest-order modes in our disk, this is approximately centered at $\sim 25$ AU.
 
-In order to predict an $\alpha_{\rm{eff}}$, \citeauthor{gammie2001}'s formula, presented in equation \eqref{eq:gammie}, also requires knowledge of the cooling rate.
+In order to predict an $\alpha_{\rm{eff}}$, \citeauthor{gammie2001}'s formula, presented in equation \eqref{eq:gammie}, also requires knowledge of the cooling time.
 This is generally not known {\it a priori} for a given disk.
-Moreover, the cooling rate may not be easily parametrized with a local or global prescription over the entire disk.
-Simulations with realistic radiative physics \citep{boley2006} have shown that cooling rates can vary substantially as a function of both radius and time in a GI active disk.
+Moreover, the cooling time may not be easily parametrized with a local or global prescription over the entire disk.
+Simulations with realistic radiative physics \citep{boley2006} have shown that cooling times can vary substantially as a function of both radius and time in a GI active disk.
 In fact, it is these areas where the cooling rates change, leading to more or less GI activity, where many interesting phenomena may occur.
-
-In addition, at any given radius, 
+In addition, at any given radius, the
 stresses, and hence $\alpha_{\rm{eff}}$, are highly variable with time.
 Figure \ref{fig:alphavar} shows the time-variability of $\alpha_{\rm{eff}}$, 
-normalized to its average value between at $t = 14$ and 18 ORP, at seven radial locations.   
-Also shown is the orbital period at each radius, provided by the length of the line at the end of each subplot. 
+normalized to its average value between $t = 14$ and 18 ORP, at seven radial locations
+for times later than 12 ORP.
+Here the time axis is normalized to the orbital time.  Because the orbital periods are nearly
+Keplerian, at smaller radii the disk is followed for many more dynamical times than at larger
+radii.  For ease of comparison, 
+only the last 14 orbit are shown for $r = 10$ and 15.
 Fluctuations in $\alpha_{\rm{eff}}$
 on the order of 100\% or more are seen at all radii
-on  time scales range from fractions to integral multiples of the dynamic (orbital) time.  The temporal averaging
+on  timescales characteristic of the local dynamical time, indicating that
+this variability about the average is local in nature.  The temporal averaging
 of figure \ref{fig:alpha_v_radius} washes out this local variability.
@@
 problem under consideration and similar problems. The $l_{max} = 512$ simulation yields a value 
-for $\alpha_{avg}$ of 0.020. This overestimates the converged value of $\alpha_{avg}
+for $\alpha_{avg}$ of 0.020. This overestimates the extrapolated converged value of $\alpha_{avg}
 = 0.017$ by $\sim 18$\%. The $l_{max} = 512$ simulation also overestimates the fractional
-power in the high-order ($m = 2$ -- 7) modes by essentially the same amount.
+power in the high-order ($m = 2$ -- 7) modes by essentially the same percentage.
 However, it should be noted that the fractional difference between successive doubling of the resolution has decreased for $l_{max} = 512$.
@@
 Furthermore, the converged value represents infinite azimuthal resolution, which is, of course, impossible to achieve in a fixed grid scheme.
-It is unclear as to what variation of $\alpha$ between two values of $l_{max}$ would represent convergence as $\alpha$ itself is a dynamically varying quantity, see figure \ref{fig:alphavar}.
+%It is unclear as to what variation of $\alpha$ between two values of $l_{max}$ would represent convergence as $\alpha$ 
+%itself is a dynamically varying quantity, see figure \ref{fig:alphavar}.
 
 Another consideration is the effect of $r$ and $z$ resolutions on the outcome of the simulation.
-In this study we have chosen to focus on azimuthal resolution because it has the greatest effect on resolving the GI structures responsible for producing the gravitational stresses.
-However, as one increases the azimuthal resolution the radius at which the cells have roughly equal sides increases as well.
+In this study we have chosen to focus on azimuthal resolution because it has the greatest effect on resolving the non-axisymmetric
+structures responsible for producing the gravitational stresses.
+However, as one increases the azimuthal resolution, the radius at which the cells have roughly equal sides increases as well.
 Ideally, this should occur near the radial range where GIs are most active.
-For $l_{max} = 512$, $\delta r = \delta \phi r$ at $\approx 13.5$ AU, which is near the radius where GIs are active.
+For $l_{max} = 512$, $\delta r = r \delta \phi $ at $\approx 13.5$ AU, which is near the radius where GIs are active.
 
-Both the gravitational torques and effective alphas are functions of the
+Both the gravitational torques and $\alpha_{\rm{eff}}$ are functions of the
 gravitational stresses within the disk.  If the Reynolds stresses are, in fact, negligible relative to the gravitational stresses,
-then the errors in the calculated torques and effective alphas that stem from limited azimuthal resolution 
+then the percentage errors in the calculated torques and $\alpha_{\rm{eff}}$ that stem from limited azimuthal resolution 
 should be the same. In this case, the  $l_{max} = 512$ simulation should also be expected to yield
@@
 with higher resolution. 
+Although computationally expensive, simulations with even higher resolution in all directions and codes using different grid geometry
+would be worthwhile.
 
-It is also worth noting that the overestimation of the effective $\alpha$ is more than likely dependent on the total strength of non-axisymmetric amplitude $\langle A_{tot} \rangle$, which may be smaller depending on the physics of a simulation.
-For example, \citet{boley2006} found weaker overall GI activity when considering realistic radiative physics.
-A similar study of the effect of azimuthal resolution in disks cooled with a realistic scheme will be presented in a follow-up paper.
-
-{\bf Substantially changed this section, might still need some work. SM}
-
-\subsection{Fragmentation and Effective Alphas}
+\subsection{Fragmentation and Effective $\alpha$s}
 
 \noindent
-{\bf (Aaron, this section is for you to write or toss, your decision.  The current text should be thought of as a filler only)}
 
-%In an $\alpha$ disk formalism the mechanism giving rise to the transport of angular momentum also acts to dissipate energy.
-Lodato \& Rice (2005) postulated that a disk in a state of gravitoturbulence would have locally balanced heating and cooling rates.
-If the heating is due to local dissipation, then $\alpha$ and $t_{cool}$ are related by
-%\begin{equation}
-%\alpha = \left| \frac{{\rm d}\ln \Omega}{{\rm d} \ln \varpi} \right|^{-2}
-%\frac{1}{\gamma(\gamma - 1) t_{cool}\Omega}.
-%\end{equation}
-%\citet{rice2005} concluded that for a given $t_{cool}$  there exists some maximum stress, quantified by $\alpha_{max}$, which if exceeded would cause a disk to fragment.
-Their predicted value $\alpha_{max} \approx 0.06$ is shown as a horizontal dotted line in figure \ref{fig:alpha_v_radius}.
-Note that both the $l_{max} = $ 64 and 128 simulations exceed this limiting value over some radial range, yet neither displayed fragmentation.
-For cases where $t_{cool}$  varies with time, \citet{clarke2007} revised this limit to $\alpha_{max} = 0.12$.
+Under certain conditions, GIs fail to self-regulate, allowing density perturbations to grow until clumps form from spiral structure (see Durisen et al.
+2007 for references).
+Under isothermal conditions, fragmentation becomes very likely whenever $Q$ drops below 1.4
+\citep{tomley1991, tomley1994, nelson1998, mayer2002}.
+When the effective adiabatic index is stiff due to inefficient cooling, shocks and mass transport can militate against fragmentation, even in disks with initially low $Q$ values \citep[e.g., ][]{boley2008}.
+To understand the degree of cooling required to effect clump formation in non-isothermal disks, \cite{gammie2001} studied the nonlinear behavior of GIs using a series of shearing sheet simulations.
+He showed that  a gravitationally unstable disk is likely to fragment whenever the cooling time is less than half the local orbital period.
+This work was followed by Rice et al.~(2005), who used global 3D SPH simulations and found that the fragmentation limit was dependent on the adiabatic index.
+For disks with $\gamma=5/3$, they found fragmentation
+occurred whenever the local cooling time was less than the local orbital period.
+For $\gamma=7/5$, the limit was twice the local orbital period.
+\cite{rice2005} proposed that this fragmentation threshold was due to a maximum stress that can  be exerted on a disk by gravitational torques.
+Using the $\alpha_{\rm{eff}}$ as defined in equation (8), they found that fragmentation could occur whenever $\alpha_{\rm eff}\gtrsim 0.06$ (the dotted line in figure 6).
+The difference between the cooling time threshold found by Gammie and the ones found by Rice et al.~are not obviously at odds, as Gammie's study was conducted in 2D and used a 2D adiabatic index of 2.
+
+Recently, these results have been called into question by \cite{meru2011a}, who suggest that our understanding of the fragmentation criterion is based on simulations that have yet to converge.
+In particular, they find that fragmentation becomes increasingly likely for longer cooling times when the resolution in their 3D SPH simulations is increased, giving weight to the claims by, e.g., \cite{boss2009}, that fragmentation is a robust result of disk instability.
+The Meru \& Bate study uses a prescribed cooling such that $t_{\rm cool}=\beta\Omega^{-1}$, where $\Omega$ is the local orbital frequency and $\beta$ is some global constant.
+This same cooling prescription was used in the Gammie and Rice et al.~studies.
+The reason for the differences is unclear, and may not be simply a resolution effect (see Lodato \& Clarke 2011, arXiv:1101.2448).
+For example, the \cite{gammie2001} simulations, which are at very high resolution due to the local limit, require the most stringent cooling of all the simulations.
+Moreover, \cite{mejia2005}, who used grid-based simulations with the same code we use here, and  \cite{boley2008}, who used radiative hydrodynamics simulations, also found results that are reasonably consistent with those of Gammie, Rice et al., and \cite{cossins2010}.
+
+The present convergence study does not address fragmentation directly, but there are several results that should be noted.
+First, although the total power in non-axisymmetric structure is roughly constant throughout the simulations, the average $Q$ value increases with increasing resolution (Table 1).
+Because both fast cooling and a low $Q$ are necessary
+for fragmentation, this suggests that the disk is becoming less susceptible to clump formation with increasing resolution.
+Second, the disks do not show a trend toward denser spiral structure as the resolution is increased, which is seen by eye in figure 2, and is quantified by the decrease in $\left<A_{2-7}\right>$.
+The high-$m$ modes are increasing in power, but remain smaller than the low-$m$ components.
+Third, our results show that as the resolution increases, the disk approaches a self-regulating state that is reasonably well-described, in the time- and space-average sense, by an $\alpha$-disk with the parameterization of \cite{gammie2001}.
+
+Some caution must be taken when applying our results generally.
+The structure of the disk at the time of the onset of GIs may have a large impact on fragmentation.
+For example, \cite{boley2009} found that for disks with the same mass and {\it total} mass accretion rate, the disk with the shallower density profile fragmented, while the other did not \citep[see also ][]{meru2011b}.
+The disks explored here evolve to a surface density profile of approximately $p=5/2$ in the outer regions for $\Sigma\propto r^{-p}$, while those explored by \citet{meru2011a} have a shallower profile ($p=1$), favoring fragmentation.
+In addition to disk structure, the thermal history of the disk also plays a role in the fragmentation threshold.
+\cite{clarke2007}  found that the fragmentation cooling limit is sensitive to the way the instability is approached.
+When the cooling time is decreased slowly, the fragmentation limit is pushed toward shorter cooling times than if a fast cooling time is suddenly prescribed.
+
+Overall, the disks simulated here are converging toward a state of self-regulation as the resolution is increased, and do not show signs of becoming more susceptible to fragmentation.
 
@@
 
-\noindent
-{\bf (All: My style of writing papers is to write the summary after everything is completed.  Therefore, I have left this
-section blank.)}
+We conducted a convergence study of a protostellar disk subject to gravitational instabilities to
+examine the distribution and amplitudes 
+of non-axisymmetric density structures arising from GIs 
+and the connection of these structures to gravitational torques,
+to the level which transport can be represented by a simple $\alpha$-disk formulation,
+and to fragmentation.
+A 3D hydrodynamics code with self-gravity
+was used to follow a disk subject to a constant global cooling time during its asymptotic state, when
+heating and cooling are roughly in balance.
+Four simulations were conducted, identical except for the number of azimuthal grid 
+points, $l_{max} = $ 64, 128, 256, and 512, used in the calculations.
+
+A Fourier decomposition of the non-axisymmetric density structure 
+in $\cos(m\phi)$, $\sin(m\phi)$ was performed in order to
+characterize the amplitude distribution,  $A_m$,
+of azimuthal structures as a function of $m$.  
+This and previous studies have shown
+that disks in the asymptotic state have power at all resolvable $m$-values.
+The maximum value of  $m$ is given by
+$m_{max} = l_{max}/2$.  Therefore simulations with larger $l_{max}$ can spread the
+power over more Fourier components.
+We find that $\langle A_{tot} \rangle$, the power summed over all $m$, is roughly the same
+for all resolutions.
+However, the distribution of power among the Fourier components
+is distinctly affected by $l_{max} $.  
+% $A_7$ is approximately equal for all resolutions.
+The fractional
+amplitude in low-order modes, $\langle A_{2-7} \rangle / \langle A_{tot} \rangle$,
+where $\langle A_{2-7} \rangle$ is 
+the sum over modes $m = 2$ through 7, 
+decreases monotonically with increasing resolution.  Over three quarters of $\langle A_{tot} \rangle$ resides
+in $m =$ 2 -- 7 for the $l_{max} = 64$ simulation, while approximately half resides in these $m$-values when $l_{max} =$ 512.
+A Richardson extrapolation suggests a limiting value of $\langle A_{2-7} \rangle / \langle A_{tot} \rangle \sim 0.43$ 
+as $l_{max}\rightarrow \infty$.
+The number of degrees of freedom available for power, $l_{max}/2$, increases linearly with the resolution.  Thus
+Fourier amplitudes are naturally spread from lower-order to higher-order as the available $m$-values increase.
+
+Non-axisymmetric density structures produce gravitational torques, driving radial flows of mass and angular momentum.
+While the total torque is dominated by low-order Fourier components for all azimuthal resolutions, the total
+time-averaged gravitational torque  is smaller for higher resolution simulations.
+This reflects the fact that while a higher fraction of the total power in larger $l_{max}$ simulations is in higher-order Fourier components, these components make only minimal contributions to the gravitational torque.
+As $l_{max}$ increases, non-axisymmetric amplitude is shifted from low-order global components, which produce relatively large gravitational torques, to high-order local structures, which tend to cancel each other when integrated over the whole disk.
+
+The effective $\alpha_{\rm{eff}}$ was determined for each simulation under the assumption that
+Reynolds stresses are negligible relative to the gravitational stresses, as found in
+previous 3D studies.  
+As is the case for gravitational torques, $\alpha_{\rm{eff}}$  decreases with increased resolution.  
+Derived values of 
+$\alpha_{\rm{eff}}$ vs. radius for the highest resolution simulation, $l_{max} = 512$, 
+are consistent with
+predictions from an analytic local theory for thin disks by \citet{gammie2001}, as is the
+converged $\alpha_{\rm{eff}}= 0.017$, time-averaged over the asymptotic phase and spatially averaged 
+over the disk. The time and spatially averaged $\alpha_{\rm{eff}}$  
+obtained for the $l_{max} = 512$ simulations overestimates the converged value
+by less than 20\%.  Gravitational torques for this same model differ with converged values by a
+similar amount.
+Locally determined values of
+$\alpha_{\rm{eff}}$ vary by factors of 2 or 3 throughout the disk on timescales characteristic
+of the local dynamical time.  Thus the predictions of Gammie are only valid over many dynamic times when averaged over a substantial volume of the disk.
+
+Although not a goal of this work, our simulations permit
+comments on studies on fragmentation and clump formation.
+Most recent studies indicate that our disk should not fragment \citep[e.g., ][]{gammie2001, rice2005, mejia2005, boley2008}.   However, some recent studies \citep[e.g., ][]{boss2009, meru2011a}
+%Meru & Bate (2011), 
+suggest that fragmentation may be a common result of gravitational instabilities in disks.
+We find that as the resolution increases, our disk approaches a self-regulating state that can be reasonably well-described, in the time- and space-average sense, by an $\alpha$-disk parameterization.
+The average Toomre $Q$ increases with increasing resolution,
+suggesting that the disk is becoming less susceptible to clump formation with increasing resolution,
+consistent with the observation that
+the disk simulations do not show a trend toward denser spiral structure as the resolution is increased.
+Overall, the disks simulated here are converging toward a state of self-regulation as the resolution is increased, and do not show signs of becoming more susceptible to fragmentation.
+
+A future paper will address the issues presented in this work using a convergence study of GI-susceptible disks subject to radiative cooling with realistic opacities.
+This will allow us to address the generality of the results cited herein.
 
 \acknowledgements
-Please add any acknowledgements here.
+
+SM would like to thank NASA for the generous funding of the
+NASA Earth and Space Science Fellowship NNX07AU82H. This
+research was also supported in part by NASA Origins of Solar
+Systems grants NNG05GN11G and NNX08AK36G. The simulations
+and analysis herein were conducted on hardware generously
+provided by the Indiana University Information Technology Services
+and was supported by the National Science Foundation under
+Grant no. ACI-0338618l, OCI-0451237, OCI-0535258, CNS-
+0521433 and OCI-0504075. This work was also supported in part
+by Shared University Research grants from IBM, Inc., to the Indiana
+University.
 
@@
 
-\begin{deluxetable}{cccccc}
-\tablecolumns{5} 
-\tablewidth{0pc} 
-\tablecaption{$Q$, Fourier Amplitudes, and $\alpha$} 
-\tablehead{ 
-   \colhead{$l_{max}$} 
-& \colhead{$Q_{avg}$}   
-& \colhead{$\langle A_{tot} \rangle$}    
-& \colhead{$\langle A_{2-7} \rangle / \langle A_{tot} \rangle$} 
-& \colhead{$\alpha_{avg}$} \\ 
-} 
-\startdata
+\begin{table}
+\begin{center}
+\begin{tabular}{ccccc}
+\multicolumn{5}{c}{$Q$, Fourier Amplitudes, and $\alpha$}\\ \hline\hline
+$l_{max}$ & $Q_{avg}$   & $\langle A_{tot} \rangle$   & $\langle A_{2-7} \rangle / \langle A_{tot} \rangle$ & $\alpha_{avg}$ \\\hline
 64   & 1.26 & 1.96 &  0.76 & 0.069 \\
@@
 512 & 1.48 & 1.95 &  0.49 & 0.020 \\ 
-\enddata 
-\caption{
-All quantities are time-averaged over 12 -- 19 ORP.
+$\infty$ & & &0.43 & 0.017 \\ \hline
+\label{tbl:ams}
+\end{tabular}
+\caption{ All quantities are time-averaged over 12 -- 18 ORP.
 $Q_{avg}$ and $\alpha_{avg}$ are also spatially
-averaged over 10 -- 40 AU.
-}
-\label{tbl:ams}
-\end{deluxetable}
+averaged over 10 -- 40 AU.  The entries for ``$\infty$'' are based on Richardson extrapolation,
+as explained in the text.}  
+\end{center}
+\end{table}
+
+%\begin{deluxetable}{cccccc}
+ %\tablecolumns{5}
+%\tablewidth{0pc} 
+%\tablecaption{$Q$, Fourier Amplitudes, and $\alpha$} 
+%\tablehead{ 
+ %  \colhead{$l_{max}$} 
+%& \colhead{$Q_{avg}$}   
+%& \colhead{$\langle A_{tot} \rangle$}    
+%& \colhead{$\langle A_{2-7} \rangle / \langle A_{tot} \rangle$} 
+%& \colhead{$\alpha_{avg}$} \\ 
+%}
+
+%\startdata
+%64   & 1.26 & 1.96 &  0.76 & 0.069 \\
+%128 & 1.24 & 1.91 &  0.62 & 0.044 \\
+%256 & 1.39 & 1.83 &  0.54 & 0.026 \\
+%512 & 1.48 & 1.95 &  0.49 & 0.020 \\
+%\enddata  
+%\label{tbl:ams}
+%\caption{
+%All quantities are time-averaged over 12 -- 19 ORP.
+%$Q_{avg}$ and $\alpha_{avg}$ are also spatially
+%averaged over 10 -- 40 AU.
+%}
+%\end{deluxetable}
+
 \newpage
@@
 \begin{figure}
-\plotone{figures/Figure1.eps}
+\plotone{Figure1.eps}
 \caption{Azimuthally averaged mid-plane Toomre Q distribution versus radius for each of the four resolutions at t $\sim$ 18 ORPs}
@@
 \begin{figure}
-\plotone{figures/Figure2.eps}
+\plotone{Figure2.eps}
 \caption
@@
 \begin{figure}
-\plotone{figures/Figure3.eps}
+\plotone{Figure3.eps}
 \caption
@@
 The logarithm of the amplitudes for the global Fourier components of the non-axisymmetric density structure plotted for different $m$-values.
-Notice that the components with the highest global amplitudes are those with low $m$-values.
+Notice that the components with the highest global amplitudes are those with low $m$-values; $m=2$ through 6 are labeled.
+The vertical dotted line corresponds with $m=7$.
+For modes with $m < 7$,  the radial range from  the inner Linblad resonance to the outer Linblad resonance exceeds the disk scale height; these quantities are roughly equal for $m = 7$.
 }
@@
 \begin{figure}
-\plotone{figures/Figure4.eps}
+\plotone{Figure4.eps}
 \caption
 {
-Convergence of fractional power.  The $x$-axis represents
+Convergence of the fractional amplitude of low-order non-axisymmetric structure.  The $x$-axis represents
 the azimuthal ``spacing'' between grid points,  normalized such that spacing of the $l_{max}$ = 512 grid
@@
 \begin{figure}
-\plotone{figures/Figure5.eps}
+\plotone{Figure5.eps}
 \caption
 {
-Torque Array}
+Total time-averaged gravitational  torques, along with contributions arising from torques
+summed over combinations of  $m = $ 1, 2, 3, 4, and 5,
+as a function of radius and azimuthal resolution. 
+Higher resolution simulations are subject to smaller total torques while the total torque is dominated by low-order $m$-value
+components for all azimuthal resolutions.
+While a higher fraction of the total power in larger $l_{max}$ simulations is in higher-order $m$-values, these components make only minimal contributions to the gravitational torque.
+}
 \label{fig:torquearray}
@@
 \begin{figure}
-\label{fig:alpha_v_radius}
-\plotone{figures/Figure6.eps}
+\plotone{Figure6.eps}
 \caption
@@
 2005).
-{\bf suggest removing the Rice line if it is not discussed in the paper.}
 }
+\label{fig:alpha_v_radius}
 \end{figure}
@@
 \begin{figure}
-\plotone{figures/Figure7.eps}
-\caption
-{Alpha variations.}
+\plotone{Figure7.eps}
+\caption{
+Local value of alpha at seven radii shown as a function of the local dynamical time.  For radii greater than 20 AU, the plot shows 
+the variation from 12 to 18 ORP.  At 10 and 15 AU, the 14 orbit periods immediately prior to 18 ORP are shown.
+Dotted lines at an ordinate value of one  represent the averages from 14 to 18 ORP at that radius.
+Note that the local $\alpha$ displays $\sim$ 100\% variations on the local dynamical time. 
+}
 \label{fig:alphavar}

### 2011-07-26 Paper_1.tex
Added Dick's changes from revision one.
	modified:   Paper_1.tex
	modified:   general.bib

@@
 
-Numerous studies have shown that thermal physics control GI amplitudes by a {\it balance} of heating and cooling \citep[e.g.,][]{tomley1991,tomley1994,pickett1998,pickett2000, pickett2003,gammie2001,boss2002,rice2003b,mejia2005,boley2006,boley2007, stamatellos2008,cossins2009}.
-Heating by GIs tends to increase $c_s$, thus increasing $Q$.
-If disk heating persists, the instability will be suppressed once $Q$ becomes sufficiently large.
-However, heating and cooling can reach a balance at nonlinear wave amplitude.
-GI-activity can sustain this balance  at a relatively constant,
+Numerous studies have shown that thermal physics control GI amplitudes by a {\it balance} of heating and cooling \citep[e.g.,][]{tomley1991,tomley1994,pickett1998,pickett2000, pickett2003,gammie2001,boss2002,rice2003b,mejia2005,boley2006,boley2007b, stamatellos2008,cossins2009}.
+Heating by GIs tends to increase $c_s$, thus increasing $Q$. If disk heating persists, the instability will be
+suppressed once $Q$ becomes sufficiently large.  On the other hand, radiative cooling tends to oppose the increase in
+$Q$ by lowering $c_s$. In this way, heating and cooling can reach a rough overall balance at nonlinear wave amplitude. GI-activity can sustain this balance  at a relatively constant,
 but unstable, value of $Q$ \citep{paczynski1978, lin1981,goldreich1965}, and 
@@
 \citet{shakura1973} proposed turbulence as the primary source of effective viscosity in gas disks
-and parameterized turbulent transport in a steady-state disk with the  parameter $\alpha$.  In this case,  the kinematic
+and parameterized turbulent transport in a steady-state disk with the  parameter $\alpha$ \citep[see also][]{pringle1981}.  In this case,  the kinematic
 viscosity is defined by $\nu = \alpha c_s H$, where  $c_s$ is the sound speed and $H$ is the disk height.
@@
 framework.  However, the  validity of an $\alpha$-disk picture requires self-consistency of local energy dissipation. 
-The locality of dissipation in GI-active disks has not been rigorously tested and remains an open question.
+The locality of dissipation in GI-active disks has not been thoroughly tested and remains an open question \citep{lodato2004,lodato2005,boley2006}.
 
@@
 This almost certainly applies to the embedded phase of both high and low-mass stars when infall from the collapsing protostellar cloud feeds mass into the disk at a rapid
-rate \citep[e.g.,][]{yorke1993, laughlin1994, yorke1999,vorobyov2005,vorobyov2006}.  
+rate \citep[e.g.,][]{yorke1993, laughlin1994, yorke1999,vorobyov2005,vorobyov2006,kratter2010}.  
 However, despite the long-distance nature of gravitational torques, theoretical and numerical results suggest that in protostellar disks where heating and cooling are in rough balance, 
@@
 $\epsilon$ by $d\epsilon/dt = \epsilon / t_{cool}$.  In contrast, we and our collaborators use either a {\it global cooling time},
-$t_{cool} = $ constant \citep[e.g.,][this paper]{pickett2003, mejia2005}, or use realistic opacities \citep[e.g.,][]{mejiaphd2004, cai2006, cai2008, boley2006, boley2007,  boley2009} which lead to something in between $t_{cool} = $ constant  and
-$t_{cool}\Omega =$ constant.
+$t_{cool} = $ constant \citep[e.g.,][this paper]{pickett2003, mejia2005}, or use realistic opacities \citep[e.g.,][]{mejiaphd2004, cai2006, cai2008, boley2006, boley2007b,  boley2009} which lead to something in between $t_{cool} = $ constant  and
+$t_{cool}\Omega =$ constant. The review by \citet{durisen2007} discusses some consequences of these various
+choices. There is a tendency for $t_{cool}\Omega =$ constant to bias a disk toward behavior like that of a
+steady-state $\alpha$-disk.
 
-\citet{gammie2001} derived an effective alpha for the case of a thin gravitationally unstable disk where the heating caused by GIs was balanced by the local cooling rate.  
+\citet{gammie2001} derived an effective $\alpha$ for the case of a thin gravitationally unstable disk where the heating
+caused by GIs was balanced by the local cooling rate \citep[see also][]{pringle1981}.  
 Assuming a local cooling time of $t_{cool}\Omega =$ constant,  Gammie found
@@
 Fully 3D simulations \citep[e.g.,][]{lodato2004, cossins2009} suggest that, at least in the case of imposed 
-$t_{cool}\Omega = $ constant cooling, $\alpha_{\rm{\rm{eff}}}$ roughly converges to the value predicted by \citeauthor{gammie2001}.
+$t_{cool}\Omega = $ constant cooling, $\alpha_{\rm{\rm{eff}}}$ roughly converges to the value given by equation (\ref{eq:gammie}).
 
@@
 
-The balance of this paper is organized as follows.
+The remainder of this paper is organized as follows.
 Section 2 provides the details of the numerical approach and defines the models.
@@
 
-Because this is a polytropic disk with an idealized cooling time, only the ratio of disk to stellar mass, given here by 0.153, is specified.
-In what follows, quantities specified with real physical units are derived assuming a 1 $M_\odot$ central star and a 0.153  $M_\odot$ disk.
-In this case, the outer rotation period (ORP) is defined at $r = 33$ AU and one ORP $\approx$ 214 years.
-The disk has an initial surface density that approximates $\Sigma \sim r^{ -1}$, with initial inner and outer radii of 2.3 and 40 AU, respectively.
+Because this is a polytropic disk with an idealized cooling time, only the ratio of disk to stellar mass, given here by
+0.153, is specified.  In what follows, quantities specified with real physical units are derived assuming a 1 $M_\odot$
+central star and a 0.153  $M_\odot$ disk. In this case, the outer rotation period (ORP) is defined at $r = 33$ AU and
+one ORP $\approx$ 214 years.  The disk has an initial surface density $\Sigma \sim r^{ -1}$, with initial inner and
+outer radii of 2.3 and 40 AU, respectively. The construction of these initial equilibrium disk models is discussed in
+\citet{pickett2003} and \citet{mejia2005}.
 
- We assume an equation of state $P = (\gamma -1)\epsilon$, where $P$ is the pressure, $\epsilon$ is the internal energy density, and $\gamma$ is the ratio of specific heats, given here by $\gamma = 5/3$.
+ We assume an equation of state $P = (\gamma -1)\epsilon$, where $P$ is the pressure, $\epsilon$ is the internal energy
+ density, and $\gamma$ is the ratio of specific heats, given here by $\gamma = 5/3$. As discussed in \citet{boley2007a},
+ the rotation states of $H_2$ are probably not excited in the cold outer parts of the disk ($\gtr 10$ AU); so $\gamma =
+ 5/3$ is an appropriate choice.
 The model disks are cooled by decreasing their internal energy according to the prescription $d\epsilon/dt = \epsilon / t_{cool}$, where $t_{cool}$ is the global cooling time, which is set to 2 ORP everywhere.
@@
 \citet{michael2010} found that freeing the star changes the torques by about 10-20\%, 
-thus the difference between torques arising from the $m=1$ mode with the star fixed and free to move are small compared to the differences found when comparing azimuthal resolution.
+thus the difference between torques arising from the $m=1$ mode with the star fixed and free to move are small compared
+to the differences found when comparing azimuthal resolution. In fact, it is expected that the $m =1$ mode would only be
+effective at higher disk-to-star mass ratios \citep{shu1990}.
 
@@
 
-In order to predict an $\alpha_{\rm{eff}}$, \citeauthor{gammie2001}'s formula, presented in equation \eqref{eq:gammie}, also requires knowledge of the cooling time.
+In order to compare an $\alpha_{\rm{eff}}$, \citeauthor{gammie2001}'s formula, presented in equation \eqref{eq:gammie}, also requires knowledge of the cooling time.
 This is generally not known {\it a priori} for a given disk.
@@
 radii.  For ease of comparison, 
-only the last 14 orbit are shown for $r = 10$ and 15.
+only the last 14 orbits are shown for $r = 10$ and 15.
 Fluctuations in $\alpha_{\rm{eff}}$
@@
 Thus,
-even in this highly GI-active region, the predictions of Gammie are only valid over many dynamic times 
+even in this highly GI-active region, Gammie's formula only seems valid over many dynamic times 
 when averaged over a substantial volume of the disk.
@@
 Under isothermal conditions, fragmentation becomes very likely whenever $Q$ drops below 1.4
-\citep{tomley1991, tomley1994, nelson1998, mayer2002}.
-When the effective adiabatic index is stiff due to inefficient cooling, shocks and mass transport can militate against fragmentation, even in disks with initially low $Q$ values \citep[e.g., ][]{boley2008}.
-To understand the degree of cooling required to effect clump formation in non-isothermal disks, \cite{gammie2001} studied the nonlinear behavior of GIs using a series of shearing sheet simulations.
-He showed that  a gravitationally unstable disk is likely to fragment whenever the cooling time is less than half the local orbital period.
+\citep{tomley1991, tomley1994, nelson1998, mayer2002}. When the effective adiabatic index is stiff due to inefficient
+cooling, shocks and mass transport can militate against fragmentation, even in disks with initially low $Q$ values
+\citep[e.g., ][]{boley2008}. To understand the degree of cooling required to effect clump formation in non-isothermal
+disks, \cite{gammie2001} studied the nonlinear behavior of GIs using a series of shearing sheet simulations.  He showed
+that  a gravitationally unstable disk is likely to fragment whenever the cooling time is less than half the local
+orbital period ($t_{cool}\Omega = 3$).
 This work was followed by Rice et al.~(2005), who used global 3D SPH simulations and found that the fragmentation limit was dependent on the adiabatic index.
@@
 
-Recently, these results have been called into question by \cite{meru2011a}, who suggest that our understanding of the fragmentation criterion is based on simulations that have yet to converge.
-In particular, they find that fragmentation becomes increasingly likely for longer cooling times when the resolution in their 3D SPH simulations is increased, giving weight to the claims by, e.g., \cite{boss2009}, that fragmentation is a robust result of disk instability.
-The Meru \& Bate study uses a prescribed cooling such that $t_{\rm cool}=\beta\Omega^{-1}$, where $\Omega$ is the local orbital frequency and $\beta$ is some global constant.
-This same cooling prescription was used in the Gammie and Rice et al.~studies.
-The reason for the differences is unclear, and may not be simply a resolution effect (see Lodato \& Clarke 2011, arXiv:1101.2448).
-For example, the \cite{gammie2001} simulations, which are at very high resolution due to the local limit, require the most stringent cooling of all the simulations.
+Recently, these results have been called into question by \cite{meru2011a}, who suggest that our understanding of the
+fragmentation criterion is based on simulations that have yet to converge.  In particular, they find that fragmentation
+becomes increasingly likely for longer cooling times when the resolution in their 3D SPH simulations is increased,
+giving weight to the claims by, e.g., \cite{boss2009}, that fragmentation is a robust result of disk instability. The
+Meru \& Bate study uses a prescribed cooling such that $t_{\rm cool}=\beta\Omega^{-1}$, where $\Omega$ is the local
+orbital frequency and $\beta$ is some global constant.  This same cooling prescription was used in the Gammie and Rice
+et al.~studies.  The reason for the differences is unclear, and may not be simply a resolution effect (see Lodato \&
+Clarke 2011, arXiv:1101.2448).  For example, the \cite{gammie2001} have revealed the most stringent cooling criterion of
+all the $t_{cool}$ fragmentation studies, and, due to the local approximation Gammie used, these simulations are also at
+very high resolution.
 Moreover, \cite{mejia2005}, who used grid-based simulations with the same code we use here, and  \cite{boley2008}, who used radiative hydrodynamics simulations, also found results that are reasonably consistent with those of Gammie, Rice et al., and \cite{cossins2010}.
@@
 for fragmentation, this suggests that the disk is becoming less susceptible to clump formation with increasing resolution.
-Second, the disks do not show a trend toward denser spiral structure as the resolution is increased, which is seen by eye in figure 2, and is quantified by the decrease in $\left<A_{2-7}\right>$.
+Second, the disks do not show a trend toward denser spiral structure as the resolution is increased which is seen by eye in figure 2, and is quantified by the decrease in $\left<A_{2-7}\right>$.
 The high-$m$ modes are increasing in power, but remain smaller than the low-$m$ components.
@@
 {
-The logarithm of the amplitudes for the global Fourier components of the non-axisymmetric density structure plotted for different $m$-values.
-Notice that the components with the highest global amplitudes are those with low $m$-values; $m=2$ through 6 are labeled.
-The vertical dotted line corresponds with $m=7$.
-For modes with $m < 7$,  the radial range from  the inner Linblad resonance to the outer Linblad resonance exceeds the disk scale height; these quantities are roughly equal for $m = 7$.
+The logarithm of the amplitudes for the global Fourier components of the non-axisymmetric density structure plotted for
+different $m$-values. Notice that the components with the highest global amplitudes are those with low $m$-values; $m=2$
+through 6 are labeled.  The vertical dotted line corresponds with $m=7$.  For modes with $m < 7$,  the radial range
+$\Delta r_{LR}$ from  the inner Linblad resonance to the outer Linblad resonance exceeds the disk scale height; the
+$\Delta r_{LR}$ and scale height are roughly equal for $m = 7$.
 }
@@
 \caption
-{Effective $\alpha$'s for the $t_{cool} =$ 2 ORP, $\gamma = 5/3$, $\Sigma \sim r^{ -1}$ disk, averaged between 12 and 18 ORPs, for four different different azimuthal grid resolutions, given by the number of azimuthal grid-points $l_{max}$.
-The dashed lines show the predictions of Gammie (2001) for $\alpha$ (upper: strongly self-gravitating limit, lower: non-self-gravitating limit), while the dashed line at $\log(\alpha) = -1.22$ represents the maximum stress a disk can maintain without fragmenting (Rice et al.
-2005).
+{Effective $\alpha$'s for the $t_{cool} =$ 2 ORP, $\gamma = 5/3$, $\Sigma \sim r^{ -1}$ disk, averaged between 12 and 18
+  ORPs, for four different azimuthal grid resolutions, given by the number of azimuthal grid-points $l_{max}$. The
+  dashed lines show the $\alpha$ from equation (\ref{eq:gammie}) (upper: strongly self-gravitating limit, lower:
+  non-self-gravitating limit), while the dotted line at $\log(\alpha) = -1.22$ represents the maximum stress a disk can
+  maintain without fragmenting according to \citet{rice2005}. 
 }
@@
 \caption{
-Local value of alpha at seven radii shown as a function of the local dynamical time.  For radii greater than 20 AU, the plot shows 
+Local value of $\alpha$ at seven radii shown as a function of the local dynamical time.  For radii greater than 20 AU, the plot shows 
 the variation from 12 to 18 ORP.  At 10 and 15 AU, the 14 orbit periods immediately prior to 18 ORP are shown.
-Dotted lines at an ordinate value of one  represent the averages from 14 to 18 ORP at that radius.
+Dotted lines at an ordinate value of unity represent the average from 14 to 18 ORP at that radius.
 Note that the local $\alpha$ displays $\sim$ 100\% variations on the local dynamical time. 

### 2011-08-30 Response_to_Referee1.txt
Dick's changes and draft response to referee.
	modified:   Paper_1.tex
	new file:   Response_to_Referee1.txt

This is the response to the first referee report for ApJ submission MS#ApJ83382

My primary concern with the analysis is the neglect of the Reynolds stress component when calculating the effective \alpha viscosity. The authors show that as the disk resolution is increased, more power is transferred into higher m-structures approaching the dissipation scale in the code. It seems likely that the shift also causes velocity correlations on this scale, which would produce extra angular momentum transport and thus a higher effective \alpha.

I do not understand the author's claim that it is too difficult or unreliable to measure Reynolds stresses. This paper would be greatly improved by showing how the Reynolds term evolves with resolution. In fact, I wonder whether all or some of the decrease in the transport at higher resolution would disappear if this latter term were included.

In addition to measuring the Reynolds stresses directly, the authors could also look at the mass accretion rate through rings in order to quantify the effective \alpha. The latter would be another way to get at the contribution from Reynolds stresses.

The authors could also measure the net heating in their simulation to quantify the effective \alpha via dissipation. This would give again more insight into the effect of resolution on the critical cooling time or the value of \alpha as a function of (Q). Even though the authors state that their results are converging and not fragmenting, a lower effective \alpha should imply less heating, and thus increase the critical cooling time. Of course, if \alpha is more and more dominated by Reynolds terms, than the heating rate and effective alpha might not decline. This is a very important distinction.  The comment in the summary that Q is increasing with increasing resolution suggests that there is more heating, and thus more dissipation.

All of my other concerns are minor, and listed in page order below:

-pg 4:  "However, heating and cooling can reach a balance at nonlinear..."
More explanation would be useful to improve the readability for non-experts

* Explanation added: In this way, heating and cooling can reach a rough overall balance at nonlinear wave amplitude. GI-activity can sustain this balance at a relatively constant, but unstable, value of Q (Paczynski 1978; Lin 1981; Goldreich & Lynden-Bell 1965), and the disk can exist in a state of quasi-equilibrium. In this state, cooling is balanced by heating arising from the inward flow of matter and shocks induced by the GIs (Gammie 2001; Lodato & Rice 2004; Rice et al. 2005; Boley et al. 2006; Cossins et al. 2009; Vorobyov 2010).

pg 5: Sunyaev is misspelled, and a citation should be given to Pringle 81 for the cited  \alpha  formalism
(SS73 is magnetic, and with a slightly different scaling)

*Spelling fixed, and Pringle cited

pg 6: A citation to Rice & Lodato 04-05 should be included as they tested the locality of transport and
dissipation.

*Citation included

pg 6: A citation to Kratter et al 2010, Walsh et al 2010 would also be appropriate for models of embedded GI
unstable accretion disks.

*Kratter citation included, however we were unable to locate the Walsh paper referenced here.

pg 7: I don't understand the emphasis on local vs "global" cooling time. I can see that a global cooling time
will make the entire disk go unstable at the same time (inner radii get to Q=1 faster), but it is no more
physical than the "local" choice, which allows for a cancellation in the energy equation and thus produces the
dimensionless criterion. The authors should provide some explanation of this issue, or of their choice. Of
course the difference between these two is irrelevant for Gammie's simulation.

*First paragraph on page 7 rewritten to attempt to elucidate this issue.

pg 8: Purely stylistic, but the phrase "the balance of this paper" see odd to me (rather than "the remainder
of this paper")

*Changed

pg 8: "alpha" should be a letter

*Changed

pg 9: "Source and flux terms..." The authors should explain what these are physically.

*Citations included which explain this in detail.

pg 9: What does is mean that the surface density approximates 1/r?

*Wording changed to be more clear.

pg 9: The authors use \gamma  = 5/3 for a protostellar disk, which should be \gamma=7/5. It would be useful to
note this as the critical cooling boundary is a function of \gamma.

*Discussion and citations to Boley et al. 2007, who cover this issue in great detail are included. In fact, it
 is not clear what the appropriate choice is when using a fixed \gamma.

pg 13: A footnote explaining Richardson extrapolation might be useful

*Citation to Numerical Recipes included.

pg 14: "only the last 14 orbit" -> orbits

*Changed

pg 15: Note that the effectiveness of the m=1 mode depends on the disk to star mass ratio. (cite Adams et al
1989)

*Citation to Shu et al. 1990 included.

pg 16: above eq 8: this seems like a lot of citation for one equation...

pg 16: The note that Gammie found that Reynolds stresses were important, might also be related to the higher resolution.

pg 20: It is not quite accurate to equate inefficient cooling with a stiff equation of state in the context of
these simulations. One can set the cooling time to be quite short, and \gamma to be quite stiff.

*What is written in the manuscript is correct. The key word is "effective." A short cooling time for a truly stiff cp/cv=gamma makes the effective gamma nonstiff.  The classic example is an isothermal gas.  We avoid possible confusion on this issue by not addressing this additional complexity.

pg 20: "militate against" Do you mean mitigate?

*No

pg 20:  It might be useful to include a footnote saying that you are using orbital periods instead of \Omega
to quote the critical values for the cooling time.

*Text modified as follows: "half the local orbital period (tcool~3/Omega)."

pg 21: The sentence starting "For example, the Gammie 2001 simulations..." is confusingly written. Also, I disagree with the description of the Meru and Bate results adding "weight to the claims by e.g. Boss..". It seems clear that unless dissipation is dramatically non-local, there must be a relation between the rate of dissipation and cooling. I don't think Boss is claiming that this doesn't exist, just that cooling isn't a problem.

*Modified as follows: "For example, the Gammie (2001) simulations have revealed the most stringent cooling
criterion of all the tcool-fragmentation studies, and due to the local approximation Gammie used, the
simulations are also at very high resolution."

Figures:

General comment: the numbers on the axes are somewhat small on all of the figures. Depending on how they are eventually sized in publication, I can imagine some of them being hard to read.

fig 3 caption: "these quantities are roughly..." I don't know what this refers to

*Rewritten for clarity.

fig 5: While I appreciate the black and white legibility of this diagram, color might be useful for online
readers (or those who print in color) to distinguish between the various curves

*Both color and B/W versions of the figures have been provided.

fig 6: "different different". The flat line looks dotted to me (not dashed). I also think it is misleading to
describe Gammie's as a prediction, and Rice et al 2005 as more of a "fact"

*This has been corrected

fig 7: "alpha" should be the letter. I find the caption description of the dotted lines confusing.

*This has been corrected

### 2011-08-30 Paper_1.tex
Dick's changes and draft response to referee.
	modified:   Paper_1.tex
	new file:   Response_to_Referee1.txt

@@
  density, and $\gamma$ is the ratio of specific heats, given here by $\gamma = 5/3$. As discussed in \citet{boley2007a},
- the rotation states of $H_2$ are probably not excited in the cold outer parts of the disk ($\gtr 10$ AU); so $\gamma =
+ the rotation states of $H_2$ are probably not excited in the cold outer parts of the disk ($> 10$ AU); so $\gamma =
  5/3$ is an appropriate choice.

### 2011-09-14 Response_to_Referee1.txt
Changed formatting for response to ref.
	modified:   Response_to_Referee1.txt
Edited Response_to_Referee1.txt via GitHub

@@
 
-My primary concern with the analysis is the neglect of the Reynolds stress component when calculating the effective \alpha viscosity.
-The authors show that as the disk resolution is increased, more power is transferred into higher m-structures approaching the dissipation scale in the code.
-It seems likely that the shift also causes velocity correlations on this scale, which would produce extra angular momentum transport and thus a higher effective \alpha.
+My primary concern with the analysis is the neglect of the Reynolds stress component when calculating the
+effective \alpha viscosity. The authors show that as the disk resolution is increased, more power is
+transferred into higher m-structures approaching the dissipation scale in the code. It seems likely that the
+shift also causes velocity correlations on this scale, which would produce extra angular momentum transport
+and thus a higher effective \alpha.
 
-I do not understand the author's claim that it is too difficult or unreliable to measure Reynolds stresses.
-This paper would be greatly improved by showing how the Reynolds term evolves with resolution.
-In fact, I wonder whether all or some of the decrease in the transport at higher resolution would disappear if this latter term were included.
+* It should be noted that the measurement presented in Table 1 and Figure 3 of the manuscript are those of the
+Fourier amplitudes, as defined in eqns 2 and 3. As can be seen by their definition, these values do not 
+directly measure the power in a given eignmode. In addtion, each particular m value makes a relatively minor 
+contribution for large m (i.e. m > 20). The measurement of A_m also includes no phase information, so A_m
+is *not* necessarily a good indicator of coherent density structures or organized velocity correlations.
 
-In addition to measuring the Reynolds stresses directly, the authors could also look at the mass accretion rate through rings in order to quantify the effective \alpha.
-The latter would be another way to get at the contribution from Reynolds stresses.
+I do not understand the author's claim that it is too difficult or unreliable to measure Reynolds
+stresses. This paper would be greatly improved by showing how the Reynolds term evolves with resolution. In
+fact, I wonder whether all or some of the decrease in the transport at higher resolution would disappear if
+this latter term were included.
 
-The authors could also measure the net heating in their simulation to quantify the effective \alpha via dissipation.
-This would give again more insight into the effect of resolution on the critical cooling time or the value of \alpha as a function of (Q).
-Even though the authors state that their results are converging and not fragmenting, a lower effective \alpha should imply less heating, and thus increase the critical cooling time.
-Of course, if \alpha is more and more dominated by Reynolds terms, than the heating rate and effective alpha might not decline.
-This is a very important distinction.
-The comment in the summary that Q is increasing with increasing resolution suggests that there is more heating, and thus more dissipation.
+In addition to measuring the Reynolds stresses directly, the authors could also look at the mass accretion
+rate through rings in order to quantify the effective \alpha. The latter would be another way to get at the
+contribution from Reynolds stresses.
+
+The authors could also measure the net heating in their simulation to quantify the effective \alpha via
+dissipation. This would give again more insight into the effect of resolution on the critical cooling time or
+the value of \alpha as a function of (Q). Even though the authors state that their results are converging and
+not fragmenting, a lower effective \alpha should imply less heating, and thus increase the critical cooling
+time. Of course, if \alpha is more and more dominated by Reynolds terms, than the heating rate and effective
+alpha might not decline. This is a very important distinction.  The comment in the summary that Q is
+increasing with increasing resolution suggests that there is more heating, and thus more dissipation.
 
@@
 
-* Explanation added: In this way, heating and cooling can reach a rough overall balance at nonlinear wave amplitude.
-GI-activity can sustain this balance at a relatively constant, but unstable, value of Q (Paczynski 1978; Lin 1981; Goldreich & Lynden-Bell 1965), and the disk can exist in a state of quasi-equilibrium.
-In this state, cooling is balanced by heating arising from the inward flow of matter and shocks induced by the GIs (Gammie 2001; Lodato & Rice 2004; Rice et al.
-2005; Boley et al.
-2006; Cossins et al.
-2009; Vorobyov 2010).
+* Explanation added: In this way, heating and cooling can reach a rough overall balance at nonlinear wave
+  amplitude. GI-activity can sustain this balance at a relatively constant, but unstable, value of Q
+  (Paczynski 1978; Lin 1981; Goldreich & Lynden-Bell 1965), and the disk can exist in a state of
+  quasi-equilibrium. In this state, cooling is balanced by heating arising from the inward flow of matter and
+  shocks induced by the GIs (Gammie 2001; Lodato & Rice 2004; Rice et al. 2005; Boley et al. 2006; Cossins et
+  al. 2009; Vorobyov 2010).
 
-pg 5: Sunyaev is misspelled, and a citation should be given to Pringle 81 for the cited  \alpha  formalism
-(SS73 is magnetic, and with a slightly different scaling)
+pg 5: Sunyaev is misspelled, and a citation should be given to Pringle 81 for the cited \alpha formalism (SS73
+is magnetic, and with a slightly different scaling)
 
@@
 
-*Discussion and citations to Boley et al. 2007, who cover this issue in great detail are included. In fact, it
- is not clear what the appropriate choice is when using a fixed \gamma.
+*Discussion and citations to Boley et al. 2007, who cover this issue in great detail are now included. In fact,
+it is not clear what the correct physical choice is for molecular hydrogen when using a fixed \gamma.
 
@@
 
-*What is written in the manuscript is correct.
-The key word is "effective." A short cooling time for a truly stiff cp/cv=gamma makes the effective gamma nonstiff.
-The classic example is an isothermal gas.
-We avoid possible confusion on this issue by not addressing this additional complexity.
+*What is written in the manuscript is correct. The key word is "effective." A short cooling time for a truly
+ stiff cp/cv=gamma makes the effective gamma nonstiff.  The classic example is an isothermal gas.  We avoid
+ possible confusion on this issue by not addressing this additional complexity.
 
@@
 
-pg 21: The sentence starting "For example, the Gammie 2001 simulations..." is confusingly written.
-Also, I disagree with the description of the Meru and Bate results adding "weight to the claims by e.g.
-Boss..".
-It seems clear that unless dissipation is dramatically non-local, there must be a relation between the rate of dissipation and cooling.
-I don't think Boss is claiming that this doesn't exist, just that cooling isn't a problem.
+pg 21: The sentence starting "For example, the Gammie 2001 simulations..." is confusingly written. Also, I
+disagree with the description of the Meru and Bate results adding "weight to the claims by e.g. Boss..". It
+seems clear that unless dissipation is dramatically non-local, there must be a relation between the rate of
+dissipation and cooling. I don't think Boss is claiming that this doesn't exist, just that cooling isn't a
+problem.
 
@@
 
-General comment: the numbers on the axes are somewhat small on all of the figures. Depending on how they are eventually sized in publication, I can imagine some of them being hard to read.
+General comment: the numbers on the axes are somewhat small on all of the figures. Depending on how they are
+eventually sized in publication, I can imagine some of them being hard to read.
 

### 2011-10-26 Response_to_Referee1.txt
Worked on referee response.
	modified:   Response_to_Referee1.txt

@@
 directly measure the power in a given eignmode. In addtion, each particular m value makes a relatively minor
-contribution for large m (i.e. m > 20). The measurement of A_m also includes no phase information, so A_m
-is *not* necessarily a good indicator of coherent density structures or organized velocity correlations.
+contribution for large m (i.e. m > 20). The measurement of A_m also includes no phase information, so A_m is
+not necessarily a good indicator of coherent density structures or organized velocity correlations.
 
@@
 
+* The difficulty arises from the fact that the Reynolds stress is typically defined as R_ij = rho*u'_i*u'_j
+  where u'_i and u'_j are the fluctuations in the velocity field. The fluctuating pieces of the velocity field
+  are defined as u'_i = u_i - u<avg>_i, but in a gravitationally unstable disk with extended non-axisymmetric
+  spiral structures, the proper definition of u<avg> (particularly in the phi direction) is unclear. We measured the
+  Reynolds stresses using several different definitions for u<avg> and obtained results that varied
+  dramatically (by an order of magnitude in some cases). The cases we tried were, using an azimuthal average
+  for u<avg>_phi and 0 for u<avg>_r, using an average +/- 2 cells in the direction of interest, using an
+  equiangular average in the direction of interest (i.e. the number of cells used in the average for the phi
+  direction increased with resolution), using a +/- 2 cell patch around the cell, and using an equiangular
+  patch. As stated earlier, results for each of these cases varied substantially. It is also not clear what is
+  the proper scale to average over, as one would like to measure the Reynolds stress caused by turbulent
+  motion, not the ordered motion of spiral arms. SAY SOMETHING HERE ABOUT THE MOST LIKELY CASE BEING A MERE FRACTION OF THE GRAV TORQUE
+
 In addition to measuring the Reynolds stresses directly, the authors could also look at the mass accretion
@@
 contribution from Reynolds stresses.
+
+* We were able to make this measurement and the mass accretion shows a similar trend to the gravitational
+  torque. In addition we calculated the predicted mass accretion based on the measured torques (see Boley et
+  al. 2006 equation 22). These are roughly in line with the measured mass accretion rates. (See attached
+  figures.)
 
@@
 time. Of course, if \alpha is more and more dominated by Reynolds terms, than the heating rate and effective
-alpha might not decline. This is a very important distinction.  The comment in the summary that Q is
-increasing with increasing resolution suggests that there is more heating, and thus more dissipation.
+alpha might not decline. This is a very important distinction. The comment in the summary that Q is increasing
+with increasing resolution suggests that there is more heating, and thus more dissipation.
 
@@
 
-*Spelling fixed, and Pringle cited
+* Spelling fixed, and Pringle cited
 

### 2011-11-03 Response_to_Referee1.txt
Finished response text.
	modified:   Response_to_Referee1.txt
Added in figures and added average mdots to response.
	modified:   Response_to_Referee1.txt
	new file:   response_to_referee1_fig1a.eps
	new file:   response_to_referee1_fig1b.eps
	new file:   response_to_referee1_fig2.eps

@@
   the proper scale to average over, as one would like to measure the Reynolds stress caused by turbulent
-  motion, not the ordered motion of spiral arms. SAY SOMETHING HERE ABOUT THE MOST LIKELY CASE BEING A MERE FRACTION OF THE GRAV TORQUE
+  motion, not the ordered motion of spiral arms. In fact, when the Reynolds stress is measured by using a +/-
+  2 cell patch in both directions, the stresses measured are a few percent of the gravitational torques. (See
+  attached figures.) response_to_referee1_fig1a.eps shows the measured torques from the Reynolds stress from
+  10 to 50 AU averaged from 12 to 18 ORP for  LMAX = 64(black), 128(blue), 256(green), and 512(red), while
+  response_to_referee1_fig1b.eps shows the measured gravitational torques averaged over the same time period.
 
@@
   torque. In addition we calculated the predicted mass accretion based on the measured torques (see Boley et
-  al. 2006 equation 22). These are roughly in line with the measured mass accretion rates. (See attached
-  figures.)
+  al. 2006 equation 22). These are roughly in line with the measured mass accretion rates. Overall the average
+  from 12 to 18 ORP of the absolute value of the measured rates follows the same trend as the torques. For
+  LMAX = 64 the average of the absolute value of the rate was 8.9 x 10^-7 m_sun/yr, for LMAX = 128 the average
+  of the absolute value of the rate was 3.7 x 10^-7 m_sun/yr, for LMAX = 256 the average of the absolute value
+  of the rate was  2.5 x 10^-7 m_sun/yr, and for LMAX = 512 the average of the absolute value of the rate was
+  2.4 x 10^-7 m_sun/yr. (See attached figure.) In response_to_referee1_fig2.eps the solid lines show the
+  predicted mdots (x10^-6 m_sun/yr) vs. radius (AU) from the measured gravitational torques, while the dashed
+  lines show the actual measured mass accretion rates from 12 to 18 ORP. The color scale is the same as in fig1.
 

### 2011-11-07 Paper_1.tex
Added Mdots to the table and some descriptive text.
	modified:   Paper_1.tex

@@
 During the initial axisymmetric phase, 
-which lasts several ORPs, the disk cools and contracts.
+which lasts several ORP, the disk cools and contracts.
 The contraction is small radially but dramatic in the vertical direction.
-Around 3 to 4 ORPs, the instabilities begin to grow to nonlinear amplitudes, and the disk
+Around 3 to 4 ORP, the instabilities begin to grow to nonlinear amplitudes, and the disk
 then undergoes a strong burst of GIs that predominantly manifest themselves in a few discrete global spiral modes.
-The disk expands violently during the burst phase, producing a significant rearrangement of the disk's mass distribution on a timescale of a few ORPs.
-This is then followed by a period of several ORPs where heating temporarily washes out some of the non-axisymmetry in the disk.
+The disk expands violently during the burst phase, producing a significant rearrangement of the disk's mass distribution on a timescale of a few ORP.
+This is then followed by a period of several ORP where heating temporarily washes out some of the non-axisymmetry in the disk.
 Finally, following this transition phase, the disk settles into a quasi-steady, long-lived asymptotic
@@
 Because of the computational overhead associated with running each simulation starting from $t = 0$, the $l_{max} =$ 128 simulation was run through the axisymmetric, burst and transition phases.
-Simulations using $l_{max} =$ 64, 256, and 512 all begin by a linear interpolation of the $l_{max} =$ 128 simulation to higher or lower azimuthal resolution at 9.6 ORPs, near the end of the transition
+Simulations using $l_{max} =$ 64, 256, and 512 all begin by a linear interpolation of the $l_{max} =$ 128 simulation to higher or lower azimuthal resolution at 9.6 ORP, near the end of the transition
 phase.
 Although this strategy saves a large amount of computational time, it limits the scope of the analysis to the asymptotic phase.
-By 12 ORPs, all four disks appear to have transitioned to the asymptotic state,
-e.g., they all display roughly steady $Q(r)$ profiles starting at $\sim 12$ ORPs.  All simulations were 
-run through $\sim 18$ ORPs.
+By 12 ORP, all four disks appear to have transitioned to the asymptotic state,
+e.g., they all display roughly steady $Q(r)$ profiles starting at $\sim 12$ ORP.  All simulations were 
+run through $\sim 18$ ORP.
 
@@
 
-Figure \ref{fig:DensityPlots}, shows mid-plane and radial densities  at $t \sim 18$ ORPs for the $l_{max} =$ 64, 128, 256, and 512 simulations.
+Figure \ref{fig:DensityPlots}, shows mid-plane and radial densities  at $t \sim 18$ ORP for the $l_{max} =$ 64, 128, 256, and 512 simulations.
 It is readily apparent that the character of the non-axisymmetric density structures differ with
@@
 In what follows, $A_m$ is averaged over time to suppress fluctuations on the dynamic timescale due to the chaotic nature of GIs in the asymptotic  phase.
-In particular, $\langle A_{tot} \rangle$ will represent the power over the radial range 10 -- 40 AU, summed from $ m = 2$ to $l_{max}/2$ and averaged over the time range 12 -- 18 ORPs.
+In particular, $\langle A_{tot} \rangle$ will represent the power over the radial range 10 -- 40 AU, summed from $ m = 2$ to $l_{max}/2$ and averaged over the time range 12 -- 18 ORP.
 The Fourier amplitudes cannot be measured at resolutions smaller than $l_{max}/2$, therefore $A_m$ is limited to modes $m \le l_{max}/2$.
@@
 \end{equation}
-Figure \ref{fig:alpha_v_radius} shows $\alpha_{\rm{eff}}$, time-averaged from 12 to 18 ORPs, calculated using equations \eqref{alpha1} and \eqref{newtonstress},
+Figure \ref{fig:alpha_v_radius} shows $\alpha_{\rm{eff}}$, time-averaged from 12 to 18 ORP, calculated using equations \eqref{alpha1} and \eqref{newtonstress},
 for the four grid resolutions.  For comparison, 
@@
 %itself is a dynamically varying quantity, see figure \ref{fig:alphavar}. 
+In addition to the convergence of $\alpha_{avg}$, table \ref{tbl:ams} shows the convergence of the $\dot{M}$
+from 10 -- 40 AU averaged from 12 to 18 ORP. The $\dot{M}_{avg}$ measurement shows convergence similar to that
+of the $\alpha_{avg}$ measurement.
 
@@
 \begin{center}
-\begin{tabular}{ccccc}
+\begin{tabular}{cccccc}
 \multicolumn{5}{c}{$Q$, Fourier Amplitudes, and $\alpha$}\\ \hline\hline
-$l_{max}$ & $Q_{avg}$   & $\langle A_{tot} \rangle$   & $\langle A_{2-7} \rangle / \langle A_{tot} \rangle$ & $\alpha_{avg}$ \\\hline
-64   & 1.26 & 1.96 &  0.76 & 0.069 \\
-128 & 1.24 & 1.91 &  0.62 & 0.044 \\
-256 & 1.39 & 1.83 &  0.54 & 0.026 \\
-512 & 1.48 & 1.95 &  0.49 & 0.020 \\ 
-$\infty$ & & &0.43 & 0.017 \\ \hline
+$l_{max}$ & $Q_{avg}$   & $\langle A_{tot} \rangle$   & $\langle A_{2-7} \rangle / \langle A_{tot} \rangle$ &
+$\alpha_{avg}$ & $\dot{M}_{avg} ~\mathrm{M_\odot/yr}$\\\hline
+64   & 1.26 & 1.96 &  0.76 & 0.069 & 8.9$ \times 10^{-7}$ \\
+128 & 1.24 & 1.91 &  0.62 & 0.044 & 3.7$ \times 10^{-7}$ \\
+256 & 1.39 & 1.83 &  0.54 & 0.026 & 2.5$ \times 10^{-7}$ \\
+512 & 1.48 & 1.95 &  0.49 & 0.020 & 2.4$ \times 10^{-7}$ \\ 
+$\infty$ & & &0.43 & 0.017 & \\ \hline
 \label{tbl:ams}
@@
 \caption{ All quantities are time-averaged over 12 -- 18 ORP.
-$Q_{avg}$ and $\alpha_{avg}$ are also spatially
+$Q_{avg}$, $\alpha_{avg}$, and $\dot{M}_{avg}$  are also spatially
 averaged over 10 -- 40 AU.  The entries for ``$\infty$'' are based on Richardson extrapolation,
@@
 \plotone{Figure1.eps}
-\caption{Azimuthally averaged mid-plane Toomre Q distribution versus radius for each of the four resolutions at t $\sim$ 18 ORPs}
+\caption{Azimuthally averaged mid-plane Toomre Q distribution versus radius for each of the four resolutions at t $\sim$ 18 ORP}
 \label{fig:Final_Q}
@@
 {
-Mid-plane disk densities at $t \sim 18$ ORPs for simulations with 64, 128, 256 and 512 azimuthal grid elements.
+Mid-plane disk densities at $t \sim 18$ ORP for simulations with 64, 128, 256 and 512 azimuthal grid elements.
 The simulations differ only in the number of  azimuthal zones used in the calculations.
@@
 {Effective $\alpha$'s for the $t_{cool} =$ 2 ORP, $\gamma = 5/3$, $\Sigma \sim r^{ -1}$ disk, averaged between 12 and 18
-  ORPs, for four different azimuthal grid resolutions, given by the number of azimuthal grid-points $l_{max}$. The
+  ORP, for four different azimuthal grid resolutions, given by the number of azimuthal grid-points $l_{max}$. The
   dashed lines show the $\alpha$ from equation (\ref{eq:gammie}) (upper: strongly self-gravitating limit, lower:

### 2011-11-09 Response_to_Referee1.txt
Dick's changes to the referees report and paper, minor tweak to Figure 5.
	modified:   Paper_1.tex
	modified:   Response_to_Referee1.txt
	modified:   figures/Figure5.eps

--- 2011-10-26 Response_to_Referee1.txt
@@
-This is the response to the first referee report for ApJ submission MS#ApJ83382
+This is the response to the first referee report for ApJ submission MS#ApJ83382.
 
@@
 Fourier amplitudes, as defined in eqns 2 and 3. As can be seen by their definition, these values do not
-directly measure the power in a given eignmode. In addtion, each particular m value makes a relatively minor
+directly measure the power in a given eignmode. In addition, each particular m value makes a relatively minor
 contribution for large m (i.e. m > 20). The measurement of A_m also includes no phase information, so A_m is
@@
   are defined as u'_i = u_i - u<avg>_i, but in a gravitationally unstable disk with extended non-axisymmetric
-  spiral structures, the proper definition of u<avg> (particularly in the phi direction) is unclear. We measured the
-  Reynolds stresses using several different definitions for u<avg> and obtained results that varied
-  dramatically (by an order of magnitude in some cases). The cases we tried were, using an azimuthal average
-  for u<avg>_phi and 0 for u<avg>_r, using an average +/- 2 cells in the direction of interest, using an
-  equiangular average in the direction of interest (i.e. the number of cells used in the average for the phi
-  direction increased with resolution), using a +/- 2 cell patch around the cell, and using an equiangular
-  patch. As stated earlier, results for each of these cases varied substantially. It is also not clear what is
-  the proper scale to average over, as one would like to measure the Reynolds stress caused by turbulent
-  motion, not the ordered motion of spiral arms. SAY SOMETHING HERE ABOUT THE MOST LIKELY CASE BEING A MERE FRACTION OF THE GRAV TORQUE
+  spiral structures, the proper definition of u<avg> (particularly in the phi direction) is unclear. We
+  measured the Reynolds stresses using several different definitions for u<avg> and obtained results that
+  varied dramatically (by an order of magnitude in some cases). The cases we tried were: using an azimuthal
+  average for u<avg>_phi and 0 for u<avg>_r, using an average +/- 2 cells in the direction of interest, using
+  an equiangular average in the direction of interest (i.e. the number of cells used in the average for the
+  phi direction increased with resolution), using a +/- 2 cell patch around the cell, and using an equiangular
+  r,phi patch. As stated earlier, results for each of these cases varied substantially. It is also not clear
+  what is the proper scale to average over, as one would like to measure the Reynolds stress caused by
+  turbulent motion, not the ordered motion of spiral arms. In fact, when the Reynolds stress is measured by
+  using a +/- 2 cell patch in both directions, the stresses measured are only a few percent of the
+  gravitational torques.  (See attached figures.) response_to_referee1_fig1a.eps shows the measured torques
+  from the Reynolds stress from 10 to 50 AU averaged from 12 to 18 ORP for LMAX = 64(black), 128(blue),
+  256(green), and 512(red), while response_to_referee1_fig1b.eps shows the measured gravitational torques
+  averaged over the same time period.
 
@@
 
-* We were able to make this measurement and the mass accretion shows a similar trend to the gravitational
-  torque. In addition we calculated the predicted mass accretion based on the measured torques (see Boley et
-  al. 2006 equation 22). These are roughly in line with the measured mass accretion rates. (See attached
-  figures.)
+* We were able to make this measurement and the mass accretion shows convergence similar to the gravitational
+  torques. In addition we calculated an expected mass accretion based on the measured torques (see Boley et
+  al. 2006 equation 22), and the gravitational torques alone are sufficient to account for the measured
+  accretion rates. For LMAX = 64 the average over 10 to 40 AU of the absolute value of the measured radial
+  mass flow rate was 8.9 x 10^-7 m_sun/yr; for LMAX = 128 it was 3.7 x 10^-7 m_sun/yr, for LMAX = 256 2.5 x
+  10^-7 m_sun/yr, and for LMAX = 512 2.4 x 10^-7 m_sun/yr. The mass inflow rates are thus converging in a
+  manner similar to the alpha measured from the gravitational torques alone. There is no indication of missing
+  contributions to alpha from Reynolds stresses. This information is now included in Table 1 and in the text
+  in Section 3.4.
 
@@
 with increasing resolution suggests that there is more heating, and thus more dissipation.
+
+* Our code conserves angular momentum with more precision than it conserves energy, and so we have more
+  confidence about changes in the radial position of mass elements due to changes in their specific angular
+  momentum than in measurements of energy dissipation rates. The similar convergence of the measured 
+  mass accretion rates and the gravitational alpha give us confidence that gravitational torques 
+  dominate the mass transport.
 
@@
 
-* Spelling fixed, and Pringle cited
+* Spelling fixed, and Pringle cited.
 
@@
 
-*Citation included
+* Citation included.
 
@@
 
-*Kratter citation included, however we were unable to locate the Walsh paper referenced here.
+* Kratter citation included, however we were unable to locate the Walsh paper referenced here.
 
@@
 
-*First paragraph on page 7 rewritten to attempt to elucidate this issue.
+* First paragraph on page 7 is rewritten to elucidate this issue.
 
@@
 
-*Changed
+* Changed.
 
@@
 
-*Citations included which explain this in detail.
+* Citations are now included where this is explained in detail.
 
@@
 
-*Wording changed to be more clear.
+* Wording changed to be more clear.
 
@@
 
-*Discussion and citations to Boley et al. 2007, who cover this issue in great detail are now included. In fact,
+* Discussion and citations to Boley et al. 2007, who cover this issue in great detail are now included. In fact,
   it is not clear what the correct physical choice is for molecular hydrogen when using a fixed \gamma.
@@
 
-*Citation to Numerical Recipes included.
+* Citation to Numerical Recipes included.
 
@@
 
-*What is written in the manuscript is correct. The key word is "effective." A short cooling time for a truly
+* What is written in the manuscript is correct. The key word is "effective." A short cooling time for a truly
   stiff cp/cv=gamma makes the effective gamma nonstiff.  The classic example is an isothermal gas.  We avoid
@@
 
-*No
+* No, "militate" is the intended meaning.
 
@@
 
-*Text modified as follows: "half the local orbital period (tcool~3/Omega)."
+* Text modified as follows: "half the local orbital period (tcool~3/Omega)."
 
@@
 
-*Modified as follows: "For example, the Gammie (2001) simulations have revealed the most stringent cooling
+* Modified as follows: "For example, the Gammie (2001) simulations have revealed the most stringent cooling
   criterion of all the tcool-fragmentation studies, and due to the local approximation Gammie used, the
@@
 
-*Rewritten for clarity.
+* Rewritten for clarity.
 
@@
 
-*Both color and B/W versions of the figures have been provided.
+* Both color and B/W versions of the figures have been provided.
 
@@
 
-*This has been corrected
+* This has been corrected.
 
@@
 
-*This has been corrected
+* This has been corrected.

### 2011-11-09 Paper_1.tex
Dick's changes to the referees report and paper, minor tweak to Figure 5.
	modified:   Paper_1.tex
	modified:   Response_to_Referee1.txt
	modified:   figures/Figure5.eps

@@
 
-The results above permit an assessment of appropriate azimuthal grids for the
-problem under consideration and similar problems. The $l_{max} = 512$ simulation yields a value 
-for $\alpha_{avg}$ of 0.020. This overestimates the extrapolated converged value of $\alpha_{avg}
-= 0.017$ by $\sim 18$\%. The $l_{max} = 512$ simulation also overestimates the fractional
-power in the high-order ($m = 2$ -- 7) modes by essentially the same percentage.
-However, it should be noted that the fractional difference between successive doubling of the resolution has decreased for $l_{max} = 512$.
-From $l_{max} =64$ to 128 the fractional difference in $\alpha_{avg}$ is 36\%.
-This decreases to 26\% for the jump from $l_{max} = 256$ to 512.
-Furthermore, the converged value represents infinite azimuthal resolution, which is, of course, impossible to achieve in a fixed grid scheme.
+The results above permit an assessment of appropriate azimuthal grids for the problem under consideration and
+similar problems. The $l_{max} = 512$ simulation yields a value for $\alpha_{avg}$ of 0.020. This
+overestimates the extrapolated converged value of $\alpha_{avg} = 0.017$ by $\sim 18$\%. The $l_{max} = 512$
+simulation also overestimates the fractional power in the high-order ($m = 2$ -- 7) modes by essentially the
+same percentage. However, it should be noted that the fractional difference between successive doubling of the
+resolution has decreased for $l_{max} = 512$. From $l_{max} =64$ to 128 the fractional difference in
+$\alpha_{avg}$ is 36\%. This decreases to 26\% for the jump from $l_{max} = 256$ to 512. Furthermore, the
+converged value represents infinite azimuthal resolution, which is, of course, impossible to achieve in a
+fixed grid scheme.
 %It is unclear as to what variation of $\alpha$ between two values of $l_{max}$ would represent convergence as $\alpha$ 
 %itself is a dynamically varying quantity, see figure \ref{fig:alphavar}.
-In addition to the convergence of $\alpha_{avg}$, table \ref{tbl:ams} shows the convergence of the $\dot{M}$
-from 10 -- 40 AU averaged from 12 to 18 ORP. The $\dot{M}_{avg}$ measurement shows convergence similar to that
-of the $\alpha_{avg}$ measurement.
+
+%DICK: I made changes here.
+In addition to the convergence of $\alpha_{avg}$, table 1 illustrates the convergence of the measured net
+inward mass transport rates $\dot{M}$ over the time interval 12 -- 18 ORP. We tabulate the value of
+$|\dot{M}|$ spatially averaged over 10 -- 40 AU. The absolute value of the net $\dot{M}$ at each $r$ is taken
+prior to the spatial average because $\dot{M}$ varies in sign with $r$. It is the magnitude of the typical net
+mass transport rates that we wish to compare. The $|\dot{M}|_{avg}$ measurements show convergence similar to
+that of the $\alpha_{avg}$ measurement based on the gravitational torques. This gives us confidence that no
+important contributions from Reynolds stresses are being overlooked. Moreover, we have used eq. (22) of
+\citet{boley2006} to estimate mass inflow rates based on the measured gravitational torques and find that
+these torques are indeed sufficient to account for the measured $\dot{M}$.
 
@@
 
+%DICK: I made changes here.
 \begin{table}
@@
 \begin{tabular}{cccccc}
-\multicolumn{5}{c}{$Q$, Fourier Amplitudes, and $\alpha$}\\ \hline\hline
+\multicolumn{5}{c}{$Q$, Fourier Amplitudes, $\alpha$, and $\dot{M}$}\\ \hline\hline
 $l_{max}$ & $Q_{avg}$   & $\langle A_{tot} \rangle$   & $\langle A_{2-7} \rangle / \langle A_{tot} \rangle$ &
-$\alpha_{avg}$ & $\dot{M}_{avg} ~\mathrm{M_\odot/yr}$\\\hline
+$\alpha_{avg}$ & $|\dot{M}|_{avg} ~\mathrm{M_\odot/yr}$\\\hline
 64   & 1.26 & 1.96 &  0.76 & 0.069 & 8.9$ \times 10^{-7}$ \\
@@
 \caption{ All quantities are time-averaged over 12 -- 18 ORP.
-$Q_{avg}$, $\alpha_{avg}$, and $\dot{M}_{avg}$  are also spatially
+$Q_{avg}$, $\alpha_{avg}$, and $|\dot{M}|_{avg}$  are also spatially
 averaged over 10 -- 40 AU.  The entries for ``$\infty$'' are based on Richardson extrapolation,

### 2011-11-30 Paper_1_submisson2.tex
Paper more or less accepted, first cut at final changes offered by referee.
	modified:   Paper_1.tex
	new file:   Paper_1_submisson2.tex
	modified:   Response_to_Referee1.txt
	modified:   general.bib

--- 2011-11-09 Paper_1.tex
@@
 
-\bibliographystyle{apj}
-\bibliography{general}
+\begin{thebibliography}{66}
+\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
+
+\bibitem[{{Balbus} \& {Papaloizou}(1999)}]{balbus1999}
+{Balbus}, S.~A. \& {Papaloizou}, J.~C.~B. 1999, ApJ, 521, 650
+
+\bibitem[{{Boley}(2009)}]{boley2009}
+{Boley}, A.~C. 2009, ApJL, 695, L53
+
+\bibitem[{{Boley} \& {Durisen}(2008)}]{boley2008}
+{Boley}, A.~C. \& {Durisen}, R.~H. 2008, ApJ, 685, 1193
+
+\bibitem[{{Boley} {et~al.}(2007{\natexlab{a}}){Boley}, {Durisen}, {Nordlund},
+  \& {Lord}}]{boley2007b}
+{Boley}, A.~C., {Durisen}, R.~H., {Nordlund}, {\AA}., \& {Lord}, J.
+  2007{\natexlab{a}}, ApJ, 665, 1254
+
+\bibitem[{{Boley} {et~al.}(2007{\natexlab{b}}){Boley}, {Hartquist}, {Durisen},
+  \& {Michael}}]{boley2007a}
+{Boley}, A.~C., {Hartquist}, T.~W., {Durisen}, R.~H., \& {Michael}, S.
+  2007{\natexlab{b}}, ApJL, 656, L89
+
+\bibitem[{{Boley} {et~al.}(2006){Boley}, {Mej{\'{\i}}a}, {Durisen}, {Cai},
+  {Pickett}, \& {D'Alessio}}]{boley2006}
+{Boley}, A.~C., {Mej{\'{\i}}a}, A.~C., {Durisen}, R.~H., {Cai}, K., {Pickett},
+  M.~K., \& {D'Alessio}, P. 2006, ApJ, 651, 517
+
+\bibitem[{{Boss}(1984)}]{boss1984}
+{Boss}, A.~P. 1984, MNRAS, 209, 543
+
+\bibitem[{{Boss}(2002)}]{boss2002}
+---. 2002, ApJ, 576, 462
+
+\bibitem[{{Boss}(2009)}]{boss2009}
+---. 2009, ApJ, 694, 107
+
+\bibitem[{{Cai} {et~al.}(2008){Cai}, {Durisen}, {Boley}, {Pickett}, \&
+  {Mej{\'{\i}}a}}]{cai2008}
+{Cai}, K., {Durisen}, R.~H., {Boley}, A.~C., {Pickett}, M.~K., \&
+  {Mej{\'{\i}}a}, A.~C. 2008, ApJ, 673, 1138
+
+\bibitem[{{Cai} {et~al.}(2006){Cai}, {Durisen}, {Michael}, {Boley},
+  {Mej{\'{\i}}a}, {Pickett}, \& {D'Alessio}}]{cai2006}
+{Cai}, K., {Durisen}, R.~H., {Michael}, S., {Boley}, A.~C., {Mej{\'{\i}}a},
+  A.~C., {Pickett}, M.~K., \& {D'Alessio}, P. 2006, ApJL, 636, L149
+
+\bibitem[{{Christodoulou}(1991)}]{christodoulou1991}
+{Christodoulou}, D.~M. 1991, \apj, 372, 471
+
+\bibitem[{{Clarke} {et~al.}(2007){Clarke}, {Harper-Clark}, \&
+  {Lodato}}]{clarke2007}
+{Clarke}, C.~J., {Harper-Clark}, E., \& {Lodato}, G. 2007, MNRAS, 381, 1543
+
+\bibitem[{{Cossins} {et~al.}(2009){Cossins}, {Lodato}, \&
+  {Clarke}}]{cossins2009}
+{Cossins}, P., {Lodato}, G., \& {Clarke}, C.~J. 2009, \mnras, 393, 1157
+
+\bibitem[{{Cossins} {et~al.}(2010){Cossins}, {Lodato}, \&
+  {Clarke}}]{cossins2010}
+---. 2010, \mnras, 401, 2587
+
+\bibitem[{{Durisen} {et~al.}(2007){Durisen}, {Boss}, {Mayer}, {Nelson},
+  {Quinn}, \& {Rice}}]{durisen2007}
+{Durisen}, R.~H., {Boss}, A.~P., {Mayer}, L., {Nelson}, A.~F., {Quinn}, T., \&
+  {Rice}, W.~K.~M. 2007, in Protostars and Planets V, B. Reipurth, D. Jewitt,
+  and K. Keil (eds.), University of Arizona Press, Tucson, 951 pp., 2007.,
+  p.607-622, ed. B.~{Reipurth}, D.~{Jewitt}, \& K.~{Keil}, 607--622
+
+\bibitem[{{Durisen} {et~al.}(2005){Durisen}, {Cai}, {Mej{\'{\i}}a}, \&
+  {Pickett}}]{durisen2005}
+{Durisen}, R.~H., {Cai}, K., {Mej{\'{\i}}a}, A.~C., \& {Pickett}, M.~K. 2005,
+  Icarus, 173, 417
+
+\bibitem[{{Durisen} {et~al.}(1986){Durisen}, {Gingold}, {Tohline}, \&
+  {Boss}}]{durisen1986}
+{Durisen}, R.~H., {Gingold}, R.~A., {Tohline}, J.~E., \& {Boss}, A.~P. 1986,
+  ApJ, 305, 281
+
+\bibitem[{{Gammie}(2001)}]{gammie2001}
+{Gammie}, C.~F. 2001, ApJ, 553, 174
+
+\bibitem[{{Goldreich} \& {Lynden-Bell}(1965)}]{goldreich1965}
+{Goldreich}, P. \& {Lynden-Bell}, D. 1965, MNRAS, 130, 125
+
+\bibitem[{{Hachisu}(1986)}]{hachisu1986}
+{Hachisu}, I. 1986, ApJS, 61, 479
+
+\bibitem[{{Kratter} {et~al.}(2010){Kratter}, {Matzner}, {Krumholz}, \&
+  {Klein}}]{kratter2010}
+{Kratter}, K.~M., {Matzner}, C.~D., {Krumholz}, M.~R., \& {Klein}, R.~I. 2010,
+  ApJ, 708, 1585
+
+\bibitem[{{Larson}(1984)}]{larson1984}
+{Larson}, R.~B. 1984, MNRAS, 206, 197
+
+\bibitem[{{Laughlin} \& {Bodenheimer}(1994)}]{laughlin1994}
+{Laughlin}, G. \& {Bodenheimer}, P. 1994, ApJ, 436, 335
+
+\bibitem[{{Laughlin} {et~al.}(1997){Laughlin}, {Korchagin}, \&
+  {Adams}}]{laughlin1997}
+{Laughlin}, G., {Korchagin}, V., \& {Adams}, F.~C. 1997, ApJ, 477, 410
+
+\bibitem[{{Laughlin} {et~al.}(1998){Laughlin}, {Korchagin}, \&
+  {Adams}}]{laughlin1998}
+---. 1998, ApJ, 504, 945
+
+\bibitem[{{Laughlin} \& {Rozyczka}(1996)}]{laughlin1996}
+{Laughlin}, G. \& {Rozyczka}, M. 1996, \apj, 456, 279
+
+\bibitem[{{Lin}(1981)}]{lin1981}
+{Lin}, D.~N.~C. 1981, ApJ, 246, 972
+
+\bibitem[{{Lin} \& {Pringle}(1987)}]{lin1987}
+{Lin}, D.~N.~C. \& {Pringle}, J.~E. 1987, \mnras, 225, 607
+
+\bibitem[{{Lin} \& {Pringle}(1990)}]{lin1990}
+---. 1990, ApJ, 358, 515
+
+\bibitem[{{Lodato} \& {Rice}(2004)}]{lodato2004}
+{Lodato}, G. \& {Rice}, W.~K.~M. 2004, MNRAS, 351, 630
+
+\bibitem[{{Lodato} \& {Rice}(2005)}]{lodato2005}
+---. 2005, MNRAS, 358, 1489
+
+\bibitem[{{Lynden-Bell} \& {Kalnajs}(1972)}]{lyndenbell1972}
+{Lynden-Bell}, D. \& {Kalnajs}, A.~J. 1972, MNRAS, 157, 1
+
+\bibitem[{{Mayer} {et~al.}(2002){Mayer}, {Quinn}, {Wadsley}, \&
+  {Stadel}}]{mayer2002}
+{Mayer}, L., {Quinn}, T., {Wadsley}, J., \& {Stadel}, J. 2002, Science, 298,
+  1756
+
+\bibitem[{{Mej{\'{\i}}a}(2004)}]{mejiaphd2004}
+{Mej{\'{\i}}a}, A.~C. 2004, PhD thesis, Indiana University
+
+\bibitem[{{Mej{\'{\i}}a} {et~al.}(2005){Mej{\'{\i}}a}, {Durisen}, {Pickett}, \&
+  {Cai}}]{mejia2005}
+{Mej{\'{\i}}a}, A.~C., {Durisen}, R.~H., {Pickett}, M.~K., \& {Cai}, K. 2005,
+  ApJ, 619, 1098
+
+\bibitem[{{Meru} \& {Bate}(2011{\natexlab{a}})}]{meru2011a}
+{Meru}, F. \& {Bate}, M.~R. 2011{\natexlab{a}}, MNRAS, 411, L1
+
+\bibitem[{{Meru} \& {Bate}(2011{\natexlab{b}})}]{meru2011b}
+---. 2011{\natexlab{b}}, MNRAS, 410, 559
+
+\bibitem[{{Michael} \& {Durisen}(2010)}]{michael2010}
+{Michael}, S. \& {Durisen}, R.~H. 2010, MNRAS, 406, 279
+
+\bibitem[{{Nelson} {et~al.}(1998){Nelson}, {Benz}, {Adams}, \&
+  {Arnett}}]{nelson1998}
+{Nelson}, A.~F., {Benz}, W., {Adams}, F.~C., \& {Arnett}, D. 1998, ApJ, 502,
+  342
+
+\bibitem[{{Nelson} {et~al.}(2000){Nelson}, {Benz}, \&
+  {Ruzmaikina}}]{nelson2000a}
+{Nelson}, A.~F., {Benz}, W., \& {Ruzmaikina}, T.~V. 2000, ApJ, 529, 357
+
+\bibitem[{{Norman} \& {Winkler}(1986)}]{norman1986}
+{Norman}, M.~L. \& {Winkler}, K.-H.~A. 1986, in NATO Advanced Research Workshop
+  on Astrophysical Radiation Hydrodynamics,, ed. K.-H.~A. {Winkler} \& M.~L.
+  {Norman} (D. Reidel Publishing Co., Boston, MA), 187
+
+\bibitem[{{Paczynski}(1978)}]{paczynski1978}
+{Paczynski}, B. 1978, \actaa, 28, 91
+
+\bibitem[{{Papaloizou} \& {Savonije}(1991)}]{papaloizou1991}
+{Papaloizou}, J.~C. \& {Savonije}, G.~J. 1991, MNRAS, 248, 353
+
+\bibitem[{{Pickett} {et~al.}(1998){Pickett}, {Cassen}, {Durisen}, \&
+  {Link}}]{pickett1998}
+{Pickett}, B.~K., {Cassen}, P., {Durisen}, R.~H., \& {Link}, R. 1998, ApJ, 504,
+  468
+
+\bibitem[{{Pickett} {et~al.}(2000){Pickett}, {Cassen}, {Durisen}, \&
+  {Link}}]{pickett2000}
+---. 2000, ApJ, 529, 1034
+
+\bibitem[{{Pickett} {et~al.}(1996){Pickett}, {Durisen}, \&
+  {Davis}}]{pickett1996}
+{Pickett}, B.~K., {Durisen}, R.~H., \& {Davis}, G.~A. 1996, ApJ, 458, 714
+
+\bibitem[{{Pickett} {et~al.}(2003){Pickett}, {Mej{\'{\i}}a}, {Durisen},
+  {Cassen}, {Berry}, \& {Link}}]{pickett2003}
+{Pickett}, B.~K., {Mej{\'{\i}}a}, A.~C., {Durisen}, R.~H., {Cassen}, P.~M.,
+  {Berry}, D.~K., \& {Link}, R.~P. 2003, ApJ, 590, 1060
+
+\bibitem[{Press {et~al.}(1992)Press, Teukolsky, Vetterling, \&
+  Flannery}]{press1992}
+Press, W., Teukolsky, S., Vetterling, W., \& Flannery, B. 1992, {Numerical
+  Recipes in C}, 2nd edn. (Cambridge, UK: Cambridge University Press)
+
+\bibitem[{{Pringle}(1981)}]{pringle1981}
+{Pringle}, J.~E. 1981, \araa, 19, 137
+
+\bibitem[{{Rice} {et~al.}(2003){Rice}, {Armitage}, {Bate}, \&
+  {Bonnell}}]{rice2003b}
+{Rice}, W.~K.~M., {Armitage}, P.~J., {Bate}, M.~R., \& {Bonnell}, I.~A. 2003,
+  MNRAS, 339, 1025
+
+\bibitem[{{Rice} {et~al.}(2005){Rice}, {Lodato}, \& {Armitage}}]{rice2005}
+{Rice}, W.~K.~M., {Lodato}, G., \& {Armitage}, P.~J. 2005, MNRAS, 364, L56
+
+\bibitem[{{Shakura} \& {Sunyaev}(1973)}]{shakura1973}
+{Shakura}, N.~I. \& {Sunyaev}, R.~A. 1973, A\&A, 24, 337
+
+\bibitem[{Shannon(1984)}]{shannon1984}
+Shannon, C. 1984, Proceedings of the IEEE, 72, 1192
+
+\bibitem[{{Shu} {et~al.}(1990){Shu}, {Tremaine}, {Adams}, \& {Ruden}}]{shu1990}
+{Shu}, F.~H., {Tremaine}, S., {Adams}, F.~C., \& {Ruden}, S.~P. 1990, ApJ, 358,
+  495
+
+\bibitem[{{Stamatellos} \& {Whitworth}(2008)}]{stamatellos2008}
+{Stamatellos}, D. \& {Whitworth}, A.~P. 2008, A\&A, 480, 879
+
+\bibitem[{{Tomley} {et~al.}(1991){Tomley}, {Cassen}, \&
+  {Steiman-Cameron}}]{tomley1991}
+{Tomley}, L., {Cassen}, P., \& {Steiman-Cameron}, T. 1991, ApJ, 382, 530
+
+\bibitem[{{Tomley} {et~al.}(1994){Tomley}, {Steiman-Cameron}, \&
+  {Cassen}}]{tomley1994}
+{Tomley}, L., {Steiman-Cameron}, T.~Y., \& {Cassen}, P. 1994, ApJ, 422, 850
+
+\bibitem[{{Toomre}(1981)}]{toomre1981}
+{Toomre}, A. 1981, in Structure and Evolution of Normal Galaxies, ed.
+  {S.~M.~Fall \& D.~Lynden-Bell}, 111--136
+
+\bibitem[{{van Albada} {et~al.}(1982){van Albada}, {van Leer}, \&
+  {Roberts}}]{albada1982}
+{van Albada}, G.~D., {van Leer}, B., \& {Roberts}, Jr., W.~W. 1982, \aap, 108,
+  76
+
+\bibitem[{{Vorobyov}(2010)}]{vorobyov2010}
+{Vorobyov}, E.~I. 2010, New Astronomy, 15, 24
+
+\bibitem[{{Vorobyov} \& {Basu}(2005)}]{vorobyov2005}
+{Vorobyov}, E.~I. \& {Basu}, S. 2005, \mnras, 360, 675
+
+\bibitem[{{Vorobyov} \& {Basu}(2006)}]{vorobyov2006}
+---. 2006, ApJ, 650, 956
+
+\bibitem[{{Yang}(1992)}]{yangphd1992}
+{Yang}, S.~X. 1992, PhD thesis, Indiana University
+
+\bibitem[{{Yorke} \& {Bodenheimer}(1999)}]{yorke1999}
+{Yorke}, H.~W. \& {Bodenheimer}, P. 1999, ApJ, 525, 330
+
+\bibitem[{{Yorke} {et~al.}(1993){Yorke}, {Bodenheimer}, \&
+  {Laughlin}}]{yorke1993}
+{Yorke}, H.~W., {Bodenheimer}, P., \& {Laughlin}, G. 1993, ApJ, 411, 274
+
+\end{thebibliography}
 

### 2011-12-06 Paper_1.tex
Changed figure. Added Dick's edits and a second response to referee.
	modified:   Paper_1.tex
	new file:   Response_to_Referee2.txt
	modified:   figures/Figure5.eps

@@
 
-% \slugcomment{Not to appear in Nonlearned J., 45.}
-
 \shorttitle{Effective Disk Alphas}
@@
 
-%
 \begin{document}
@@
 effective $\alpha$ also depends upon the magnitude of the stresses, thus
-$\alpha_{eff}$ also decreases with increasing resolution. Our converged
-$\alpha_{eff}$ is consistent with predictions from
+$\alpha_{\rm eff}$ also decreases with increasing resolution. Our converged
+$\alpha_{\rm eff}$ is consistent with predictions from
 an analytic local theory for thin disks by Gammie (2001), but only over many
@@
 This almost certainly applies to the embedded phase of both high and low-mass stars when infall from the collapsing protostellar cloud feeds mass into the disk at a rapid
-rate \citep[e.g.,][]{yorke1993, laughlin1994, yorke1999,vorobyov2005,vorobyov2006,kratter2010}.  
+rate \citep[e.g.,][]{yorke1993, laughlin1994, yorke1999,vorobyov2005,vorobyov2006,kratter2010,walch2010}.  
 However, despite the long-distance nature of gravitational torques, theoretical and numerical results suggest that in protostellar disks where heating and cooling are in rough balance, 
@@
 We note, however, that the these studies all assumed
-a {\it local cooling time}, $t_{cool}\Omega =$ constant, 
-where $t_{cool}$ is defined by the rate at which cooling decreases the disk's internal energy density
-$\epsilon$ by $d\epsilon/dt = \epsilon / t_{cool}$.  In contrast, we and our collaborators use either a {\it global cooling time},
-$t_{cool} = $ constant \citep[e.g.,][this paper]{pickett2003, mejia2005}, or use realistic opacities \citep[e.g.,][]{mejiaphd2004, cai2006, cai2008, boley2006, boley2007b,  boley2009} which lead to something in between $t_{cool} = $ constant  and
-$t_{cool}\Omega =$ constant. The review by \citet{durisen2007} discusses some consequences of these various
-choices. There is a tendency for $t_{cool}\Omega =$ constant to bias a disk toward behavior like that of a
+a {\it local cooling time}, $t_{\rm cool}\Omega =$ constant, 
+where $t_{\rm cool}$ is defined by the rate at which cooling decreases the disk's internal energy density
+$\epsilon$ by $d\epsilon/dt = \epsilon / t_{\rm cool}$.  In contrast, we and our collaborators use either a {\it global cooling time},
+$t_{\rm cool} = $ constant \citep[e.g.,][this paper]{pickett2003, mejia2005}, or use realistic opacities \citep[e.g.,][]{mejiaphd2004, cai2006, cai2008, boley2006, boley2007b,  boley2009} which lead to something in between $t_{\rm cool} = $ constant  and
+$t_{\rm cool}\Omega =$ constant. The review by \citet{durisen2007} discusses some consequences of these various
+choices. There is a tendency for $t_{\rm cool}\Omega =$ constant to bias a disk toward behavior like that of a
 steady-state $\alpha$-disk.
@@
 caused by GIs was balanced by the local cooling rate \citep[see also][]{pringle1981}.  
-Assuming a local cooling time of $t_{cool}\Omega =$ constant,  Gammie found
+Assuming a local cooling time of $t_{\rm cool}\Omega =$ constant,  Gammie found
 \begin{equation}
-\alpha_{\rm{eff}} = [\gamma(\gamma - 1) \slantfrac{9}{4}\Omega t_{cool}]^{-1},
+\alpha_{\rm{eff}} = [\gamma(\gamma - 1) \slantfrac{9}{4}\Omega t_{\rm cool}]^{-1},
 \label{eq:gammie}
@@
 Fully 3D simulations \citep[e.g.,][]{lodato2004, cossins2009} suggest that, at least in the case of imposed 
-$t_{cool}\Omega = $ constant cooling, $\alpha_{\rm{\rm{eff}}}$ roughly converges to the value given by equation (\ref{eq:gammie}).
+$t_{\rm cool}\Omega = $ constant cooling, $\alpha_{\rm{eff}}$ roughly converges to the value given by equation (\ref{eq:gammie}).
 
@@
  5/3$ is an appropriate choice.
-The model disks are cooled by decreasing their internal energy according to the prescription $d\epsilon/dt = \epsilon / t_{cool}$, where $t_{cool}$ is the global cooling time, which is set to 2 ORP everywhere.
+The model disks are cooled by decreasing their internal energy according to the prescription $d\epsilon/dt = \epsilon / t_{\rm cool}$, where $t_{\rm cool}$ is the global cooling time, which is set to 2 ORP everywhere.
 
@@
 %Table \ref{tbl:ams} 
-Table 1 lists $Q_{avg}$, the arithmetic mean of $Q$ for each of the asymptotic disks, time-averaged from 12 -- 18 ORP and spatially averaged over 10 -- 40 AU.
+Table 1 lists $Q_{\rm avg}$, the arithmetic mean of $Q$ for each of the asymptotic disks, time-averaged from 12 -- 18 ORP and spatially averaged over 10 -- 40 AU.
 From $\sim$ 12 -- 50 AU, $Q \sim 1.0$ -- 1.4 and thus the disk is subject to GIs over this full range.
@@
 %\begin{equation}
-%\alpha_{\rm{eff}} = [\gamma(\gamma - 1) \slantfrac{9}{4}\Omega t_{cool}]^{-1}
+%\alpha_{\rm{eff}} = [\gamma(\gamma - 1) \slantfrac{9}{4}\Omega t_{\rm cool}]^{-1}
 %\end{equation}
-%where $\gamma$ is the two dimensional adiabatic index and $t_{cool}$ is a cooling parameter. 
+%where $\gamma$ is the two dimensional adiabatic index and $t_{\rm cool}$ is a cooling parameter. 
 %The two dimensional index relates
@@
 
-Numerical uncertainties render Reynolds stresses difficult to determine accurately. 
-This difficulty is mitigated by the fact that 
-previous studies have found Reynolds stresses to be small relative to gravitational stresses (see, 
-for example, figure 5 of Lodato \& Rice 2004 and figure 10 of Boley et al. 2006).  This  suggests that the Reynolds stress
-can be neglected in the evaluation of $\alpha_{\rm{eff}}$.   For this reason, 
-here we omit
-$T_{\varpi\phi}^{Reyn}$ from the calculation of $\alpha_{\rm{eff}}$. 
-We note that \cite{gammie2001}  found
-Reynolds and gravitational stresses contributed equally to $\alpha_{\rm{eff}}$ in his study.  However,
-his was a local, shearing-sheet, 2D study.  As such, it was unable to capture global effects, such
-as those seen in the cited 3D studies.
+As discussed in section \ref{sec:converge}, numerical uncertainties render Reynolds stresses difficult to
+determine accurately.  This difficulty is mitigated somewhat by the fact that previous global three
+dimensional studies have found Reynolds stresses to be small relative to gravitational stresses (see, for
+example, figure 5 of Lodato \& Rice 2004 and figure 10 of Boley et al. 2006).  Here we omit
+$T_{\varpi\phi}^{Reyn}$ from the calculation of $\alpha_{\rm{eff}}$ and show in section \ref{sec:converge}
+that, in fact, the gravitational torques are sufficient to account for all the mass transport.  We note that
+\cite{gammie2001} found Reynolds and gravitational stresses contributed equally to $\alpha_{\rm{eff}}$ in his
+study.  However, his was a local, shearing-sheet, 2D study.  As such, it was unable to capture global effects,
+such as those seen in the cited 3D studies.
 
@@
 To determine whether the converged value of $\alpha_{\rm{eff}}$ is consistent with predictions, we examine 
-the spatially (10 -- 40 AU) and temporally (12 -- 18 ORP) averaged  $\alpha_{avg}$,
+the spatially (10 -- 40 AU) and temporally (12 -- 18 ORP) averaged  $\alpha_{\rm avg}$,
 provided in table 1, for each value of  $l_{max}$. A Richardson extrapolation applied to these values  
-converges to $\alpha_{avg} = 0.017$ (over the same radial  and time range)
+converges to $\alpha_{\rm avg} = 0.017$ (over the same radial  and time range)
 as the grid spacing goes to zero.   This falls
-between the $\alpha_{avg}$ values for the Gammie curves 0.014 (strongly self-gravitating) and 0.027 (non-self-gravitating).
+between the $\alpha_{\rm avg}$ values for the Gammie curves 0.014 (strongly self-gravitating) and 0.027 (non-self-gravitating).
 
@@
 
-\subsection{Convergence and $l_{max}$}
+\subsection{Convergence and $l_{max}$}\label{sec:converge}
 
 The results above permit an assessment of appropriate azimuthal grids for the problem under consideration and
-similar problems. The $l_{max} = 512$ simulation yields a value for $\alpha_{avg}$ of 0.020. This
-overestimates the extrapolated converged value of $\alpha_{avg} = 0.017$ by $\sim 18$\%. The $l_{max} = 512$
+similar problems. The $l_{max} = 512$ simulation yields a value for $\alpha_{\rm avg}$ of 0.020. This
+overestimates the extrapolated converged value of $\alpha_{\rm avg} = 0.017$ by $\sim 18$\%. The $l_{max} = 512$
 simulation also overestimates the fractional power in the high-order ($m = 2$ -- 7) modes by essentially the
@@
 resolution has decreased for $l_{max} = 512$. From $l_{max} =64$ to 128 the fractional difference in
-$\alpha_{avg}$ is 36\%. This decreases to 26\% for the jump from $l_{max} = 256$ to 512. Furthermore, the
+$\alpha_{\rm avg}$ is 36\%. This decreases to 26\% for the jump from $l_{max} = 256$ to 512. Furthermore, the
 converged value represents infinite azimuthal resolution, which is, of course, impossible to achieve in a
@@
 %DICK: I made changes here.
-In addition to the convergence of $\alpha_{avg}$, table 1 illustrates the convergence of the measured net
+In addition to the convergence of $\alpha_{\rm avg}$, table 1 illustrates the convergence of the measured net
 inward mass transport rates $\dot{M}$ over the time interval 12 -- 18 ORP. We tabulate the value of
@@
 prior to the spatial average because $\dot{M}$ varies in sign with $r$. It is the magnitude of the typical net
-mass transport rates that we wish to compare. The $|\dot{M}|_{avg}$ measurements show convergence similar to
-that of the $\alpha_{avg}$ measurement based on the gravitational torques. This gives us confidence that no
+mass transport rates that we wish to compare. The $|\dot{M}|_{\rm avg}$ measurements show convergence similar to
+that of the $\alpha_{\rm avg}$ measurement based on the gravitational torques. This gives us confidence that no
 important contributions from Reynolds stresses are being overlooked. Moreover, we have used eq. (22) of
@@
 these torques are indeed sufficient to account for the measured $\dot{M}$.
+
+The principal reasons for not including the contributions from the Reynolds stress are twofold:
+
+1) As
+elucidated in the above paragraph, the gravitational torques are sufficient to account for the measured
+$\alpha$ and $\dot{M}$, and all three converge as resolution is increased. There does not appear to be an
+additional source of stress required to explain the observed mass transport.
+
+2) It is difficult to measure the Reynolds stress accurately in a nonlocal, strongly nonlinear calculation. The difficulty arises from the fact that the Reynolds
+stress is typically defined as 
+\begin{equation}
+\label{eq:reynstress}
+R_{ij} = \rho u'_i u'_j
+\end{equation}
+where $u'_i$ and $u'_j$ are the fluctuations in the velocity field. The fluctuating pieces of the velocity
+field are defined as $u'_i = u_i - u_i({\rm avg})$, but in a gravitationally unstable disk with extended
+non-axisymmetric spiral structures, the proper definition of the ``mean'' flow, $u_i({\rm avg})$, is unclear.
+
+In attempting to measure the Reynolds stresses, we tried using several different
+definitions for $u_{\phi}({\rm avg})$ and obtained results that varied dramatically. We used several local
+averaging schemes as well as an azimuthal average. Although Reynolds stresses are presented for other global
+simulations, \citep[e.g.][]{lodato2004}, the precise methods used to determine
+the stress values are not always elucidated. Because of this, it is difficult to make a direct comparison.
+The particulars of the method chosen can have a significant effect on the computed values of the
+Reynolds stress.
 
@@
 
-\subsection{Fragmentation and Effective $\alpha$s}
+\subsection{Fragmentation and Effective Alphas}
 
@@
 that  a gravitationally unstable disk is likely to fragment whenever the cooling time is less than half the local
-orbital period ($t_{cool}\Omega = 3$).
+orbital period ($t_{\rm cool}\Omega = 3$).
 This work was followed by Rice et al.~(2005), who used global 3D SPH simulations and found that the fragmentation limit was dependent on the adiabatic index.
@@
 orbital frequency and $\beta$ is some global constant.  This same cooling prescription was used in the Gammie and Rice
-et al.~studies.  The reason for the differences is unclear, and may not be simply a resolution effect (see Lodato \&
-Clarke 2011, arXiv:1101.2448).  For example, the \cite{gammie2001} have revealed the most stringent cooling criterion of
-all the $t_{cool}$ fragmentation studies, and, due to the local approximation Gammie used, these simulations are also at
-very high resolution.
-Moreover, \cite{mejia2005}, who used grid-based simulations with the same code we use here, and  \cite{boley2008}, who used radiative hydrodynamics simulations, also found results that are reasonably consistent with those of Gammie, Rice et al., and \cite{cossins2010}.
+et al.~studies.
+The reason for the differences is unclear, and may not be simply a resolution effect \citep{lodato2011}.
+For example, the \citeauthor{gammie2001} work revealed the most stringent cooling criterion of
+all the $t_{\rm cool}$ fragmentation studies, and, due to the local approximation Gammie used, these simulations are also at
+very high resolution. Moreover, \cite{mejia2005}, who used grid-based simulations with the same code we use
+here, and  \cite{boley2008}, who used radiative hydrodynamics simulations, also found results that are
+reasonably consistent with those of \citet{gammie2001}, \citet{rice2005}, and \cite{cossins2010}. Very recent
+work by \citet{paardekooper2011} supporting the \citeauthor{gammie2001} fragmentation criterion suggests that
+convergence results for fragmentation may be sensitive to the presence of sharp edges and the manner in which
+fragmentation is approached.
 
@@
 
-Overall, the disks simulated here are converging toward a state of self-regulation as the resolution is increased, and do not show signs of becoming more susceptible to fragmentation.
+Overall, the disks simulated here are converging toward a state of self-regulation as the resolution is
+increased, and do not show signs of becoming more susceptible to fragmentation.
 
@@
 
-We conducted a convergence study of a protostellar disk subject to gravitational instabilities to
-examine the distribution and amplitudes 
-of non-axisymmetric density structures arising from GIs 
-and the connection of these structures to gravitational torques,
-to the level which transport can be represented by a simple $\alpha$-disk formulation,
-and to fragmentation.
-A 3D hydrodynamics code with self-gravity
-was used to follow a disk subject to a constant global cooling time during its asymptotic state, when
-heating and cooling are roughly in balance.
-Four simulations were conducted, identical except for the number of azimuthal grid 
+We conducted a convergence study of a protostellar disk subject to gravitational instabilities to examine the
+distribution and amplitudes of non-axisymmetric density structures arising from GIs and the connection of
+these structures to gravitational torques, to the level which transport can be represented by a simple
+$\alpha$-disk formulation, and to fragmentation.  A 3D hydrodynamics code with self-gravity was used to follow
+a disk subject to a constant global cooling time during its asymptotic state, when heating and cooling are
+roughly in balance.  Four simulations were conducted, identical except for the number of azimuthal grid
 points, $l_{max} = $ 64, 128, 256, and 512, used in the calculations.
 
-A Fourier decomposition of the non-axisymmetric density structure 
-in $\cos(m\phi)$, $\sin(m\phi)$ was performed in order to
-characterize the amplitude distribution,  $A_m$,
-of azimuthal structures as a function of $m$.  
-This and previous studies have shown
-that disks in the asymptotic state have power at all resolvable $m$-values.
-The maximum value of  $m$ is given by
-$m_{max} = l_{max}/2$.  Therefore simulations with larger $l_{max}$ can spread the
-power over more Fourier components.
-We find that $\langle A_{tot} \rangle$, the power summed over all $m$, is roughly the same
-for all resolutions.
-However, the distribution of power among the Fourier components
-is distinctly affected by $l_{max} $.  
+A Fourier decomposition of the non-axisymmetric density structure in $\cos(m\phi)$, $\sin(m\phi)$ was
+performed in order to characterize the amplitude distribution, $A_m$, of azimuthal structures as a function of
+$m$.  This and previous studies have shown that disks in the asymptotic state have power at all resolvable
+$m$-values.  The maximum value of $m$ is given by $m_{max} = l_{max}/2$.  Therefore simulations with larger
+$l_{max}$ can spread the power over more Fourier components.  We find that $\langle A_{tot} \rangle$, the
+power summed over all $m$, is roughly the same for all resolutions.  However, the distribution of power among
+the Fourier components is distinctly affected by $l_{max} $.
 % $A_7$ is approximately equal for all resolutions.
-The fractional
-amplitude in low-order modes, $\langle A_{2-7} \rangle / \langle A_{tot} \rangle$,
-where $\langle A_{2-7} \rangle$ is 
-the sum over modes $m = 2$ through 7, 
-decreases monotonically with increasing resolution.  Over three quarters of $\langle A_{tot} \rangle$ resides
-in $m =$ 2 -- 7 for the $l_{max} = 64$ simulation, while approximately half resides in these $m$-values when $l_{max} =$ 512.
-A Richardson extrapolation suggests a limiting value of $\langle A_{2-7} \rangle / \langle A_{tot} \rangle \sim 0.43$ 
-as $l_{max}\rightarrow \infty$.
-The number of degrees of freedom available for power, $l_{max}/2$, increases linearly with the resolution.  Thus
-Fourier amplitudes are naturally spread from lower-order to higher-order as the available $m$-values increase.
+The fractional amplitude in low-order modes, $\langle A_{2-7} \rangle / \langle A_{tot} \rangle$, where
+$\langle A_{2-7} \rangle$ is the sum over modes $m = 2$ through 7, decreases monotonically with increasing
+resolution.  Over three quarters of $\langle A_{tot} \rangle$ resides in $m =$ 2 -- 7 for the $l_{max} = 64$
+simulation, while approximately half resides in these $m$-values when $l_{max} =$ 512.  A Richardson
+extrapolation suggests a limiting value of $\langle A_{2-7} \rangle / \langle A_{tot} \rangle \sim 0.43$ as
+$l_{max}\rightarrow \infty$.  The number of degrees of freedom available for power, $l_{max}/2$, increases
+linearly with the resolution.  Thus Fourier amplitudes are naturally spread from lower-order to higher-order
+as the available $m$-values increase.
 
@@
 
-The effective $\alpha_{\rm{eff}}$ was determined for each simulation under the assumption that
-Reynolds stresses are negligible relative to the gravitational stresses, as found in
-previous 3D studies.  
-As is the case for gravitational torques, $\alpha_{\rm{eff}}$  decreases with increased resolution.  
-Derived values of 
-$\alpha_{\rm{eff}}$ vs. radius for the highest resolution simulation, $l_{max} = 512$, 
-are consistent with
-predictions from an analytic local theory for thin disks by \citet{gammie2001}, as is the
-converged $\alpha_{\rm{eff}}= 0.017$, time-averaged over the asymptotic phase and spatially averaged 
-over the disk. The time and spatially averaged $\alpha_{\rm{eff}}$  
-obtained for the $l_{max} = 512$ simulations overestimates the converged value
-by less than 20\%.  Gravitational torques for this same model differ with converged values by a
-similar amount.
-Locally determined values of
-$\alpha_{\rm{eff}}$ vary by factors of 2 or 3 throughout the disk on timescales characteristic
-of the local dynamical time.  Thus the predictions of Gammie are only valid over many dynamic times when averaged over a substantial volume of the disk.
+The effective $\alpha_{\rm{eff}}$ was determined for each simulation under the assumption that Reynolds
+stresses are negligible relative to the gravitational stresses, as found in previous 3D studies.  As is the
+case for gravitational torques, $\alpha_{\rm{eff}}$ decreases with increased resolution.  Derived values of
+$\alpha_{\rm{eff}}$ vs. radius for the highest resolution simulation, $l_{max} = 512$, are consistent with
+predictions from an analytic local theory for thin disks by \citet{gammie2001}, as is the converged
+$\alpha_{\rm{eff}}= 0.017$, time-averaged over the asymptotic phase and spatially averaged over the disk. The
+time and spatially averaged $\alpha_{\rm{eff}}$ obtained for the $l_{max} = 512$ simulations overestimates the
+converged value by less than 20\%.  Gravitational torques for this same model differ with converged values by
+a similar amount.  Locally determined values of $\alpha_{\rm{eff}}$ vary by factors of 2 or 3 throughout the
+disk on timescales characteristic of the local dynamical time.  Thus the predictions of Gammie are only valid
+over many dynamic times when averaged over a substantial volume of the disk.
 
-Although not a goal of this work, our simulations permit
-comments on studies on fragmentation and clump formation.
-Most recent studies indicate that our disk should not fragment \citep[e.g., ][]{gammie2001, rice2005, mejia2005, boley2008}.   However, some recent studies \citep[e.g., ][]{boss2009, meru2011a}
-%Meru & Bate (2011), 
-suggest that fragmentation may be a common result of gravitational instabilities in disks.
-We find that as the resolution increases, our disk approaches a self-regulating state that can be reasonably well-described, in the time- and space-average sense, by an $\alpha$-disk parameterization.
-The average Toomre $Q$ increases with increasing resolution,
-suggesting that the disk is becoming less susceptible to clump formation with increasing resolution,
-consistent with the observation that
-the disk simulations do not show a trend toward denser spiral structure as the resolution is increased.
-Overall, the disks simulated here are converging toward a state of self-regulation as the resolution is increased, and do not show signs of becoming more susceptible to fragmentation.
+Although not a goal of this work, our simulations can be used to comment on studies of fragmentation and clump
+formation. Previous work indicates that our disk should not fragment due to the long cooling times that we
+use \citep[e.g.,][]{gammie2001, rice2005, mejia2005, boley2008}.
+However, some recent studies \citep[e.g.,][]{boss2009,meru2011a} suggest that previously determined fragmentation criteria too restrictive due to poor
+resolution. We find here that as the resolution is increased, our disk approaches a self-regulating state that
+can be reasonably well-described, in the time- and space-average sense, by an $\alpha$-disk
+parameterization. The average Toomre $Q$ also increases with increasing resolution, suggesting that the disk
+becomes less susceptible to clump formation as the resolution is increased.  This is consistent with our
+simulations not showing a trend toward denser spiral structure with increasing resolution. Moreover, recent
+work by \citet{paardekooper2011} showed that if a high cooling rate is approached slowly, the
+\citet{gammie2001} cooling criterion is recovered for all resolutions considered.  We do note that they find
+non-convergence with resolution when sharp edges in disks are present, but as the authors suggest, this may be
+numerically driven (see also \citet{lodato2011} on numerical convergences issues in regards to fragmentation). Overall, the disks simulated here are converging toward a state of self-regulation as the
+resolution is increased, and do not show signs of becoming more susceptible to fragmentation.
 
@@
 
-SM would like to thank NASA for the generous funding of the
-NASA Earth and Space Science Fellowship NNX07AU82H. This
-research was also supported in part by NASA Origins of Solar
-Systems grants NNG05GN11G and NNX08AK36G. The simulations
-and analysis herein were conducted on hardware generously
-provided by the Indiana University Information Technology Services
-and was supported by the National Science Foundation under
-Grant no. ACI-0338618l, OCI-0451237, OCI-0535258, CNS-
-0521433 and OCI-0504075. This work was also supported in part
-by Shared University Research grants from IBM, Inc., to the Indiana
-University.
+We would like to thank the anonymous referee for very useful comments that substantially improved this manuscript. SM
+would like to thank NASA for the generous funding of the NASA Earth and Space Science Fellowship
+NNX07AU82H. This research was also supported in part by NASA Origins of Solar Systems grants NNG05GN11G and
+NNX08AK36G. The simulations and analysis herein were conducted on hardware generously provided by the Indiana
+University Information Technology Services and was supported by the National Science Foundation under Grant
+no. ACI-0338618l, OCI-0451237, OCI-0535258, CNS- 0521433 and OCI-0504075. This work was also supported in part
+by Shared University Research grants from IBM, Inc., to the Indiana University.
+This work used the Extreme Science and Engineering Discovery Environment (XSEDE), which is supported by National Science Foundation grant number OCI-1053575
 
@@
 \multicolumn{5}{c}{$Q$, Fourier Amplitudes, $\alpha$, and $\dot{M}$}\\ \hline\hline
-$l_{max}$ & $Q_{avg}$   & $\langle A_{tot} \rangle$   & $\langle A_{2-7} \rangle / \langle A_{tot} \rangle$ &
-$\alpha_{avg}$ & $|\dot{M}|_{avg} ~\mathrm{M_\odot/yr}$\\\hline
+$l_{max}$ & $Q_{\rm avg}$   & $\langle A_{tot} \rangle$   & $\langle A_{2-7} \rangle / \langle A_{tot} \rangle$ &
+$\alpha_{\rm avg}$ & $|\dot{M}|_{\rm avg} ~\mathrm{M_\odot/yr}$\\\hline
 64   & 1.26 & 1.96 &  0.76 & 0.069 & 8.9$ \times 10^{-7}$ \\
@@
 \caption{ All quantities are time-averaged over 12 -- 18 ORP.
-$Q_{avg}$, $\alpha_{avg}$, and $|\dot{M}|_{avg}$  are also spatially
+$Q_{\rm avg}$, $\alpha_{\rm avg}$, and $|\dot{M}|_{\rm avg}$  are also spatially
 averaged over 10 -- 40 AU.  The entries for ``$\infty$'' are based on Richardson extrapolation,
@@
 \caption
-{Effective $\alpha$'s for the $t_{cool} =$ 2 ORP, $\gamma = 5/3$, $\Sigma \sim r^{ -1}$ disk, averaged between 12 and 18
+{Effective $\alpha$'s for the $t_{\rm cool} =$ 2 ORP, $\gamma = 5/3$, $\Sigma \sim r^{ -1}$ disk, averaged between 12 and 18
   ORP, for four different azimuthal grid resolutions, given by the number of azimuthal grid-points $l_{max}$. The
